### 2019-01-19 Articles/HamiltonianMechanicsIsConservationOfEntropy/titlepage.tex
Revision submitted to BJPS

\documentclass[11pt]{article}

\usepackage[margin=1.25in]{geometry}

\begin{document}

\title{Hamiltonian mechanics is conservation of information entropy}
\author{Gabriele Carcassi, Christine A. Aidala \\ University of Michigan}

\maketitle

\begin{abstract}
	In this work we show the equivalence between Hamiltonian mechanics and conservation of information entropy. We will show that distributions with coordinate independent values for information entropy require that the manifold on which the distribution is defined is charted by conjugate pairs (i.e.~it is a symplectic manifold). We will also show that further requiring that the information entropy is conserved during the evolution yields Hamilton's equations.
\end{abstract}

\section*{Acknowledgements}

We would like to thank Josh Hunt, David J. Baker and Gordon Belot for feedback. We would especially like to thank Mario Hubert: without his inspiration and guidance this work would have not been possible. This article is part of a larger project, Assumptions of Physics, that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.\footnote{See http://assumptionsofphysics.org/book for further details}

\end{document}