### 2020-04-01 JustTheMath/informationGeometry/informationGeometry.tex
Putting together elements of information geometry

\documentclass{article}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
% Remove line spaces between items of enumerate and itemize
\usepackage{enumitem}
\setlist{noitemsep}

\newtheorem{assump}{Assumption}
\renewcommand*{\theassump}{\Roman{assump}}

\newtheorem{axiom}[equation]{Axiom}
\newtheorem{defn}[equation]{Definition}
\newtheorem{prop}[equation]{Proposition}
\newtheorem{coro}[equation]{Corollary}
\newtheorem{thrm}[equation]{Theorem}

%\theoremstyle{definition}

\newenvironment{remark}{\emph{Remark}.}{}
\newenvironment{rationale}{\emph{Rationale}.}{\qed}
\newenvironment{justification}{\emph{Justification}.}{\qed}
\renewenvironment{proof}{\emph{Proof}.}{\qed}

\title{Just the math: information geometry}

\date{\vspace{-5ex}}
\begin{document}

\maketitle

\begin{abstract}
Short summary of information geometry.
\end{abstract}

\section{Parametric model}

\begin{defn}
	We call $\mathcal{S}$ the set of all possible probability density functions.
\end{defn}

\begin{remark}
	We can think of $\mathcal{S}$ as the set of function that are discontinuous at countably many points whose integral is unitary.
\end{remark}

\begin{defn}
	A \textbf{divergence} is a function $D : \mathcal{S} \times \mathcal{S} \to \mathbb{R}$ with the following properties:
	\begin{enumerate}
		\item $D(a, b) \geq 0$ for all $a,b \in \mathcal{S}$
		\item $D(a, b) = 0$ if and only if $a=b$
	\end{enumerate}
\end{defn}

\begin{remark}
	A divergence need not be symmetrical. A divergence need not satisfy the triangle inequality. A divergence is not a distance.
\end{remark}

\begin{defn}
	Given a manifold $M$, a \textbf{continuous parametric model} is a function $P_M : M \to \mathcal{S}$ that returns a probability density function (PDF) of a random variable $X$.
\end{defn}

\begin{defn}
	For a parametric model, we can define the diverge in terms of the point. That is, $D : M \times M \to \mathbb{R}$ where $D[a | b] = D(P_M(a), P_M(b)$ for all $a,b \in M$. Moreover, we can define left and right derivatives as follows:
	\begin{itemize}
		\item $D[\partial_i \cdot | \cdot] : TM \times M \to \mathbb{R}$ such that $D[ (\partial_i) \xi^i(a) | \xi^j(b)] = \partial_i D[\xi^i(a) | \xi^j(b)]$
		\item $D[\cdot | \partial_j \cdot] : M \times TM \to \mathbb{R}$ such that $D[ \xi^i(a) | (\partial_j) \xi^j(b)] = \partial_j D[\xi^i(a) | \xi^j(b)]$
	\end{itemize} 
\end{defn}

%\bibliographystyle{plain}
%\bibliography{bibliography}

\end{document}