### 2020-12-01 Articles/2019-QuantumComposition/Submissions/PhysLetters/tensor.tex
After Lorenzo's review

%ATTENZIONE: PER COMPILARE BISOGNA FARE A MANO
%dvipdf tensor.dvi ;mv tensor.pdf temp.pdf 
%Dovremmo forse aggiungere due parole sulle C* algebras: non sono stateintrodotte proprio per evitare il tensor product in infinitedimensional systems?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The four postulates of quantum mechanics are three
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[aps,prl,amsmath,amssymb,twocolumn]{revtex4}
%\documentclass[twocolumn,aps,showpacs,prl,groupedaddress]{revtex4}
%\usepackage{amssymb,stackengine}
\usepackage{amssymb}
\usepackage{color}
\usepackage{graphicx}
\usepackage{epsfig,amssymb,amsmath,amsthm}
%\usepackage[active]{srcltx}
%\usepackage[hypertex,linkcolor=red]{hyperref}

\usepackage{color}
\usepackage{cancel}
\usepackage{tikz-cd}

\theoremstyle{plain}
\newtheorem{thrm}{Theorem}[section]
\newtheorem{form}[thrm]{Formalization}
\newtheorem{prop}[thrm]{Proposition}
\newtheorem{lem}[thrm]{Lemma}

\theoremstyle{definition}
\newtheorem{defn}[thrm]{Definition}
\newtheorem{post}{Postulate}[]
\renewcommand*{\thepost}{(\alph{post})}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\newcommand{\blue}{\color{blue}}  %NON LINEAR
\newcommand{\red}{\color{red}}
\newcommand{\cyan}{\color{cyan}}
\newcommand{\green}{\color{green}} %THEORY quantum state estimation
\newcommand{\yellow}{\color{yellow}} %THEORY quantum channel estimation

\DeclareMathOperator{\spn}{span}

\newcommand{\pj}[1] {\underbar{$#1$}}
%\newcommand{\pj}[1] {\overline{#1}}

\def\>{\rangle}
\def\<{\langle}
\def\ca{_{\cal A}}
\def\cb{_{\cal B}}
\def\cc{_{\cal C}}
\def\comment#1{}
%\def\comment#1{ [{\bf Comment Lor:} {\sf #1}]}
\def\commentg#1{ [{\bf Comment Gabriele:} {\sf #1}]}
\def\labell#1{\label{#1}}
%\def\labell#1{\label{#1}{\mbox{{\tiny #1}}}}
%\def\section#1{{\par\em #1:--- }}
\def\togli#1{}
\def\sh{\mbox{sh}}
\def\iden{\openone}
\begin{document}

	%\fbox{{\scriptsize Preprint. \today}}
	\title{The four postulates of quantum mechanics are three}
	\author{Gabriele Carcassi$^1$, Lorenzo Maccone$^{2}$ and Christine A.
		Aidala$^1$ }\affiliation{\vbox{1.~Physics Department, University of
			Michigan, 450 Church Street, Ann Arbor, MI 48109-1040,
			United States}\\
		\vbox{2.~Dip.~Fisica and INFN Sez.\ Pavia, University of Pavia, via
			Bassi 6, I-27100 Pavia, Italy}}
	\begin{abstract}
		The tensor product postulate of quantum mechanics states that the
		Hilbert space of a composite system is the tensor product of the
		components' Hilbert spaces. All current formalizations of quantum
		mechanics that do not contain this postulate contain some equivalent
		postulate or assumption (sometimes hidden). Here we give a natural
		definition of composite system as a set containing the component
		systems and show how one can logically derive the tensor product
		rule from the state postulate and from the measurement postulate. In
		other words, our paper reduces by one the number of postulates
		necessary to quantum mechanics.
	\end{abstract}
	\pacs{}
	% Measurement theory (quantum mechanics), 03.65.Ta Mechanics quantum,
	% 03.65.-w noise quantum, 42.50.Lc quantum information, 03.67.Ac
	% Quantum information, 03.67.-a Quantum fluctuations, 42.50.Lc quantum
	% mechanics, 03.65.Ta quantum optics, 42.50.Gy
	\maketitle

	In this paper we derive the tensor product postulate (which, hence,
	loses its status of postulate) from two other postulates of quantum
	mechanics: the state postulate and the measurement postulate.  The
	tensor product postulate does not appear in all axiomatizations of
	quantum mechanics: it has even been called ``postulate 0'' in some
	literature \cite{zurek}. A widespread belief is that it is a direct
	consequence of the superposition principle, and it is hence not a
	necessary axiom. {\em This belief is mistaken}: the superposition
	principle is encoded into the quantum axioms by requiring that the
	state space is a {\em linear} vector space. This is, by itself,
	insufficient to single out the tensor product, as other linear
	products of linear spaces exist, such as the direct product, the
	exterior/wedge product or the direct sum of vector spaces, which is
	used in classical mechanics to combine state spaces of linear systems.
	These are all maps from linear spaces to linear spaces but they differ
	in how the linearity of one is mapped to the linearity of the others
	\footnote{For example, in the tensor product
		$a \otimes (b+c) = a \otimes b + a \otimes c$ while in the direct
		product $a \times (b+c) = a \times b + 0 \times c$ where $0$ is the
		zero vector. Also, in the tensor product
		$r (a \otimes b) = (r a) \otimes b = a \otimes (r b)$ while in the
		direct product $r (a \times b) = r a \times r b$, where $r$ is a
		scalar.}.  This belief may have arisen from the seminal book of
	Dirac \cite{diracbook}, who introduces tensor products (Chap.~20) by
	appealing to linearity. However, he adds the seemingly innocuous
	request that the product among spaces be distributive (rather,
	bilinear), which is equivalent to postulating tensor products (or
	linear functions of them). It is not an innocuous request. For example
	it does not hold where the composite vector space of two linear spaces
	is described by the direct product, e.g.~in classical mechanics, for
	two strings of a guitar: it is not distributive.  [General classical
	systems, not only linear ones, are also composed through the direct
	product.] Of course, Dirac is not constructing an axiomatic
	formulation, so his `sleight of hand' can be forgiven. In contrast,
	von Neumann (\cite{vonneumannbook} Chap.~VI.2, also \cite{jauch})
	introduces tensor products by noticing that this is a natural choice
	in the position representation of wave mechanics (where they were
	introduced in \cite{weyl,epr}), and then {\em explicitly postulates}
	them in general: ``This rule of transformation is correct in any case
	for the coordinate and momentum operators [...]  and it conforms with
	the [observable axiom and its linearity principles], we therefore
	postulate them generally.''  \cite{vonneumannbook}.  More mathematical
	or conceptually-oriented modern formulations
	(e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}) introduce this
	postulate explicitly.  An interesting alternative is provided in
	\cite{ballentinebook,ballentinepaper}: after introducing tensor
	products, Ballentine verifies a posteriori that they give the correct
	laws of composition of probabilities. Similarly, Peres uses
	relativistic locality \cite{peres}. While these procedures seemingly
	bypass the need to postulate the tensor product, they do not guarantee
	that this is the {\em only} possible way of introducing composite
	systems in quantum mechanics. In the framework of quantum logic,
	tensor products arise from some additional conditions \cite{matolcsi}
	which (in contrast to what is done here) are not connected to the
	other postulates.  In \cite{marmo,aerts} tensor products were obtained
	by specifying additional physical or mathematical requirements.
	\comment{ {\em Il resto della frase da spostare in supp material} In
		quantum field theory one tends to avoid problems connected with
		tensor products of infinite dimensional spaces by focusing on
		algebraic commutation structures, e.g.~\cite{giddins,roos}.  In
		particular, the recent MIP*=RE result \cite{mipre} implies that, in
		infinite dimensions, the tensor product is strictly less
		computationally powerful than the commutation structures,
		emphasizing the difference among these two structures, at least for
		the infinite-dimensional case. We will consider the non-relativistic
		setting here.}

	We start from the natural definition of a composite system as the set
	of two (or more) quantum systems. The composite system is therefore
	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing
		else}. The first key insight is that the first two postulates of
	quantum theory (introduced below) already assume that the preparation
	of one system is independent from the preparation of another
	(statistical independence). In fact, we cannot even talk about a
	system in the first place if we cannot characterize it independently.
	The second key insight is that, since all measurement outcomes and all
	states of the subsystems plus the composite live in the same
	probability space, we will automatically have a map $M$ that for a
	pair of subsystems' states gives us the composite state that
	corresponds to the statistically independent case. These insights are
	enough to characterize mathematically the state space of the
	composite: the linearity given by the Hilbert space, together with the
	fact that the composite system is fully described by the observables
	of $A$ and $B$, allows to extend the construction to the composite
	states that do not satisfy statistical independence (the entangled
	states). So the work consists of two interrelated efforts: a physical
	argument that starts from the first two postulates and leads to the
	necessary existence of the composition map $M$ and its properties
	together with a formal argument that shows how $M$ leads to the tensor
	product.

	\comment{mettere nel supplementary: We will mainly focus on
		kine\-mati\-cal\-ly-inde\-pen\-dent systems, namely no
		superselection rules or other restrictions to the state space are
		present: it is possible to prepare each subsystem of a composite
		system in a state that is independent of the other systems
		(preparation independence).  This is the only case in which the
		tensor product can be properly employed: the Hilbert space of
		composite systems that have restrictions is {\em not} the tensor
		product of the component spaces, but a subspace of it (e.g.~the
		anti-symmetric subspace for fermions). Typically this is ignored in
		the literature, since the tensor product formalism is very
		convenient and is often used also in these cases, and superselection
		rules are typically avoidable \cite{susskind,zanardi,zanardilloyd}.
	}

	This map $M$ acts on the state spaces of the subsystems. Each pure
	state is identified by a ray $\pj{\psi}$, a subspace of the system's
	Hilbert space spanned by all vectors $\psi$ differing by their
	(nonzero) modulus and phase: a two-dimensional subspace (a plane). In
	the same way, constraining the observable $X$ to a particular outcome
	value $x_0$ means identifying the subspace spanned by all
	non-normalized eigenvectors $|x_0\>$ of arbitrary phase such that
	$X | x_0 \> = x_0 |x_0\>$. As we need to map states and observables
	between subsystems and composite system, the map $M$ derived from the
	probability space is really a map between subspaces, namely the
	projective spaces. We then use projective geometry to show that the
	map $M$ on the projective spaces corresponds to a map $m$ on the
	vector spaces which is the tensor product.

	This is done by showing that preparation independence and statistical
	independence of the systems imply three conditions on the map $m$:
	(H1)~totality: the map is defined on all states of the subsystems;
	(H2)~bilinearity: the map is bilinear thanks to the fundamental
	theorem of projective geometry; (H3)~span surjectivity: the span of
	the image of map coincides with the full composite Hilbert space.  We
	then prove that, if these three conditions H1, H2 and H3 hold, then
	the map $m$ is the tensor product, namely the Hilbert space of the
	composite system is a tensor product of the components: the tensor
	product ``postulate'', which hence loses its status of a postulate. An
	overview of all these logical implications is given in
	Fig.~\ref{f:fig}. The rest of the paper contains the sketch of this
	argument, including all the physical arguments outlined above. The
	supplementary material contains the mathematical details.

	%%BoundingBox: 45 418 630 588
	%(%old BoundingBox: 45 382 649 547)
	\begin{figure}[ht]
		\epsfxsize=1.\hsize\leavevmode\epsffile{fig.eps}
		%\includegraphics[width=\linewidth, trim={0.2in 5.8in 2.5in 0.1in}, clip=true]{fig.eps}
		\caption{Schematic depiction of the logical implications used in
			this paper. FTPG stands for ``Fundamental Theorem of Projective
			Geometry''.  \label{f:fig}}\end{figure}

	We start with the axiomatization of quantum mechanics based on the
	following postulates
	(e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}): (a)~The pure state
	of a system is described by a ray $\pj{\psi}$ corresponding to a set
	of non-zero vectors $|\psi\>$ in a complex Hilbert space, and the
	system's observable properties are described by self-adjoint operators
	acting on that space; (b)~The probability that a measurement of a
	property $X$, described by the operator with spectral decomposition
	$\sum_{x,i}x|x_i\>\<x_i|/|\<x_i|x_i\>|$ ($i$ a degeneracy index),
	returns a value $x$ given that the system is in state $\pj{\psi}$ is
	$p(x|\psi)=\sum_i|\<\psi|x_i\>|^2/|\<x_i|x_i\>\<\psi|\psi\>|$ (Born
	rule). (c)~The state space of a composite system is given by the
	tensor product of the spaces of the component systems; (d)~The time
	evolution of an isolated system is described by a unitary operator
	acting on a vector representing the system state,
	$|\psi({t})\>=U_{t}|\psi({t}=0)\>$ or, equivalently, by the
	Schr\"odinger equation. The rest of quantum theory can be derived from
	these axioms. While some axiomatizations introduce further postulates,
	we will be using only (a) and (b) to derive (c), so the above are
	sufficient to our aims.

	\togli{This axiomatization implicitly contains a definition of
		``quantum system'' which is crucial for what follows, so we need to
		clarify the assumptions that it contains. We will use the following
		definition for a quantum
		system\togli{$\stackon[1pt]={\mbox{\tiny
					def}}$}$\stackrel{\mbox{\tiny def}}=${\em ``a quantum degree
			of freedom with $d$ (possibly discrete, or continuous, infinite)
			mutually exclusive (commuting) values for each of its properties.
			Its mathematical description is through a Hilbert space of
			dimension $d$ which contains all the states that describe the
			values of its possible properties. In accordance with the
			postulate (a), these values correspond to a basis of the space,
			given by the eigenvectors of the observable corresponding to that
			property''}. \commentg{We may have to revise to be more clear.
			Where is this used?} }

	Note that we limit ourselves to kinemati\-cal\-ly-inde\-pen\-dent
	systems, where all state vectors $|\psi\>$ in the system's Hilbert
	space $\cal H$ describe a valid state, {\em unconditioned on anything
		else}. We call this condition ``preparation independence'' and it
	should be noted that the tensor product applies only in this case. For
	example, the composite system of two electrons is not the tensor
	product, rather the anti-symmetrized tensor product, precisely because
	the second electron cannot be prepared in the same state of the first.
	We note that restrictions due to superselection rules arise either
	from practical (not fundamental) limitations on the actions of the
	experimenter \cite{susskind,zanardi,zanardilloyd} or from the use of
	ill-defined quantum systems. In the example above, the field is the
	proper quantum system and the electrons are its excitations.
	\footnote{We emphasize that the kinematic independence is inequivalent
		to dynamical independence (or isolation).  Indeed if two systems
		interact, their interaction may lead to dynamical restrictions in
		the state spaces.  Here we will not consider dynamical evolution,
		which is contained in postulate (d).}

	The definition of a composite system as containing {\em only} the
	collection of the subsystems means that any preparation of both
	subsystems independently must correspond to the preparation of the
	composite system. Since states are defined by postulate (a) as rays in
	the respective Hilbert spaces, there must exist a map
	$M : \pj{\cal A} \times \pj{\cal B } \to \pj{\cal C}$ that takes a
	pair of states for the subsystems ($\pj{\cal A}$ and $\pj{\cal B }$
	represent the projective spaces and the Cartesian product is the set
	of all possible pairs) and returns a state in the projective space
	$\pj{\cal C}$ for the composite. To visualize the geometrical meaning
	of $M$ directly within the Hilbert spaces, given a ray (a complex
	plane) in each of $\cal A$ and $\cal B$, $M$ returns a ray (a complex
	plane) in $\cal C$.  Our final goal will be to find a map
	$m:{\cal A}\times{\cal B}\to{\cal C}$ that acts on vectors in the
	Hilbert spaces $\cal A$, $\cal B$ and $\cal C$ consistently with $M$.
	Namely, $\pj{m(a,b)}=M(\pj{a},\pj{b})$ where the underline sign
	indicates the elements in the projective space. Again geometrically,
	$m$ takes a vector in each of $\cal A$ and $\cal B$, and returns a
	vector in $\cal C$ and we want this to be consistent with $M$ such
	that vectors picked from the same rays will return vectors in the same
	ray. We will prove that the map $m$ is the tensor product. We focus on
	pure states here: the argument can be extended to mixed states using
	standard tools \cite{ballentinebook}.

	The map $M$ must be injective: as said above, different states of the
	subsystems must correspond, by definition of composite system, to
	different states of the composite. Moreover, preparation independence
	implies that $M$, and hence $m$, must be total maps (condition H1):
	each subsystem of the composite system can be independently prepared
	and gives rise to a state of the composite.  H1 is not sufficient to
	identify the tensor product: by itself it does not even guarantee that
	the map $m$ is linear.

	Postulate (b) contains the connection between quantum mechanics and
	probability theory. It must then implicitly contain the axiomatization
	of probability, e.g.~see \cite{ballentinepaper,ballentinebook,cox}.
	One of the axioms of probability theory (axiom 4 in
	\cite{ballentinepaper}) asserts that the joint probability events $a$
	and $b$ given $z$ is $p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$.
	Consider $p(a \wedge b | \psi \wedge b)$ which represents the
	probability of measurement outcomes $a$ on system $A$ and $b$ on
	system $B$ given that system $A$ was prepared in $\psi$ and system $B$
	in $b$.  We have
	$p(a \wedge b | \psi \wedge b) = p(a | \psi \wedge b \wedge b) p(b |
	\psi \wedge b) = p(a | \psi \wedge b ) p(b | \psi \wedge b)$. The Born
	rule tells us that $p(a | \psi \wedge b) = |\<a|\psi\>|^2$ and that
	$p(b | \psi \wedge b) = |\<b|b\>|^2 = 1$, where $|a\>$, $|b\>$ are the
	normalized eigenstates relative to outcomes $a$ and $b$, and $|\psi\>$
	is the normalized state vector. We have:
	\begin{align}
	p(a\wedge b|\psi \wedge b)&=p(a|\psi)\\
	p(a\wedge b|a \wedge \phi)&=p(b|\phi)
	\end{align}
	In other words, since the probability for a measurement on one system
	depends only on its pure state, the Born rule requires that the
	measurement of one system is independent from the preparation of the
	other. We call this property ``statistical
	independence'' \footnote{One can also prove that the measurements on
		the components are independent as well (see Supplementary material),
		but we only strictly need preparation here.}.  It characterizes the
	map $M$, since $M(\pj{a}, \pj{b})$ corresponds to the composite state
	where $A$ and $B$ are prepared in the states $|a\>$ and $|b\>$.
	Define $M_b(\pj{a}) = M(\pj{a},\pj{b})$. From the Born rule we find
	\begin{align} &\Big|\Big\<M\left(\pj{a},\pj{b}\right)\Big|M\left(\pj{\psi},\pj{b}\right)\Big\>_{\cal C}\Big|^2
	=\Big|\Big\<M_b\left(\pj{a}\right)\Big|M_b\left(\pj{\psi}\right)\Big\>_{\cal C}\Big|^2
	\nonumber \\&
	=\left|\<a|\psi\>_{\cal A}\right|^2
	\labell{questa},
	\end{align}
	where the first and second terms contain the inner product in the
	composite space $\cal C$. [This is not a new assumption: it follows
	from the measurement postulate (b) for the composite system.] This
	means that, when one subsystem is prepared in an eigenstate of what is
	measured there, the state space of the other is mapped preserving the
	square of the inner product.  This implies orthogonality and the
	hierarchy of subspaces are preserved through $M_b$, making $M_b$ a
	colinear transformation by definition. Geometrically, recall that
	$M_b$ maps rays to rays. The fact that $M_b$ is colinear means that it
	also maps higher order subspaces to higher order subspaces (lines to
	lines, planes to planes, and so on) while preserving inclusion (if a
	line is within a plane, the mapped line will be within the mapped
	plane). In this case, the fundamental theorem of projective geometry
	\cite{fun} applies, which tells us that a unique semi-linear map $m_b$
	that acts on the vectors exists in accordance with $M_b$.  Moreover,
	conservation of probability further constrains it to be either linear
	or antilinear. This tells us that the corresponding $m$ is either
	linear or antilinear in the first argument. Namely, if equation
	\eqref{questa} holds, then
	\begin{align}
	\<a|\psi\>=\<m(a,b)|m(\psi,b)\>\labell{h2}\;
	\\\mbox{ or }
	\<a|\psi\>=\<m(\psi,b)|m(a,b)\> \labell{h2b}.
	\end{align}
	In this setting, the antilinear case \eqref{h2b} corresponds to a
	change of convention (much like a change of sign in the symplectic
	form for classical mechanics) and can be ignored.  Given a Hilbert
	space, in fact, we can imagine replacing all vectors and all the
	operators with their Hermitian conjugate, mapping vectors into duals
	$|\psi\>^\dag=\<\psi|$. These changes would effectively cancel out
	leaving the physics unchanged: the two equations $A|w\>=B|z\>$ and
	$\<w|A^\dag=\<z|B^\dag$ are equivalent.  (For example, in his first
	papers Schr\"odinger used both signs in his equation: effectively
	writing {\em two} equivalent equations with complex-conjugate
	solutions \cite{sch}). We can repeat the same analysis for the second
	argument of $m$ to conclude that it is a bilinear map, condition (H2).

	The last condition, span surjectivity (H3), follows directly from the
	definition of a composite system. Since it is composed {\em only} of
	the component systems, for any state $c$ of the composite system, we
	must find at least one pair $|a\>$, $|b\>$ such that
	$p(a\wedge b | c)\neq 0$.  Span-surjectivity follows: namely the span
	of the map applied to all states in the component systems spans the
	composite system state space. In other words, the composite does not
	contain states that are totally independent of (i.e.~orthogonal to)
	the states of the components.

	We have obtained the conditions H1, H2 and H3 from the state postulate
	(a), the measurement postulate (b) and the definitions of composite
	and independent systems. We now prove that these three conditions
	imply that the (up to now unspecified) composition rule $m$ is the
	tensor product. More precisely, given a total, span-surjective,
	bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ that maps the
	Hilbert spaces $\cal A$, $\cal B$ of the components into the Hilbert
	space $\cal C$ of the composite and that preserves the square of the
	inner product, we find that $\cal C $ is equivalent to
	$\cal A \otimes \cal B $ and that $m=\otimes$.

	Proof. Step 1: the bases of the component systems are mapped to a
	basis of the composite system. Because of totality property (H1) and
	because the square of the inner product is preserved, we can conclude
	that, given two orthonormal bases $\{|a_i\>\}\in{\cal A}$ and
	$\{|b_j\>\}\in{\cal B}$,
	$|\<m(a_i,b_j)|m(a_k,b_\ell)\>|^2=\delta_{ik}\delta_{j\ell}$, namely
	$\{|m(a_i,b_j)\>\}$ is an orthonormal set in $\cal C$.  Moreover, the
	surjectivity property (H3) guarantees that in $\cal C$ no vectors are
	orthogonal to this set. This implies that it is a basis for $\cal C$.

	Step 2: use the universal property. The tensor product is uniquely
	characterized, up to isomorphism, by a universal property regarding
	bilinear maps: given two vector spaces $\cal A$ and $\cal B$, the
	tensor product ${\cal A}\otimes{\cal B}$ and the associated bilinear
	map $T : \cal A \times \cal B \to {\cal A}\otimes{\cal B}$ have the
	property than any bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$
	factors through $T$ uniquely.  This means that there exists a {\em
		unique} $\hat i$, dependent on $m$, such that $\hat i \circ T=m$.
	In other words, the following diagram commutes:
	\begin{center}
		\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "\hat{i}"] \\
			& \mathcal{C}
		\end{tikzcd}
	\end{center}
	Since $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is a
	bilinear operator (property H2), thanks to the universal property of
	the tensor product we can find a unique linear operator
	$\hat{i} : \mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that
	$m(a, b) = \hat i(a \otimes b)$. The set $\{ \hat i(a_i\otimes b_j)$
	with $|a_i\>$ and $|b_j\>$ orthonormal bases for $\cal A$ and
	${\cal B}\}$ forms a basis for $\cal C$, since
	$\hat i(a_i\otimes b_j)=m(a_i,b_j)$ and we have shown above that the
	latter is a basis.  Thus,
	\begin{align} 
	&\<\hat i(a_i\otimes b_j)|\hat i(a_k\otimes b_\ell)
	\>_{\mathcal{C}}=\<m(a_i,
	b_j)|m(a_k,b_\ell)\>_\mathcal{C} \nonumber\\& =
	\delta_{ik}\delta_{j\ell}
	= \<a_i\otimes
	b_j| a_k \otimes b_\ell\>_{\otimes},
	\labell{ecco}\; 
	\end{align}
	where we used the orthonormality of the bases and the fact that
	$|a_i\otimes b_j\>$ is a basis of the tensor product space
	${\cal A}\otimes{\cal B}$. Since the function $\hat{i}$ is a linear
	function that maps an orthonormal basis of ${\cal A}\otimes{\cal B}$
	to an orthonormal basis of $\cal C$, $\hat{i}$ is a an isomorphism (a
	bijection that preserves the mathematical structure) between
	${\cal A}\otimes{\cal B}$ and $\cal C$. As
	${\cal C}\cong{\cal A}\otimes{\cal B}$ are isomorphic as Hilbert
	spaces, they are mathematically equivalent: we can loosely say that
	$\hat i$ is the identity. So we can directly use the
	tensor product to represent the composite state space. This means that
	the map $m : \cal A \times \cal B \to \cal C$ is equivalent to the map
	$\otimes : \cal A \times \cal B \to \cal A \otimes \cal B$ in the
	sense that $m \circ \hat i^{-1} = \otimes$.$\square$

	\togli{\comment{Commento finale da aggiungere?  E' interessante
			perche' tutto cio' ci dice che non otteniamo necessariamente il
			prodotto tensore, ma il prodotto tensore modulo una fase locale
			che e' fisicamente irrilevante. Viene da chiedersi se c'e' una
			qualche notazione (evidentemente piu' generale di quella di Dirac)
			che ci permetta di eliminare questa ambiguita' fisicamente
			irrilevante... La notazione di Dirac gia' elimina la necessita' di
			avere una rappresentazione per descrivere gli stati. Pero'
			evidentemente non elimina del tutto la ambiguita' di
			rappresentazione perche' un cambio di coordinate sia sugli stati
			che sugli operatori mi lascia invariata la fisica, ma non lascia
			invariata la notazione di Dirac...  Probabilmente la notazione che
			elimina l'ambiguita' e' quella che utilizza le matrici densita'
			invece dei vettori di stato (vedi Ozawa \cite{ozawa} e Holevo): le
			matrici densita' normalizzate sono phase-independent:
			$\rho=|\psi\>\<\psi|$. Attenzione: l'interpretazione di una
			matrice mista $\rho=\sum_ip_i|\psi_i\>\<\psi_i|$ come ``il sistema
			e' nello stato $|\psi_i\>$ con probabilita' $p_i$'' e' una
			CONSEGUENZA della regola di Born, quindi nella formalizzazione del
			postulato degli stati in termini di matrici densita', questa
			interpretazione non puo' apparire perche' e' una conseguenza di un
			altro postulato.}}

	\vskip 1\baselineskip A few comments on the proof: it is based on the
	universal property of the tensor product, which uniquely characterizes
	it. In step 1 we show that the bilinear map $m$ maps subsystems' bases
	into the composite system basis. We also know that there exists a
	tensor product map $ {T}=\otimes$ that can compose the vectors in
	$\cal A$ and $\cal B$.  In step 2 we use the universal property: since
	$m$ is a bilinear map, we are assured that there exists a unique
	$\hat i$ such that $\hat i \circ T=m$. Since we show that $\hat i$ is
	an isomorphism, then $\hat{i}$ bijectively maps vectors in $\cal C$
	onto vectors in the tensor product space. Namely $m={T}=\otimes$.

	We conclude with some general comments. The tensor product structure
	of quantum systems is not absolute, but depends on the observables
	that are accessible \cite{zanardi,zanardilloyd}. This is due to the
	fact that an agent that has access to a set of observables will define
	quantum systems differently from an agent that has access to a
	different set of observables. Where one agent sees a single system, an
	agent that has access to less refined observables (and is then limited
	by some superselection rules) can consider the same system as composed
	of multiple subsystems. \comment{Supplementary material? A typical
		example \cite{tellerbook} comes from quantum field theory. It is
		customary in basically all quantum optics literature to treat
		different modes of the radiation field (e.g.~the output of two
		lasers) as independent systems composed through the tensor product.
		Clearly the electromagnetic field is a single system and an agent
		who is able to access an optical mode that is a linear combination
		of the two will give a quantum description for it that cannot easily
		accommodate tensor products. Similarly, an agent can consider two
		electrons as two systems, joined with the tensor product, whenever
		they are distinguishable for all practical purposes (e.g.~the
		electrons are in widely separated physical locations). Yet, in
		principle, electrons are just excitations of a field, and the `true'
		quantum system is the field, not the single electrons
		\cite{teller,tellerbook}.  So, in quantum field theory, the quantum
		systems that should be joined through tensor products are the
		different fields and {\em not} the particles, which are just
		excitations (states) of the fields. In the words of Teller
		(\cite{tellerbook}, pg.22), tensor products can be safely used only
		if there is a ``primitive thisness'', which is captured in the
		definition of system.}

	It has been pointed out before that the quantum postulates are
	redundant: in \cite{masanes,hartle} it was shown that the measurement
	postulate (b) can be derived from the others (a), (c), (d). Here
	instead we have shown how the tensor product postulate (c) can be
	logically derived from the state postulate (a), the measurement
	postulate (b) and a reasonable definition of independent systems, and
	we have described the logical relations among them.  Of course, we do
	not claim that this is the {\em only} way to obtain the tensor product
	postulate from the others.

	{\it Acknowledgements:} L.M. acknowledges useful discussions with
	M.~Ozawa, P.~Zanardi, S.~Lloyd, D.~Zeh, G.~Auletta, A.~Aldeni and
	funding from the Attract project through the Eu Horizon 2020 research
	and innovation programme under grant agreement No 777222. G.C. and
	C.A.A. would like to thank M. J. Greenfield for reviewing the
	mathematical details and acknowledge funding from the MCubed program
	of the University of Michigan.

	\begin{references}
		\bibitem{zurek} W.H. Zurek, Quantum Darwinism, Nature Phys. {\bf 5},
		181 (2009).
		\bibitem{diracbook}P.A.M. Dirac, The principles of quantum mechanics,
		(Clarendon Press, Oxford, 1966).
		\bibitem{vonneumannbook}J. von Neumann, Mathematical Foundations of
		Quantum Mechanics (Princeton Univ.  Press, 1955).
		\bibitem{jauch}J.M. Jauch, Foundations of quantum mechanics
		(Addison-Welsey, 1968), pg.~176.
		\bibitem{weyl} H. Weyl, Gruppentheorie und Quantenmechanik (Hirzel,
		Leipzig, 1928); translated by H. P. Robertson, The Theory of Groups
		and Quantum Mechanics (Methuen, London, 1931); reprinted by Dover,
		p. 91.
		\bibitem{epr}A. Einstein, B. Podolsky, N. Rosen, Can
		quantum-mechanical description of physical reality be considered
		complete?, Phys. Rev. {\bf 47}, 777 (1935).
		\bibitem{ozawa}M. Ozawa, {Uncertainty relations for noise and
			disturbance in generalized quantum measurements}, Ann. Phys.  {\bf
			311}, 350 (2004).
		\bibitem{masanes}L. Masanes, T.D. Galley, M.P. M\" uller, The
		measurement postulates of quantum mechanics are operationally
		redundant, Nat. Commun. {\bf 10}, 1361 (2019).
		\bibitem{wootters}W.K. Wootters, Optimal Information Transfer and
		Real-Vector-Space Quantum Theory. In: Chiribella G., Spekkens R.
		(eds) Quantum Theory: Informational Foundations and Foils,
		Fundamental Theories of Physics, vol 181. Springer, Dordrecht
		(2016).
		\bibitem{nielsenchuang}M. A. Nielsen and I. L. Chuang, Quantum
		Computation and Quantum Information (Cambridge University Press,
		Cambridge, 2000).
		\bibitem{ballentinebook}L.E. Ballentine, Quantum Mechanics, a modern
		development (World Scientific, 2014).
		\bibitem{ballentinepaper}L.E. Ballentine, Probability theory in
		quantum mechanics, Am. J. Phys. {\bf 54}, 883 (1986).
		\bibitem{peres}A.~Peres, Classical interventions in quantum systems.
		II. Relativistic invariance, Phys. Rev. A {\bf 61}, 022117 (2000).
		\bibitem{matolcsi} T. Matolcsi, Tensor product of Hilbert lattices and
		free orthodistributive product of orthomodular lattices, Acta Sci.
		Math. (Szeged), {\bf 37}, 263 (1975).
		\bibitem{marmo} F.M. Ciaglia, A. Ibort, G. Marmo, On the Notion of
		Composite System, In: Nielsen F., Barbaresco F. (eds) Geometric
		Science of Information, Lecture Notes in Computer Science, vol
		11712. Springer (2019).
		\bibitem{aerts} D. Aerts, I. Daubechies, Physical justification for
		using the tensor product to describe two quantum systems as one
		joint system, Helv. Phys. Acta {\bf 51}, 661 (1979).
		\bibitem{susskind}Y. Aharonov, L. Susskind, Charge superselection
		rule, Phys. Rev. {\bf 155}, 1428 (1967).
		\bibitem{zanardi}P. Zanardi, Virtual Quantum Subsystems, Phys. Rev.
		Lett {\bf 87}, 077901 (2001).
		\bibitem{zanardilloyd} P. Zanardi, D.A. Lidar, S. Lloyd, Quantum
		Tensor Product Structures are Observable Induced, Phys. Rev. Lett.
		{\bf 92},060402 (2004).
		\bibitem{cox}R.T. Cox, The Algebra of Probable Inference (J. Hopkins
		press, 1961).
		\bibitem{fun} E. Artin: Geometric algebra, Interscience Publishers Inc
		(1957)
		\bibitem{sch}E. Schr\"odinger, Annalen der Physik {\bf 102}, 81
		(1926); English translation in E. Schr\"odinger, {\em Collected
			papers on Wave Mechanics} (Blackie \& Son, London, 1928).
		\bibitem{hartle}J.B. Hartle, Quantum Mechanics of Individual Systems,
		Am. J.  Phys. {\bf 36}, 704 (1967).
	\end{references}

\end{document}

\begin{prop}[Mixed states]
	Up to now we have only considered pure states, described by rays
	$\pj{\psi}$ or vectors $|\psi\>$ in the Hilbert space.  Using
	standard techniques we can extend the analysis to more general mixed
	states, described by a density matrix $\rho$. 
\end{prop}

Using the Born rule (which, we remind, is one of the postulates we
employ), it is simple to show that a density matrix (i.e.~a positive,
trace one operator) can be given the interpretation of the system
being in a pure state with some probability. Indeed, any density
matrix can be always written as $\rho=\sum_ip_i|\psi_i\>\<\psi_i|$,
where the $|\psi_i\>$ are not necessarily orthogonal and where
$p_i\geqslant 0$ sum up to one: $\sum_ip_i=1$. We now show that, a
system which is in a state $|\psi_i\>$ with probability $p_i$ can be
described by a density matrix $\rho$ \cite{nielsenchuang}.

Consider an arbitrary observable with eigenstate $|a\>$. The Born rule
says that, the probability of finding the related eigenvalue $a$
given that the system is in a state described by a normalized vector
$|\psi_i\>$, is $p(a|i)=|\<\psi_i|a\>|^2$. But if the system is in
state $|\psi_i\>$ with probability $p_i$, a trivial application of the
Bayes rule, marginalizing over $i$, tells us that
\begin{align} &p(a)=\sum_ip(a\wedge
i)=\sum_ip(a|i)p_i=\sum_ip_i|\<\psi_i|a\>|^2
\nonumber\\&=
\mbox{Tr}[|a\>\<a|\sum_ip_i|\psi_i\>\<\psi_i|]=\mbox{Tr}[|a\>\<a|\rho]
\labell{bo}\;.
\end{align}
As is customary, we interpret the last term as the Born rule for
density matrices, which then says that a system described by the
density matrix $\rho$ is equivalent (thanks to the arbitrariness of
$|a\>$) to a system described by a state $|\psi_i\>$ with probability
$p_i$. Note that the decomposition $\rho=\sum_ip_i|\psi_i\>\<\psi_i|$
of a mixed state $\rho$ into pure states $|\psi_i\>,p_i$ is not
unique, except for the pure states themselves where one of the $p_i$
is one.  This is an interesting peculiar characteristic of quantum
mechanics.

In the discussion above, we have not specified the Hilbert space where
the vectors live. If we suppose that $|\psi_i\>$ are vectors in the
tensor product Hilbert space of two subsystems, then the discussion we
have given for pure states in our paper trivially extends to mixed
states: namely we can conclude that the mixed states of composite
systems are density matrices (positive operators with trace one) that
act on the tensor product Hilbert space.

\vskip 1\baselineskip