### 2020-12-23 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Adding section on density/riemmanian geometry

\documentclass[letterpaper]{article}

\usepackage{amsmath, hyperref, enumerate, comment}
\usepackage{amsfonts, esint}
\usepackage{graphicx}
\usepackage{mathrsfs} %for script font

\usepackage[backend=biber, style=authoryear, url=false, doi=false, eprint=false]{biblatex}
%\usepackage[backend=biber,  style=authoryear-icomp, url=false, doi=false, eprint=false]{biblatex}
\addbibresource{bibliography.bib}

\AtEveryBibitem{%
  \clearfield{note}%
  \clearfield{series}%
  \clearfield{pages}%
}

\newcommand{\diff}{\mathop{}\!\mathrm{d}} %defines exterior derivative 
\renewcommand{\div}[1]{\gv{\nabla} \cdot #1} % for divergence
\newcommand{\curl}[1]{\gv{\nabla} \times #1} % for curl
\renewcommand{\vector}[1]{\ensuremath{\vec{#1}}} % redefines vector to vec
\newcommand{\integral}{\int}

\begin{document}

\title{From physical conditions to Hamiltonian and Lagrangian mechanics}
\author{TBD}

\date{\today}

\maketitle

\begin{abstract}
	TBD
\end{abstract}

\tableofcontents

\section{Introduction}
\label{introduction}

% TODO: Lagrangian assumption vs kinematic assumption. Investigate what happens when the transformation is non-linear.

Philosophers have recently shown that we can interpret Lagrangian and Hamiltonian mechanics as theoretically equivalent in a precise mathematical sense. Within a subclass of classical mechanical systems known as the \textit{hyper-regular domain}, the two formalisms are categorically equivalent \parencites[]{Teh}{Barrett2}. Categorical equivalence provides a natural interpretive criterion for theoretical equivalence because it yields mutual inter-translatability between theory formulations. At first glance, this equivalence motivates the view that there are no physically significant differences between these two formalisms, at least within the hyper-regular domain. Yet, it allows for physical and conceptual differences to arise outside the hyper-regular domain. 
%Once identified, these differences can even affect how we interpret hyper-regular systems.

Within this wider domain, \textcites[]{North} has argued for privileging Hamiltonian mechanics, on the grounds that it requires less mathematical structure to formulate. In contrast, \textcites[]{Curiel} recommends privileging Lagrangian mechanics, arguing that it isolates physically-significant kinematical constraints that all classical systems must satisfy. Here, we will argue that there are indeed physical grounds for privileging Hamiltonian mechanics over Lagrangian mechanics, their categorical equivalence notwithstanding. Whereas North's argument relies on a problematic criterion for comparing structure---based on the size of symmetry groups \parencites[]{Swanson}---our argument avoids this pitfall. We show that a system needs to satisfy strictly fewer physical conditions in order to be Hamiltonian. To be Lagrangian, systems must satisfy an additional physical condition, making Lagrangian mechanics a proper subset of Hamiltonian mechanics. Hamiltonian mechanics thereby captures a wider range of systems than Lagrangian mechanics, and this difference has interpretive consequences even within the hyper-regular domain. Furthermore, the construction at the heart of this argument provides an interpretation of classical mechanical systems in terms of their invariant phase space density, along with a novel interpretation of the principle of stationary action within the Hamiltonian formalism.

Understanding the relationship between Hamiltonian and Lagrangian mechanics depends partly on how one defines ``classical systems." Different definitions can result in different interpretive relationships. For instance, restricting classical systems to the hyper-regular domain precludes certain interpretive considerations. To avoid verbal disputes, it is essential to clarify what falls within the scope of ``classical systems." One way to delimit classical systems is to specify a set of equations that they satisfy---such as Hamilton's equations or the Euler--Lagrange equations. We will instead follow a different approach by delimiting classical systems based on the physical conditions that they satisfy. Section~\ref{infinitesimal} introduces the condition of \textit{infinitesimal reducibility}, meaning that the state of the whole system is equivalent to specifying the states of its parts (and vice versa). Infinitesimal reducibility distinguishes classical from quantum systems, where the latter are irreducible (we cannot specify the whole state by specifying the states of its subsystems). In the widest sense, a classical system is any infinitesimally reducible system.
%a reducible system comprises subsystems whose states jointly determine the state of the whole system.

Section~\ref{deterministic} introduces the joint condition of deterministic and reversible evolution. We will show that a classical system which satisfies this condition is Hamiltonian: its phase space is a symplectic manifold that evolves according to Hamiltonian evolution  (i.e. according to a symplectomorphism). Following a deterministic and reversible evolution is equivalent to being a conservative system, i.e. one that conserves energy. Hence, our discussion of Hamiltonian mechanics necessarily excludes dissipative systems, i.e. ones that do not conserve energy. Likewise, our derivation of Lagrangian mechanics in Section~\ref{motion} will also take place within the context of conservative systems. Consequently, we will not consider systems that satisfy inhomogenous Euler--Lagrange equations. We, of course, do not deny the sense in which the Lagrangian formulation applies to dissipative systems, but this is not what we will mean by ``Lagrangian mechanics" \parencites[]{Smith}{Curiel}.\footnote{Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems, although he seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.} Hence, our claim about the fundamentality of Hamiltonian mechanics is restricted to the non-dissipative regime. We will view dissipative systems as Newtonian, but neither Lagrangian nor Hamiltonian.

% General decomposition of conservative and non-conservative: https://scholar.harvard.edu/files/rsyuan/files/ijbc.pdf

%Further motivation for focusing on non-dissipative systems: dissipative systems contain more information than contained in the Lagrangian alone, namely, the additional information contained in the Rayleigh potential. Or, as Curiel puts it, the systems involve a ``force $1 $-form that is not exact and that is not Lie-derived along the Lagrangian vector field solving the inhomogeneous system" [311].

We begin in Section~\ref{privileging} with an informal overview of our argument for privileging Hamiltonian mechanics. We introduce the key physical conditions that distinguish Hamiltonian from Lagrangian mechanics. Section~\ref{fundamentality} explains how our position differs from North's. Specifically, our argument does not require commitments to fundamental structure or perfectly natural properties. Section~\ref{derivation} describes in detail the derivation of Hamiltonian mechanics for systems that satisfy two physical conditions: infinitesimal reducibility and deterministic/reversible evolution. We then show (Section~\ref{motion}) that Lagrangian systems must satisfy an additional physical condition, which we call motion equivalence. Section~\ref{density} explains how the invariant phase space density plays a special role in the Hamiltonian framework. In contrast, we show that in configuration space, the invariance of the density is coordinate-dependent (arising from the fact that the Jacobian equals unity in Cartesian coordinates but not in general). Finally, Section~\ref{Lagrangian} considers the status of Lagrangian mechanics within our framework. We provide a geometric interpretation of the action principle, which applies even outside the hyper-regular domain. We also consider and rebut objections stemming from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.
% In configuration space, it is in some sense an accident that the density is invariant

%Note that the photon does not belong to the hyper-regular domain. Furthermore, if you are not in the hyper-regular domain, then there is no longer a one-to-one map between position and velocity.

%It seems as though North is explicitly not assuming that her argument applies only to the hyper-regular domain [page 85, especially footnote 50]. Part of her argument involves motivating the idea that we will not in general be able to construct an equivalent Lagrangian formulation, and this involves appealing to systems outside the hyper-regular domain.
% To-do: Figure out how North defines classical systems. It might be in terms of the equations of motion that hold for classical particles [87]. North is arguing that ``symplectic structure is a simpler, more natural, more general underlying structure" that ``suffices for" this theory of classical mechanics [87]. Seems similar to GC position.

% In the introduction, should we give glosses on hyper-regular, infinitesimally reducible, deterministic/reversible, and motion equivalence? Otherwise, discussion might be less accessible.

\section{Privileging Hamiltonian mechanics}
\label{privileging}

%[To do: Do we agree with the following claim defended by North: ``as far as we can tell, we need only symplectic structure to do classical mechanics" [74]. If so, is our position in tension with the more general formulation of Lagrangian mechanics noted by Swanson and Halvorson, in terms of the almost tangent structure of the tangent bundle? [See their footnote 9, pages 4-5]

In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must clarify what we mean by ``fundamental." Of the various notions of fundamentality, we will mean something entirely non-mysterious: a formalism is more fundamental provided that it is logically more general. In this case, the Lagrangian framework is less general because it requires that an additional condition hold, compared to the Hamiltonian framework. Hence, the Lagrangian framework is less fundamental because it is a proper subset of the Hamiltonian framework.

We now outline the three main steps of our argument, summarized in Figure~\ref{diagram}. The left column lists three basic physical conditions: infinitesimal reducibility, determinism/reversibility, and motion equivalence. Each condition entails physically significant properties, described in the middle column. Each physical property corresponds to a formal mathematical framework depicted in the right column: including phase space, Hamiltonian mechanics, and Lagrangian mechanics. In this way, our framework develops a correspondence between physical conditions and their associated mathematical formalisms. We thereby gain a straightforward interpretation of the physical significance of these formalisms.
%Of the various notions of fundamentality, there is at least the following non-mysterious definition:

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{Diagram.png}
\caption{Physical conditions and corresponding properties and formalisms}
\label{diagram}
\end{figure}

First, a system is \textit{infinitesimally reducible} provided that the states of its constituent, infinitesimal parts completely specify the overall state of the system. Colloquially, we can say that the whole state is equal to the sum of its parts, where we take these infinitesimal parts to be classical particles. When this condition is satisfied, we can define a distribution $\rho$ over particle states. Furthermore, since particle states are independent of a choice of units, this distribution must be invariant. We show in Section~\ref{infinitesimal} that to define this invariant distribution, we require a symplectic manifold, which serves as the phase space of classical mechanics. The physical significance of this symplectic manifold is thereby its connection with the physical condition of infinitesimal reducibility. [To do: Clarify the scope of our treatment: is it restricted to $N$-particle classical mechanics? Or does it also handle rigid bodies and classical continual mechanics? Cite Mark Wilson 2013 on these issues]. 
% % I think I could spend more time clarifying reducibility, infinitesimal. Maybe define separately first?

Secondly, a system has \textit{deterministic and reversible evolution} provided that the present state of the system uniquely determines both its future and past states. This condition entails that the invariant distribution $\rho$ is in fact a conserved phase space density. We can thus model such systems using Hamiltonian mechanics, defined on a phase space with a symplectomorphism. These two conditions thereby characterize Hamiltonian mechanics as the mathematical framework for infinitesimally reducible systems with deterministic and reversible evolution. Section~\ref{deterministic} justifies this correspondence. [To do: on the topic of failures of determinism, cite Earman  \textit{a primer on determinism}, Baez singularities paper, or Norton's dome paper].

Finally, some classical systems satisfy a further condition, which we call \textit{motion equivalence}. Motion equivalence holds when the system's trajectory in physical space determines---and is determined by---its evolution in phase space. In other words, knowledge of the trajectory in physical space (e.g. Euclidean three-space $\mathbb{R}^3$) suffices for knowledge of the trajectory in phase space, and vice versa. Physically, knowledge of the trajectory in physical space corresponds to knowledge of position and velocity, while knowledge of the trajectory in phase space requires knowing the system's momentum. Hence, motion equivalence entails a correspondence between velocity and momentum, which formally leads to the framework of Lagrangian mechanics in the hyper-regular domain (Section~\ref{motion}). As a result, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.

Importantly, some systems fail to satisfy the condition of motion equivalence. It is not always true that we can determine the evolution of a system in phase space from its trajectory in physical space. This is because knowing a system's position and velocity is not in general sufficient for knowing its momentum. For instance, knowledge of the physical trajectory of a photon, when treated as a particle, is insufficient for knowledge of its phase space evolution, since we cannot determine the momentum of a photon from only its trajectory through space. In general, different Lagrangians---differing even by more than a constant---can correspond to the same particle trajectories while not agreeing on the total energy, even up to a constant.\footnote{\textcites[1185--1186]{Barrett2} discusses this in the context of the failure of equivalence between the vector field formulations of Lagrangian and Hamiltonian mechanics. In contrast, the trajectories in phase space (encoded by the Hamiltonian vector field) uniquely define the Hamiltonian up to a constant.} Hence, the set of classical systems that satisfy motion equivalence is a proper subset of the classical systems that satisfy the conditions for Hamiltonian mechanics. For at least this reason, Hamiltonian mechanics is more fundamental than Lagrangian mechanics.
[To do: verify that the scenario of two different Lagrangians---giving rise to the same trajectories in physical space with different energies---arises only when the motion equivalence condition fails. See Curiel 2014, page 42 for his take on this kind of scenario.]

%Figure~\ref{diagram} summarizes our characterization of Hamiltonian and Lagrangian mechanics, in terms of physical conditions and corresponding mathematical structures. Section~\ref{derivation} justifies the logical connection between these physical conditions and corresponding mathematics. Before delving into this derivation, Section~\ref{fundamentality} discusses its philosophical significance.

% One way to motivate the motion equivalence condition: in general, the same trajectory in physical space can correspond to different trajectories in phase space: knowledge of the position and velocity is not sufficient for knowledge of the momentum (e.g. if you do not know the mass of the system). Formally, different Lagrangians can have the same Lagrangian vector field but give rise to Hamiltonians with different Hamiltonian vector fields, as Barrett notes \parencites*[1186]{Barrett2}. In contrast, if two Hamiltonian systems agree on the Hamiltonian vector field, then they must agree on the Hamiltonian, up to an arbitrary constant (whereas two Lagrangian systems can have the same Lagrangian vector field but not agree on the Lagrangian or total energy, even up to a constant)).

% Overall summary: A classical system is an infinitesimally reducible system. The state space of these systems can be modeled by a symplectic manifold (the physical assumption of infinitesimal reducibility corresponds to the mathematical structure of symplectic manifolds). Classical Hamiltonian systems satisfy the additional conditions of having deterministic and reversible evolutions. These two physical assumptions entail the existence of a symplectic morphism on the state space manifold, making it a contact manifold. Classical Lagrangian systems satisfy a further assumption: keeping track of the trajectories in physical space is equivalent to tracking the evolution of states in phase space.

\subsection{(Non-metaphysical) fundamentality and structure}
\label{fundamentality}

% North argues that ``the structure of physics is the (minimal) structure that is assumed among dynamically equivalent---in classical mechanics: canonically transformed---formulations of a theory. Structure comprises those quantities that remain intact when we transform one allowable coordinatization of the laws into another" [78]

% North's epistemic method for figuring out the structure: ``look at what is invariant under allowable transformations"; this tells us what the structure is [73].

As indicated at the start of Section~\ref{privileging}, we use ``fundamentality" in a non-metaphysical fashion. On our usage, differences in fundamentality amount to differences in generality.\footnote{Of course, we allow for other non-metaphysical uses of ``fundamental," tracking differences in epistemic or conceptual primacy. For instance, we can think of the canonical position and momentum observables as ``epistemically fundamental" because they suffice to completely specify a classical state, fixing its values for all other observables. \textcites[31, 200]{Ruetsche} refers to this as ``fundamental in the physicist's sense." Section~\ref{density},discusses another sense in which the Hamiltonian formalism is more fundamental than Lagrangian mechanics.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires that fewer physical conditions are satisfied. Compared with North's \parencites*[]{North} account of structure, our definition has two distinct advantages. First, it avoids metaphysical disputes surrounding perfectly natural or joint-carving properties, which empiricists and pragmatists decry. Secondly, it avoids a technical objection that Swanson and Halvorson \parencites*[]{Swanson} levy against North's more precise mathematical criterion for comparing structure.

Methodologically (but not metaphysically), our approach has many parallels to North's. We agree with North that it is fruitful to identify both invariants and---more generally---``the mathematical structure needed to formulate the theory in an invariant, frame-independent way" \parencites*[65]{North}. Additionally, we agree that it is worthwhile to identify what particular mathematical structure is required by a formulation \parencites[78]{North}. Indeed, we undertake precisely this sort of investigation in Section~\ref{derivation}. However, North advocates a metaphysical account of structure, about which we remain agnostic. According to North, ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description" \parencites*[66]{North}. These notions of fundamentality and intrinsicality correspond to differences in joint-carving, with more fundamental or intrinsic features being more joint-carving, or in Lewis's \parencites*[]{Lewis1983} terminology, more natural.

Owing to our epistemic inabilities to access these putative joints, we do not endorse these metaphysical dimensions of structure or fundamentality. While we will argue that Hamiltonian mechanics is significant partly because it elucidates invariant distributions (Section~\ref{density}), this does not entail that additional mathematical structure is physically insignificant or surplus structure. Additional structure may correspond to an additional physically-significant assumption that applies to the system under consideration. As we will see, precisely this scenario occurs in the context of Lagrangian mechanics. Hence, whereas North views Lagrangian mechanics as containing ``excess, superfluous structure" \parencites*[75]{North}, we take this additional structure as unnecessary but still physically significant. It corresponds to the physical assumption of motion equivalence, which applies to many classical systems. 
%This characterizes the relationship between Hamiltonian and Lagrangian formalizations of a classical system. 
% classical mechanics also deploys physically significant representations that are not intrinsic to mechanical systems [to do: mention some examples, e.g. in a footnote. The Lagrangian? Cartesian coordinates?].

% Emphasize contrast: whereas North is arguing that we should believe that the state space of Hamiltonian mechanics corresponds to that of a classical world (using Occam's razor; inference to the best explanation), we are not trying to figure out which state space structure to believe in more. 
% As Swanson and Halvorson put it, North's argument presupposes ``that statespace structure actually corresponds to some kind of real, physical structure. Many physicists and philosophers of science view statespace as a formal tool for efficiently encoding certain modal commitments of scientific theories, and aid for making calculations and predictions, but not part of the theory's ontology. North's position presupposes some form of statespace realism" [1]. North assuming that there is a fact of the matter regarding the state space of the world.

Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real" fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure as part of the theory's ontology, but as formal device for encoding physical information or even merely as a calculational tool.} Whether Hamiltonian phase space structure is metaphysically fundamental does not matter for the questions we seek to answer. Regardless of its metaphysical status, Hamiltonian mechanics is physically more fundamental, and we will argue for its epistemic and conceptual primacy as well. [To do: only keep preceding remark if somewhere later in the essay, we elucidate these epistemic and conceptual dimensions.] Indeed, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [mention these three formalisms and provide references]. Each of these frameworks has different pragmatic costs and benefits. Hence, it is plausible that there may be no single, unique necessary mathematical structure for mechanics, but rather a family of related structures. There might not be a single canonical minimal structure for formalizing the relevant physical models and physical assumptions. Furthermore, each of these formalisms references some ``unphysical" coordinates or representational means, whose physical significance is borne out by their fruitfulness in modeling. 
%The mathematical structure that you ``need" might not be unique. Neither one of these three formalizations of Hamiltonian mechanics is necessarily more ``fundamental" or joint-carving than the others. Nevertheless, there are compelling physical and pragmatic motivations for preferring the extended phase space formulation of Hamiltonian mechanics. In this formulation, we gain an invariant Hamiltonian that is the conjugate variable of the proper time, while the normal Hamiltonian (which is not coordinate-invariant) still plays the role of generating time-translations.

North also introduces a technical notion of comparative structure, based on the symmetry group of statespace. According to this criterion, statespace A has more structure than statespace B if the symmetry group of A is a proper subgroup of B's symmetry group \parencites[87-88]{North}.\footnote{This statement follows the charitable reconstruction of North's criterion given by \textcites[]{Swanson}. North sometimes speaks in terms of comparing the dimensions of symmetry groups, which they note leads to trivial counterexamples.} A smaller symmetry group entails that it is in some sense ``harder" to preserve the space's structure, indicating that it has more mathematical structure to be preserved. These symmetry groups are defined in terms of the coordinate transformations on the space that preserve geometric structure, amounting to re-coordinatizations (redescriptions) of that structure. For Lagrangian systems, North takes the relevant symmetry group to be that of the point transformations, and for Hamiltonian systems, she assumes that canonical transformations play this role. Since point transformations are a proper subgroup of canonical transformations, North's criterion entails that the statespace of Lagrangian mechanics has more structure than the statespace of Hamiltonian mechanics---in virtue of the former having a smaller symmetry group.

However, Swanson and Halvorson \parencites*[]{Swanson} develop a series of problems for North's argument. Most compellingly, they argue that we cannot in general assume that point transformations and canonical transformations are the proper symmetry groups for Lagrangian and Hamiltonian mechanics, respectively. As they note, Galilean boosts are often symmetry transformations of Lagrangian systems (in the sense of preserving the equations of motion), but they are not point transformations. Additionally, the two-body (Kepler) problem admits canonical transformations that map physically distinct solutions to each other (specifically, solutions whose orbits possess the same energy but different eccentricity). Hence, we cannot straightforwardly compare the structure of Lagrangian and Hamiltonian mechanics in terms of the symmetry groups of their statespaces.

Advantageously, our criterion for comparing Hamiltonian and Lagrangian mechanics avoids the criticisms that Swanson and Halvorson develop against North. For the sense in which Hamiltonian mechanics has less structure than Lagrangian mechanics is simply one of logical strength: a system must satisfy strictly fewer physical conditions in order to be Hamiltonian. Whereas to be Lagrangian, the system must also satisfy the assumption of motion equivalence. The assumption of motion equivalence places an additional constraint on the Hamiltonian, a constraint that need not be satisfied in general.

Nevertheless, based on the equivalence between the tangent bundle and cotangent bundle formulations of Lagrangian and Hamiltonian mechanics, one might wonder whether there could be a ``correspondingly general formulation of Lagrangian mechanics" that is analogous to the symplectic but-not-cotangent-bundle formulation of Hamiltonian mechanics \parencites[1182]{Barrett2}. For, if there were, then this formulation of Lagrangian mechanics might be more general than the cotangent bundle formulation of Hamiltonian mechanics. And thus, there would be an analogous sense in which Lagrangian mechanics is more fundamental than Hamiltonian mechanics, undercutting our argument. Fortunately, it is implausible that there could be a suitable generalization of Lagrangian mechanics. In going from the cotangent bundle formulation of Hamiltonian mechanics to the symplectic manifold formulation, we lose some globally-defined structure. This is analogous to moving from the space of globally-Euclidean manifolds to the space of locally-Euclidean manifolds. Yet, it is not clear how Lagrangian mechanics could remain well-defined while throwing away global structure from the tangent bundle. For the action seems to be inherently a globally-defined quantity, as an integral defined over a trajectory.

[To do: figure out best place to slip in the following background information, which is relevant for the preceding objection:] Barrett shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics are categorically equivalent provided that we define categories of models over the tangent and cotangent bundle formulations, respectively \parencites*[1181-82]{Barrett2}. However, he also shows that if we consider the more general formulation of Hamiltonian mechanics on a symplectic manifold (which does not require a cotangent bundle), then the two formulations are not categorically equivalent \parencites*[1182-83]{Barrett2}. This parallels the construction of Hamiltonian mechanics in Section~\ref{derivation}.

% Plausibly, move the following directly after our claim that our notion of fundamentality/structure comparison avoids the objections of Swanson/Halvorson against North's:

% Way to motivate that the motion equivalence assumption further constrains the Hamiltonian: the Hamiltonian already defines a relationship between conjugate momentum and velocity (through ($dq/dt = \partial H / \partial p$) ). Motion equivalence provides a further constraint on this relationship, and hence a further constraint on the Hamiltonian.

% As defined, our notions of fundamentality and structure are not metaphysical in nature.

%Finally, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [mention these three formalisms and provide references]. Each of these frameworks has different pragmatic costs and benefits. Hence, unlike North, we are comfortable acknowledging that the mathematical structure that you ``need" might not be unique. There might not be a single canonical minimal structure for formalizing the relevant physical models and physical assumptions. Neither one of these three formalizations of Hamiltonian mechanics is necessarily more ``fundamental" or joint-carving than the others. And in some sense, each of these formalisms references some ``unphysical" coordinates or representational means. Nevertheless, there are compelling physical and pragmatic motivations for preferring the extended phase space formulation of Hamiltonian mechanics. In this formulation, we gain an invariant Hamiltonian that is the conjugate variable of the proper time, while the normal Hamiltonian (which is not coordinate-invariant) still plays the role of generating time-translations.
% One con of this extended phase space formulation: the Hamiltonian constraint is both a constraint and the generator of time evolution; hence, it must be defined in the neighborhood of the actual states, so the values it provides outside of these neighborhoods should be unphysical, yet they encode the equations of motion.
% Should we remark on the claim that the only structure we need is whatever is sufficient to define densities locally? Hence, we need just the symplectic form but not the cotangent bundle?
% E.g. in the extended phase space formulation, the equations of motion come from derivatives of the Hamiltonian constraint over the variables, but this constraint is defined on unphysical points. These unphysical points are plausibly a mathematical artifact, although they supply a nice tool/method for understanding physically significant aspects of the representation.

% One way to motivate the motion equivalence condition: in general, the same trajectory in physical space can correspond to different trajectories in phase space: knowledge of the position and velocity is not sufficient for knowledge of the momentum (e.g. if you do not know the mass of the system). Formally, different Lagrangians can have the same Lagrangian vector field but give rise to Hamiltonians with different Hamiltonian vector fields, as Barrett notes \parencites*[1186]{Barrett2}. In contrast, if two Hamiltonian systems agree on the Hamiltonian vector field, then they must agree on the Hamiltonian, up to an arbitrary constant (whereas two Lagrangian systems can have the same Lagrangian vector field but not agree on the Lagrangian or total energy, even up to a constant)).

% Claim to think about: GC was arguing on 10-12 that you could handle classical systems in the Hamiltonian framework with just a differential manifold and no symplectic structure. This would contradict North's indispensability claim that Hamiltonian symplectic structure is necessary for treating classical systems. E.g., North footnote 31: ``The symplectic form is crucial to determining the motion, since the same Hamiltonian can induce different flows for different symplectic forms" [76]. North takes this as an argument for why we cannot eliminate the symplectic form in an attempt to get at even less structure. We need it to define the dynamics.

\section{Deriving Hamiltonian and Lagrangian mechanics}
\label{derivation}

% [To do: figure out how to flag to the reader that they can skip/skim some parts of this derivation section without loss of understanding for later sections]. E.g., `Readers who are more interested in the significance of invariant phase space densities or the geometric action principle can feel at peace skipping ahead.' But, presumably, part of the stuff on invariant densities is relevant for Section~\ref{density}? Figure out how to finesse this

This section justifies the physical and mathematical relationships in Figure~\ref{diagram}.
Sections~\ref{infinitesimal}-\ref{deterministic} show how to derive the formalism of Hamiltonian mechanics from the assumptions of infinitesimal reducibility and deterministic/reversible evolution. Section~\ref{motion} shows how systems that satisfy the further condition of motion equivalence can be treated within the formalism of Lagrangian mechanics. We thus demonstrate in detail our methodology of deriving  relevant mathematical structures from underlying physical conditions.

% Is it worth emphasizing/highlighting any particularly important parts of the derivation up front?

\subsection{Infinitesimal reducibility}
\label{infinitesimal}

First, we will show that a classical phase space arises if and only if a system satisfies the following physical condition:

\begin{quotation}
\noindent
\textit{Infinitesimal reducibility}: the state of the system is reducible to the state of its infinitesimal parts. That is, giving the state of the whole system is equivalent to giving the state of its parts, which is equivalent to giving the states of those parts, and so on.
\end{quotation}

\noindent
This assumption characterizes classical systems as those whose infinitesimal parts completely specify the state of the system. [At least in a footnote: comment on how this characterization of infinitesimal particles relates to standard philosophy of physics interpretive disputes over the status of point particles in classical mechanics. See Butterfield Pointillisme article; Butterfield characterizes particles as extensionless.]

Let $\mathcal{C}$ be the state space for the whole system. Any point $c \in \mathcal{C}$ represents a particular state of the whole system. We define a classical \textit{particle} as the limit of recursive subdivision. A classical system then consists of a collection of classical particles. Let $\mathcal{S}$ be the state space for the particles, so that each point in $\mathcal{S}$ describes a possible state of a classical particle. We assume that this state space $\mathcal{S}$ is a manifold. Then for each state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts. That is, the state of the whole system is a distribution over its infinitesimal parts. This distribution describes the proportion of classical particles in any particular state. Formally, if $\mathcal{S}$ is a manifold, $\int_U \rho_c d\mathcal{S}$ gives us the fraction of the system whose parts are in the region $U$ of the state space $\mathcal{S}$. 
%[is region U in the state space C or S. I assume in S.].

We define the \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state. Formally, $\xi^a: \mathcal{S} \to \mathbb{R}^n$, yielding a real-valued vector for any given particle state, with $n$-many degrees of freedom.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates,' but we will reserve that terminology for the state variables that define units.} We will define a \textit{coordinate} $q \in \xi^a$ as a particular state variable that defines a unit.  For instance, $q$ might be position, defining units of meters. In general, the coordinates are the subset of the state variables that define the unit system, needed to measure properties of the system.

On physical grounds, the distribution $\rho_c$ must transform as both a density and a scalar function. As a function defined over the state variables, the distribution transforms as a density, i.e. $\rho_c$ changes as a function of the Jacobian determinant that relates the new coordinates to the old coordinates. As a function defined over the points in phase space, the distribution transforms as a scalar, i.e. $\rho_c$ is invariant under all coordinate transformations. Hence, the Jacobian determinant must in fact equal one. In other words, $\rho_c$ is a function only of the particle state $s$ and does not depend on the choice of coordinates.\footnote{Formally, we can express this argument as a chain of equalities: $\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$, where $\rho_c(\xi^a)$ and $\rho_c(\hat{\xi}^b)$ represent the same distribution as expressed over the different coordinates and $s(\xi^a)$ and $s(\hat{\xi}^b)$ represent the same particle state as expressed by the different coordinates. Note that $\rho_c(s(\hat{\xi}^b))$ shows how the distribution depends on the coordinates only through the particle state.} We will now show that in order for the distribution to satisfy these two constraints, the classical phase space must be at least a symplectic manifold. In other words, having symplectic manifold structure is a necessary condition for the distribution over particle states to transform as a density and scalar function.
%This is because the distribution is a function only of the states, rather than of the particular choice of coordinates. 
 % Classical phase space (i.e. a symplectic manifold) is the only space that satisfies these constraints on the distribution. 
%(in manifold terminology, these are known as coordinates)

The argument begins with the simplest case: suppose that a single coordinate $q $ suffices to define the unit system for the particle state space. Then, the state variables are given by a set $\xi^a = \{ q, k^i \}$, where \textit{prima facie} the index $i$ can range over more than one number. We will show that, in fact, there can only be a single $k^i$ variable in this case, conjugate to the single $q$ coordinate. Assuming that the particle density $\rho_c$ is a scalar, it follows that we can arbitrarily change the unit defined by $q$ without changing the density $\rho_c$. 
% We further assume that $q$ fully defines the units for all state variables. [Is there a way to physically motivate this assumption, beyond this being a constitutive role of the position-like state variables?]

Now, suppose that we change the units by transforming the coordinate $q$ to a new coordinate $\hat{q}$. Since---by assumption---$q$ fully defines the unit system, $\hat{q}$ must do so as well, making it a function just of the old coordinate: $\hat{q}=\hat{q}(q)$. Call the new units $\hat{\xi}^b = \{ \hat{q}, \hat{k}^i\}$. As shown above, the Jacobian determinant $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must equal one. Assume for contradiction that there is only one state variable, so that there are no $k^i$ variables. Then we would have $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right| = 1$. But this would mean the unit change cannot be arbitrary. Therefore we must have two or more state variables, so that there must be at least one $k^i$ variable.

\begin{comment}
%Longer justification for the Jacobian determinant equaling one.

 As defined, $\rho_c$ is a function from the particle state space $\mathcal{S}$ to the reals. We can define $\rho_c$ on the units by taking advantage of an invertible function $s: \mathbb{R}^n \to \mathcal{S}$ mapping a particular value of the units $\hat{\xi}^b$ to a point in the particle phase space $\mathcal{S}$. Hence, by definition $\rho_c(\hat{\xi}^b):= \rho_c(s(\hat{\xi}^b))$. [Is this function $s$ the inverse of the coordinate mapping $\xi^b$?]. Then, since the physical state is independent of the coordinates, the state $s(\hat{\xi}^b)$ is the same state as $s(\xi^a) $ defined in terms of the original units $\xi^a$. Again, by definition, $\rho_c(s(\xi^a))=\rho_c(\xi^a) $. Since the distribution transforms as a density, it follows that $\rho_c(\xi^a)$ must equal the product of the Jacobian determinant $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| $ times the density defined over the transformed units: $\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$. We can summarize the argument so far as a chain of equalities:

\begin{equation*}
\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)
\end{equation*}

\noindent
Therefore the Jacobian $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must equal one.

\end{comment}

Next, we show that in fact there must be exactly one $k$ variable, conjugate to the single $q$ variable. Using the fact that the Jacobian determinant equals one, we determine how the state variables $k^i$ must change in order to compensate the change from $q$ to $\hat q$. Since $\hat{q}$ depends only on $q$, $\frac{\partial \hat{q}}{\partial k^a} = 0$. Hence, the Jacobian is an upper triangular block matrix:

\begin{equation}
%\begin{aligned}
|J| = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \begin{vmatrix}
\frac{\partial \hat{q}}{\partial q} & \frac{\partial \hat{q}}{\partial k^a} \\
\frac{\partial \hat{k}^b}{\partial q} & \frac{\partial \hat{k}^b}{\partial  k^a} 
\end{vmatrix} = \begin{vmatrix}
\frac{\partial \hat{q}}{\partial q} & 0 \\
\frac{\partial \hat{k}^b}{\partial q} & \frac{\partial \hat{k}^b}{\partial  k^a}
\end{vmatrix}
%\end{aligned}
\end{equation}

\noindent
The determinant of a triangular block matrix equals the product of determinants of the block matrices along the diagonal: $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| =  \left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^b}{\partial k^a}\right|$. For the Jacobian determinant to equal one, we must therefore have $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| =   \left|\frac{\partial \hat{q}}{\partial q}\right|^{-1}$. Since $\hat q$ is a function only of $q$, $\frac{\partial \hat{q}}{\partial q}$ is a total derivative, so we obtain the inverse by swapping the numerator and denominator. Hence, we see that $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| = \left|\frac{\partial q}{\partial \hat q} \right|$. 
%  $\left|\frac{\partial \hat{q}}{\partial q}\right|^{-1} = \left|\frac{\partial q}{\partial \hat q} \right|$.

This equation provides only a single constraint, placed on the determinant of the transformation from $k^a$ to $\hat{k}^b$. Hence, if there were three or more $\xi^a$ variables, this constraint would be insufficient to recover the transformation of the $k^a$ uniquely. Yet in this case, $\hat q$  would not fully define the units for all other state variables, which is contrary to assumption. Consequently, there must be exactly two state variables: $q$ and a single $k$. So in this case of a single $q$ variable, coordinate-independent areas and densities can only be defined on a two-dimensional manifold.

\begin{comment}
\begin{equation}
\left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^b}{\partial k^a}\right| - \left|0\right| \left|\frac{\partial \hat{k}^b}{\partial q}\right|
= \left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^b}{\partial k^a}\right|.
\end{equation}
 so $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right|\left|\frac{\partial \hat{k}^b}{\partial k^a} \right|$: 
\end{comment}

To generalize, we say two coordinates are \textit{independent} if changing the units for one does not change the units for the other. Suppose our particle state space $\mathcal{S}$ is such that its units are fully defined by $n$ independent coordinates $q^i$. Suppose you change the first coordinate $q^1$ without changing the others. Then we will find a variable $k_1$ that changes as before so that the densities are invariant. Now change the second coordinate $q^2$ in the same way while also fixing $k_1$. Then we find a corresponding $k_2$. In the end, we will find that $\mathcal{S}$ is $2n$-dimensional, and the state variables are $\xi^a = \{ q^i, k_i \}$. We define an \textit{independent degree of freedom} as the space charted by a pair of such variables.

We can now show that $\mathcal{S}$ must have at least the structure of a symplectic manifold. Invariant densities can only be assigned to infinitesimal areas. Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface, so that we can define integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$, for $\Sigma \subset \mathcal{S}$. Because degrees of freedom are---at least locally---independent, the total number of states in a volume is the product of the possible configurations of each degree of freedom. This means the volume form is given by $\omega^n$. Importantly, the volume form cannot be degenerate because otherwise there would be a `volume' in state space that could not contain states. But such a region of state space would---by definition---not be a volume. Therefore the form $\omega$ itself cannot be degenerate (for otherwise, we would have states that are not counted by it). Next, we show that $\omega$ must be closed, using the fact that degrees of freedom are independent. If we translate a surface along an independent degree of freedom, this does not change the area of the surface, meaning that the number of states on that surface is invariant. Consider a parallelepiped, whose opposite sides have equal areas. Taking the integral over this closed surface (over the parallelepiped), opposite sides provide equal and opposite contributions. Hence, the integral over this closed volume must equal zero. This shows that the 2-form $\omega$ must be closed. Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate, and therefore $\mathcal{S}$ is a symplectic manifold. By convention, we set $\omega = \hbar dq^i \wedge dk_i = dq^i \wedge dp_i$ where $p_i = \hbar k_i$.
% % the `motivation' section of the wiki page on symplectic manifolds provides a different style of derivation for the 2-form being non-degenerate and closed.

%since the degrees of freedom are independent, the number of states on a surface does not change if we translate that state across independent degrees of freedom. If we imagine a parallelepiped, the integral over the surface must be zero (the integrals over opposite sides are equal and opposite) [explain/justify this more clearly. Is this coming from the two-form being antisymmetric?]. 
%
%if degenerate form, would have states that are not counted by the form.

The conclusion is that, under the infinitesimal reducibility assumption, a state $c$ of the system is a coordinate-invariant distribution over the state space $\mathcal{S}$ of its infinitesimal parts. If the state space of the particles is a manifold, then it must be a symplectic manifold: symplectic manifold structure is necessary for the distribution to be coordinate-invariant. 
% Which is the only type of manifold that supports coordinate-invariant distributions

\subsection{Deterministic and reversible evolution}
\label{deterministic}

Next, we assume that the classical systems of interest are deterministic and reversible. This amounts to satisfying the following joint condition:

\begin{quotation}
\noindent
\textit{Deterministic and reversible evolution assumption}: given the present state of the system, all future (determinism) and past (reversible) states are uniquely identified.

\end{quotation}

[Should we footnote that this condition leaves out some classical systems, e.g. Norton's dome, Earman's space invaders?]

In our case, this applies to both the state of the whole system and to the state of its parts. Let $\lambda: \mathbb{R} \to \mathcal{S}$ be the evolution over time of the state of a particle. Assuming deterministic and reversible evolution, the initial state $s_0 = \lambda(t_0)$, at the initial time $t_0$, uniquely identifies the state of the system. Moreover, if $\rho(\lambda(t_0), t_0)$ is the density associated to the initial state at the initial time, we must have that $\rho(\lambda(t), t) = \rho(\lambda(t_0), t_0)$, i.e. the density remains the same throughout the evolution. In other words, all particles that begin in a given initial state must end up in the same final state and vice-versa.  
%[could two different particles occupy the same initial state?].
%[what exactly does `initial particle' mean? Do we mean the initial state of the composite system or just a particular particle?]

Now consider the integral $\int_{\Sigma} \rho_c \omega(d\Sigma)$. The region $\Sigma$ will be mapped to a new region, but the fraction of the system found in both the old and new regions must be the same. Since both the integral and $\rho_c$ must remain constant during the evolution, the two-form $\omega$ must also remain the same. That is, the areas in phase space must be mapped to areas of equal size, and independent degrees of freedom must be mapped to independent degrees of freedom (otherwise, volumes would not be mapped to equal volumes, which would violate determinism). The evolution is therefore a symplectomorphism and corresponds to Hamiltonian evolution. Intuitively, this argument is the inverse of Liouville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes (i.e. sets of states are mapped to sets of states of equal size) which leads to Hamiltonian evolution. In summary, we have shown that any system that satisfies the conditions of both infinitesimal reducibility and deterministic/reversible evolution is a classical Hamiltonian system.

\subsection{Motion equivalence}
\label{motion}

Having recovered classical Hamiltonian mechanics, the laws of evolution are further restricted when the following condition holds:

% [to do: make the footnote regarding connection to Marsden and Abraham more precise. Might want to flag that their theorem is expressed in terms of the Lagrangian vector field and corresponding Hamiltonian vector field (obtained via Legendre transformation) having the same base integral curves. These base integral curves encode how particles evolve over time. Should I also elaborate on how one can interpret motion equivalence in terms of the theories having the same empirical content? This is how Barrett interprets the Marsden and Abraham theorem 3.6.2]
\begin{quotation}
\noindent
\textit{Motion equivalence assumption}: the motion of the system (i.e. trajectories in physical space-time) suffices to recover its dynamics (i.e. evolutions in state space) and vice-versa.\footnote{ \textcites[]{Abraham1978}, Theorem 3.6.2 shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics describe the same evolution of particle positions over time (see \textcites[1180-1181]{Barrett2} for discussion). Motion equivalence can be understood as the converse of this theorem: assuming that Lagrangian and Hamiltonian mechanics agree on particle evolutions, we derive that the Legendre transformation is well-defined in the hyper-regular domain.}
\end{quotation}

If we focus on the parts of the system, this means that for every evolution in phase space there should be one and only one trajectory in physical space. Note that each space variable $x^i$ is a coordinate, i.e. a state variable that defines a unit. In fact, the trajectories can be fully described by these units and only these units. For, to specify a trajectory, we need epistemic access to a unit system, where the number of independent units is determined by the dimensionality of physical space. It follows that $x^i = q^i$, each $q^i$ is paired with a conjugate $p_i$, and each state $\{q^i, p_i\}$ corresponds to one and only one trajectory in physical space. At each point $x^i$, then, there must be infinitely many trajectories, one for each possible combination of $\{p_i\}$. Since these trajectories are differentiable in $x^i$, we can define a velocity $v^i = d_t x^i$. If the equations of motions were such that $v^i=v^i(q^i)$, then motion equivalence would fail as the full trajectory would be determined only by the $q^i$. Hence, we must have $v^i=v^i(q^j, p_k)$, so that knowledge of the state in phase space suffices for knowledge of the trajectory. If motion equivalence holds, then this relationship must be invertible, meaning that knowledge of the trajectory suffices for knowledge of the phase space state. For systems that satisfy the motion equivalence condition, then, the following relationships hold at any given time:
\begin{equation}
\begin{aligned}
x^i &= q^i \\
v^i &= d_t x^i = v^i(q^j, p_k)
\end{aligned}
\end{equation}

In this case---which we call \textit{weak equivalence}---$v^i(q^j, p_k)$ is invertible at every $q^i$. Therefore, for fixed q, the Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible. Since $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j} = \frac{\partial}{\partial p_i} (\frac{\partial H}{\partial p_j})$ equals the Jacobian $\frac{\partial v^i}{\partial p_j}$. Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere, which we take to be positive. These are exactly the cases where a Lagrangian can be constructed, and where the Lagrangian leads to a unique solution (via the stationary action principle). Hence, weak equivalence suffices for the existence of a hyper-regular Lagrangian.

Next, we show that this Lagrangian describes massive particles under scalar potentials. If we look at the whole distribution (so that $q$ is no longer held fixed), we have $\rho(q^i, p_j) = |J| \rho(x^i, v^j) = \left|\frac{\partial v^i}{\partial p_j}\right| \rho(x^i, v^j)$ since
\begin{equation}
\begin{aligned}
|J| &= \begin{vmatrix}
\frac{\partial x^i}{\partial q^j} & \frac{\partial x^i}{\partial p_j} \\
\frac{\partial v^i}{\partial q^j} & \frac{\partial v^i}{\partial p_j}
\end{vmatrix}
= \begin{vmatrix}
\delta^i_j & 0 \\
\frac{\partial v^i}{\partial q^j} & \frac{\partial v^i}{\partial p_j}
\end{vmatrix} \\
&= \left|\delta^i_j\right| \left|\frac{\partial v^i}{\partial p_j}\right| 
= \left|\frac{\partial v^i}{\partial p_j}\right|.
\end{aligned}
\end{equation}
Note that while the value given by $\rho(q^i, p_j)$ is coordinate independent, the value given by $\rho(x^i, v^j)$ depends on the choice of coordinate through $\left|\frac{\partial v^i}{\partial p_j}\right|$. If $x^i$ truly sets the unit system by itself, then $\left|\frac{\partial v^i}{\partial p_j}\right|$ must be a function of position only. Similar considerations will also hold for marginal distributions (i.e. distributions on a subset of the coordinates), which means all components of $\frac{\partial v^i}{\partial p_j}$ must be a function of position only. We set:
\begin{equation}
\begin{aligned}
\frac{\partial v^i}{\partial p_j} = \frac{1}{m} g^{ij} \\
\frac{\partial p_j}{\partial v^i} = m g_{ji}
\end{aligned}
\end{equation}
where $m$ is the unit conversion constant between velocity and conjugate momentum, while $g_{ij}$ represents the linear dependency of the velocity on the momentum.

If we integrate, we have:
\begin{equation}
\begin{aligned}
v^i = \frac{1}{m} g^{ij}(p_j - A_j) \\
p_j = m g_{ji} v^i + A_j
\end{aligned}
\end{equation}
where $A_j$ are arbitrary functions. Note that:
\begin{equation}
v^i = d_t q^i = \frac{\partial H }{\partial p_i} = \frac{1}{m} g^{ij}(p_j - A_j) \\
\end{equation}
We integrate yet again and find:
\begin{equation}
H = \frac{1}{2m} (p_j - A_j) g^{ij}(p_j - A_j) + V \\
\end{equation}
where V is another arbitrary function. We recognize this as the Hamiltonian for massive particles under potential forces.

[To do: Connect this with Curiel equation 6.2 for the Hamiltonian.]

\section{The physical significance of invariant densities}
\label{density}

* ``Fundamentality" because it makes the physics more prominent. Some people use coordinate-free/independence as necessary conditions for fundemantaly/physical significance. This means coordinates do not appear in the definition. Yet, densities require coordinate for the definition, therefore cannot be coordinate-free/independent, therefore there are some physically significant quantities that do not satisfy that. Why fundamental??? (reverse mathematics? fundamental because it recovers the correct math from the physical assumption?)

---
Section on densities and riemmanian geometry. (Gabriele)

* A paragraph to show how densities work. In general are coordinate dependent and you can't simply compare values at different points. This is the case in Lagrangian mechanics because both differentials in position vary controvariantly.

* Contrast this with densities over position and velocity. But! We can understand some aspects of Riemmanian geometry in terms of how states densities are represented over position and velocity. Hyper-regularity allows us to express densities from "states" to "kinematic variables". The fact that this expression depends only on position (and not on velocity), tells us we have a Riemannnian manifold. Norm of the metric tensor gives us the ``conversion'' factor. Note that norm of the metric tensor in general depends on the coordinate system and is in general point-dependent. There always exists a coordinate system in which the norm of the metric tensor is constant. The conversion factor for marginal distributions is linked to the eigenvalues along each coordinate. Flatness means there exists a coordinate system where we can factorize the distribution (if it factorizes in phase space), which is exactly the Cartesian coordinate system. (Note, we can always factorize in phase space) -> Therefore this focus on densities also allows us to get a better intuition even on the tangent space. [Still need to incorporate much of this discussion]

\subsection{Density and Riemmanian geometry}

It is helpful to take a quick look at how densities work in general so that we can better understand what is so special about them in the context in Hamiltonian mechanics. By density we mean the intuitive physical notion of a quantity divided by region it is found, like mass per volume or charge per area. Over a continuum, this is an integrable function $\rho(x)$. Mathematically, the concept is captured by the Radon–Nikodym derivative: we have a region $U$, we have two measures $m$ and $V$ and the derivative satistifies $m(U) = \int_U \rho dV$. This makes it clear that the units of $\rho$ depends on the units of both $m$ and $V$. More precisely, if $\rho(x, y)$ is a function of multiple variables, we will want $\int_U \rho(x, y) dx dy$ so that the units of $\rho$ are consistent with the units of the arguments. For example, in polar coordinates $\rho(r, \theta)$ will return mass over meter-radians.

This object is coordinate dependent because its definition requires us to make a choice of coordinates. Worse than that, it is not coordinate invariant: the density changes with the Jacobian. For example, changing polar to Cartesian we have $\rho(x,y) = |J| \rho(r, \theta) = r \rho(r, \theta)$. Since the Jacobian is a function of the point, this makes the value of the density, in general, not directly comparable at different points. If we have a uniform density in $x$ and $y$, we will have an $r$ dependence in $r$ and $\theta$. This dependence is introduced precisely because the area identified by $dr$ and $d\theta$ is $dr \,r d\theta$, which is a function of $r$.

We can look at this through the lens of differential geometry. First of all, note that in differential geometry there is also an object called density, but this is not the function $\rho$ we are looking for. A density in differential geometry is essentially a volume form (i.e. an alternating function from $n$ vectors to a real number) that is unaffected by changes in orientation. What we are looking for is really the component of a volume form $\nu = \rho(x^i) dx^1 \wedge dx^2 \wedge ... \wedge dx^n$. Note that $\nu$ combines the density and the volume, and that is why it is coordinate-free. Comparing the integrals we have $m(U) = \int \nu$ so essentially $\nu = dm$ which is the infinitesimal mass contribution we are integrating. But this means the value returned by $\nu$ does not depend only on how must stuff is there at the point, but on the size of the volume identified by the $n$ vectors given to the form.

In Riemannian geometry, we have a canonical volume form $dV = \epsilon \sqrt{|g_{ab}|} dx^1 \wedge dx^2 \wedge ... \wedge dx^n$ where $|g_{ab}|$ is the determinant of the metric tensor as expressed in local coordinates and $\epsilon$ is $1$ or $-1$ depending on the orientation. It can also be expressed as $\sqrt{|g_{ab}|}\epsilon_{a_1 ... a_n}$ in component notation where $\epsilon_{a_1 ... a_n}$ is the Levi-Civita symbol. We can then express any differential density $\nu = \hat{\rho} dV$ as the product of the volume density times a scalar function $\hat{\rho}$. By simply looking at the math, we may be mislead to think we were able to express the density in a coordinate independent way, but this is not the case. The quantity $|g_{ab}|$ depends on the choice of coordinates and is, in general, a function of the point. However, we can always find a set of coordinates that, at least locally, makes that quantity constant. The determinant $|g_{ab}|$ is the Jacobian of the transformation to those coordinates. For example, in Cartesian coordinates $|g_{ab}| = 1$ and the volume form $dV = dx \wedge dy$; in polar coordinates $|g_{ab}| = r^2$ and the volume form $dV = r dr \wedge d\theta$. This means that $\hat{\rho}$ is really expressing the density in a particularly convenient Cartesian-like coordinate system: one in which we can compare the density point-wise. We can also see that if we look at the units of $\hat{\rho}$, which is terms of inverse volume as expressed in the privileged system.\footnote{This is similar to what happens with the concept of proper time: it is defined in the comoving frame, so it is coordinate dependent. Yet, the expression $d\tau = \frac{1}{c} \sqrt{g_{ab} dx^a dx^b}$ is coordinate invariant.}

We can now look at what happens for distributions over position and velocity. Note that the velocities transform in the same way (i.e. controvariantly) than differentials in position. Therefore $\rho(x^i, v^j) = \sqrt{|g_{ab}|} \sqrt{|g_{ab}|} \rho(\hat{x}^i, \hat{v}^j) = |g_{ab}| \rho(\hat{x}^i, \hat{v}^j)$ $dx^a$. The determinant of the metric tensor is therefore directly the volume of the tangent bundle. The relationship to the density as expressed in phase space is therefore the following:
\begin{equation}
	\rho(q^i, p_j) = |g_{ab}| \rho(x^i, v^j).
\end{equation}
A change of coordinates will induce a controvariant change in the velocity and a covariant change in the momentum. The left side will remain invariant. The two terms on the right size will change in the opposite way so that their product is invariant. Focusing on densities, then, brings about a physically motivated way to understand the relationship between the cotangent and the tangent bundle, where the metric tensor plays a role in defining the geometry of the second in a similar way that the symplectic form defines the geometry of the first: it defines the count of states so that densities can be properly tracked.

---

Section on Ham. Not everything physical has to be coordinate independent or invariant.

* In phase space, we do have coordinate invariance and point-wise comparison.  Momentum changes covariantly. It is the density that is truly fundamental, not the coordinates in general. Though, the q's are priviledged from epistemic access. The p's are not, in fact we can define gauge transformation as things that keep the q's the same but change the p.

*[Could save the following for discussion of gauge redundancy in section 5.2] Not all ``state variables" are physical. Hamiltonian mechanics does not really care what the q's and p's are. It is the motion assumption that tells us that q's are position, simply because they define the unit system. P is a guage quantity (we care about the trajectory, the velocity). In this view, p is simply a ``label'' we can use to express the density over states. Therefore we do care about the physicality of the invariant distribution more than the physicality of the state variables themselves. -> Invariant density is really the important physical element/object in Hamiltonian mechanics.

* Units are physically significant. In fact, the ability to express the densities in units is what gives us the symplectic structure. It's a physical requirement that we must be able to express the density in some unit.

* Invariant densities: physical significance connected to coordinate independence. [incorporated]

* Density comparison is independent of coordinates. We can compare densities point-wise without worry. [incorporated]

* Makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous. Areas/volumes are measures of number of states. They are coordinate invariant. Also the Shannon/Gibbs entropy are coordinate independent (see physicality of quantities that are coordinate invariant)

* Close the section with a more general discussion on physicality/fundamentality and all those other physical goodness.

We have now substantiated the claim of Section~\ref{privileging} that Hamiltonian mechanics requires less structure than Lagrangian mechanics, in virtue of requiring fewer physical conditions. While this provides one reason for privileging Hamiltonian mechanics, it is not the only one. A further---possibly even more important---reason stems from the relationship between Hamiltonian mechanics and invariant densities. As we argue in this section, viewing physics through the lens of invariant densities has much to commend it. In general, it provides a better understanding of classical systems. Specifically, we will see that invariant densities are of fundamental physical significance due to their coordinate-independence [invariance?]. Hamiltonian mechanics---unlike Lagrangian mechanics---allows us to compare these densities point-wise in any coordinate system. Although the density is physically significant because it is coordinate-invariant, we will nevertheless argue that coordinate-invariance is not a necessary condition for physical significance. In fact, the ability to express densities in some choice of units is closely connected with symplectic structure. Hence, we will argue that some coordinate-dependent representational choices are nevertheless physically significant. This view is at odds with metaphysically-driven demands to represent physical theories in exclusively ``intrinisic" terms [get citations: Field 1980; Chen papers; North]. 
% Section~\ref{action} uses the invariant density to provide a geometrical interpretation of the action principle.

In the Hamiltonian formalism, the density has the following desirable property: we can compare the density point-wise across phase space. If the scalar value of the density at one point is larger than another, this entails that the density is higher at that point of phase space. Physically, this means that there is more physical stuff at this point. Likewise, if the scalar value of the density is the same at two different points, this entails that there is the same amount of stuff at those points. For brevity, will refer to this as the property of \textit{point-wise comparison}. Since the density depends in general on the choice of units, point-wise comparison does not hold universally across representations of state space. Instead, for the density of states to have this property, we require a coordinate-independent definition of the density. As the derivation in Section~\ref{deterministic} shows, we must define our state space using a symplectic manifold in order to define an invariant density. In other words, to define a density with the property of point-wise comparison, we must use canonical coordinates $(q, p)$ defined on a symplectic manifold. As we will discuss further in Section~\ref{gauge}, the conjugate momentum $p$ is itself a gauge-dependent quantity. Hence, the real physical significance of canonical coordinates is that they allow us to define an invariant phase space density $\rho $.

It is instructive to contrast this situation with Lagrangian mechanics. In this case, the density depends on the position and velocity state variables $(q,\dot{q}) $. Crucially, these state variables transform contravariantly, rather than covariantly. Consequently, the density $\rho(q,\dot{q})$ also varies contravariantly. These transformation properties preclude the density---in the Lagrangian formalism---from being coordinate-invariant in general. Of course, there exist special coordinate systems that allow for faithful point-wise comparison of the density. For instance, Cartesian coordinates---for a system in an inertial frame of reference---support this formal representational property. Yet, we still need to know which coordinate system we are in (or have transformed to) before making these point-wise comparisons. In general, non-Cartesian or non-inertial coordinate systems will not allow for point-wise comparisons of the density. The Hamiltonian formalism circumvents this further epistemic requirement by defining the density using quantities that vary covariantly. Hence, we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in.

[Should we briefly illustrate or gesture at an illustration of a failure of point-wise comparison? E.g., consider a situation where the density is a constant $k$ as a function of Cartesian coordinates (x, y). If we transform to polar coordinates $(r,\theta) $, then the scalar value of the density is no longer a constant (due to the Jacobian of this transformation not being unity). For instance, the scalar value of the density increases as we go further out, even though the density is physically the same. Hence, to determine whether there is more stuff at one location in state space vs. another, we need to know which coordinate system we are in, and how the density depends on these coordinates (i.e. we can't just compare the values of $\rho$ in these coordinates and see which is larger).]

% Connect this back with fundamentality: epistemic notion of the fundamental (don't need to know which coordinate system we are in), along with a physical sense: supporting a physical invariant, namely the invariant phase space density. Showing that this is more physically significant than the state variables themselves.

Section~\ref{fundamentality} analyzed fundamentality in terms of generality. We argued that Hamiltonian mechanics is more fundamental than Lagrangian mechanics in part because it is more general. By considering how Hamiltonian mechanics supports faithful point-wise density comparisons, we have seen two further notions of fundamentality. The first is epistemic: by representing the density on a symplectic manifold, we do not need to know which coordinate system we have chosen to work in. Any and all coordinate systems will support the formal property of point-wise comparisons [what about time-dependent coordinate transformations?]. Not only is this formal property methodologically advantageous, but it also bespeaks a further physical significance. This brings us to a second, physical notion of fundamentality: the phase space density $\rho$ is physically fundamental because it is a coordinate-invariant. Assuming that coordinate-invariance is sufficient for physical significance, the density thereby represents a genuine physical property of classical systems. It quantifies the amount of stuff in each phase space state. [to do: should we make this precise in terms of representational theory of measurement? E.g., the particular kind of scale that this representation defines and the meaningful comparisons it supports?].

% Here, we consider an epistemic sense of ``fundamental." In this sense, Hamiltonian mechanics is ``fundamental" because it shows that densities are coordinate-independent quantities. Assuming that coordinate-independence is sufficient for being physically significant, this shows that these densities are physically significant. More precisely, they are physical quantities or properties of the system.

[To do: expand on the following point: ] This is why statistical mechanics works best on phase space (we have a uniform measure over q/p). See the relevant * point above.

% Point of the following paragraph is to draw a contrast with North, who also seeks to identify physically significant quantities in terms of coordinate-invariants. We are pointing out that coordinate-invariance is not (on our view) a necessary condition for physical significance.

While it is important to provide a physical interpretation of invariants, this is not to say that only invariant quantities are physically meaningful.  For instance, the coordinates $q$ define the units, and units are physically significant despite not being invariant. Units are physically significant because they allow us to assign numbers to a physical quantity. They are physically significant in part because they correspond to measurable quantities. Indeed, as the derivation in Section~\ref{deterministic} shows, a necessary condition for defining the symplectic structure is that we can express the density in a particular choice of units. If we did not have the physical ability to define the density in some units---although any particular choice is arbitrary---we would not be able to define this structure. [To do: make this argument more clear]

Whereas North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}, some choices of the coordinates $q$ are less physically meaningful because they are not properly connected with measurement. Some arbitrary mixings of $q$'s and $p$'s preserve the symplectic manifold but are not as physically significant. For this reason, the $q$'s are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative $p$'s are not. Whereas symplectic geometry creates a mathematical, syntactic symmetry between the $q$ and $p$ variables, this is not a physically significant symmetry from the standpoint of measurement.
%cite van Fraassen on the measurement point; e.g. APA session on Halvorson book

Further support for the preceding point: could discuss how Cartesian coordinates can be physically significant, in virtue of position and velocity being independent degrees of freedom, whereas this independence does not hold generically across coordinate systems. (In contrast, position and conjugate momentum variables always characterize separable/independent degrees of freedom).

% Have mostly done the following: [To do: expand on the following point, and try to eliminate the notion of ``simplicity" and replace it with a more epistemic notion. Ability to make comparisons of density, independently of which coordinate system you are in or knowledge of being in a ``good" coordinate system. Josh: see 11-30-20 meeting notes where I think we discussed this.]: Compared with Lagrangian mechanics, canonical pairs (position and conjugate momentum) allow a simpler comparison of densities over space.

[Not sure if/where the following claim is relevant: ] This difference in the number of assumptions corresponds to a difference in the number of commitments. When working with a Hamiltonian but non-Lagrangian structure, we are not committed to the q's being positions (although is this in tension with the remarks on the significance of the coordinate and unit systems?). In contrast, for systems where motion equivalence applies, we gain the further commitment of treating the q's as positions.

% By definition, a dynamical system is one where we can write xdot as a function of x (i.e. the current state determines the evolution of the system). Having a Hamiltonian system further means that we can express this function xdot(x) using the Hamiltonian.

\section{Interpreting Lagrangian mechanics}
\label{Lagrangian}

By analyzing classical systems in terms of their invariant phase space density, we gain new insights into Lagrangian mechanics. Section~\ref{action} shows how the action principle can be understood geometrically in the extended phase space as a consequence of ``state density" conservation. We show that the action principle holds even when the action cannot be expressed as a function of just position and velocity, since it can always be expressed as a function of position and conjugate momentum. Section~\ref{gauge} discusses how the Lagrangian is highly gauge-redundant, and therefore less physical than the Hamiltonian. This provides a further argument for privileging Hamiltonian mechanics in the interpretation of classical systems. Finally, Section~\ref{Curiel} considers Curiel's \parencites*[]{Curiel} argument for privileging Lagrangian mechanics, which he alleges encodes the necessary and sufficient structure for capturing the physically significant kinematical constraints of classical systems. This raises a number of objections to our own position, which we rebut. In short, we argue that we have given an eminently natural construction of Lagrangian mechanics, viewed as a special case of Hamiltonian mechanics. Moreover, our argument shows that Lagrangian mechanics is in some sense less natural than Hamiltonian mechanics because the former is contingent on having a bijection between velocity and conjugate momentum.

\subsection{A geometric interpretation of the action principle}
\label{action}

% Somewhere below, emphasize that you have a unique trajectory only if the action is a minimum or maximum (if it is just stationary but not a minimum or maximum, then you do not have a unique trajectory)

Hamilton's principle of stationary action serves as a foundational principle for Lagrangian mechanics. According to this principle, infinitesimal variations of the action $ \mathscr{A}$ along the system's trajectory $\gamma $ equal zero: $\delta  \mathscr{A}[\gamma] = 0 $. Although central to Lagrangian mechanics, we can actually derive the stationary action principle using Hamiltonian mechanics. Our derivation simply assumes that 1) the system trajectory follows a deterministic and reversible evolution and 2) the system satisfies the motion equivalence condition. We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a similar argument illustrating the generalization, see \textcites[110]{Arnold1989}. [To do: verify that Arnold can legitimately be read as generalizing the argument below.]}
% Note on the generalization: cannot use curl. Need to use a form that measures the number of states on a surface. The displacement field kills this form (eats the degenerate direction that does not form degrees of freedom).
% Something that I thought was a second assumption but turns out that this follows just from double role played by time variable: 2) the system's motion always advances in time (so that $dt/dt = 1$).

Knowing that the system is deterministic and reversible, it follows that the density of states is invariant. This means that possible trajectories of the system are never pushed toward each other or pulled apart. In other words, the flux of trajectories through any closed surface is always zero. Formally, the displacement of the states in time is divergenceless: $\nabla \cdot \vector{S} = 0 $. Here, $\vec{S} := (\frac{d q }{d t }, \frac{d p }{d t }, \frac{d t }{d t })$ defines a vector field in the extended phase space $(q, p , t ) $. $\vector{S}$ characterizes the displacement of states in time. In particular, it is tangent to the trajectory $\gamma [q, p, t] (t)$, in extended phase space, where the variable $t$ functions as both a coordinate and affine parameter of the trajectory.

Since $\vector{S}$ is divergenceless, there exists another vector field $-\vector{\theta}$ such that $\vector{S} = curl(-\vector{\theta} )$.\footnote{For reasons that will become clear, we choose $-\vector{\theta}$ rather than $\vector{\theta}$.} As a vector field in the extended phase space, $\vector{\theta}$ can be written in terms of its components $\vector{\theta} = (\theta_q, \theta_p, \theta_t) $. Since the curl of the gradient of a scalar function is always zero, the displacement vector $\vector{S}$ is invariant under the gauge transformation $\vector{\theta'} =\vector{\theta} + grad(f)$, where $f$ is an arbitrary scalar function. Hence, we can choose $f$ such that $\partial f/ \partial p = -\theta_p $. Without loss of generality, we can thereby redefine $\vector{\theta}$ as $(\theta_q, 0, \theta_t) $.

Treating time $t$ as both a coordinate and an affine parameter (parameterizing the trajectory), the time component of the displacement, $S_t = dt/ dt = 1$.\footnote{Considering $dt/dt$, the time variable in the numerator functions as a coordinate, while the time variable in the denominator functions as a parameter. We could instead use a parameter $\tau$, and the derivation would remain valid.} This places another constraint on $\vector{\theta} $, since $\vector{S} = -curl(\vector{\theta}) $. In components, this means that $S_t = - (\frac{\partial}{\partial q} \cdot \theta_p - \frac{\partial}{\partial p} \cdot \theta_q) = - (\frac{\partial}{\partial q} \cdot 0 - \frac{\partial}{\partial p} \cdot \theta_q) = \frac{\partial q}{\partial p}$. Hence, $S_t = 1$ entails that $\frac{\partial q}{\partial p} = 1$. Integrating this expression, we find that $\theta_q = p$ up to addition by an arbitrary constant function $g(q)$. Again using the gauge freedom of $\vector{S}$ we can set $g(q)=0$, so that $\theta_q = p$. Thus, so far, we have shown that $\vector{\theta} = (p, 0,\theta_t) $.

Next, we introduce a function $H$ and let $ \theta_t \equiv -H$. Hence, we write $-\vector{\theta} = (-p, 0, H) $ as the vector potential for the displacement vector $\vector{S} $. By matching components of $\vector{S}$ and $-curl(\theta)$, we show that H is the Hamiltonian for the system. First, by definition, $S_q \equiv \frac{dq}{dt} $ and $S_p \equiv  \frac{dp}{dt} $. Secondly, since $\vector{S} = -curl(\vector{\theta}) $, we calculate that $S_q = -\frac{\partial \theta_t}{\partial p} = \frac{\partial H}{\partial p}$ and $S_p = \frac{\partial \theta_t}{\partial q}=-\frac{\partial H}{\partial q}$. Matching terms, we see that $\frac{\partial H}{\partial p} =\frac{dq}{dt} $ and $-\frac{\partial H}{\partial q} =\frac{dp}{dt}$. These simply are Hamilton's equations, which justifies interpreting $H$ as the Hamiltonian.

% Since the system is deterministic and reversible, it satisfies Hamilton's equations: $\frac{\partial H}{\partial p} =\frac{dq}{dt} \equiv S_q$ and $-\frac{\partial H}{\partial q} =\frac{dp}{dt} \equiv S_p $. Additionally, since $\vector{S} = -curl(\vector{\theta}) $, we know that $S_q = -\frac{\partial \theta_t}{\partial p}$ and $S_p = \frac{\partial \theta_t}{\partial q}$. Integrating the first equation for $\theta_t$ gives $-H$ up to an arbitrary constant function of q, while the second gives $\theta_t = -H$ up to an arbitrary constant function of p. For consistency, these two constant functions must vanish, so that $\theta_t = -H$ [check argument here; do we actually need to appeal to the gauge freedom of S again?]. Hence, it follows that we can take $-\vector{\theta} = (-p, 0, H) $ as the vector potential for the displacement vector $\vector{S} $.

At this point, we return to the action $ \mathscr{A}$, defined as the integral of the Lagrangian over a trajectory beginning at time $t_0$ and ending at time $t_1$: $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt$. Provided that the Legendre transform is well-defined (i.e. the motion equivalence assumption is satisfied), then $H = p \dot{q} - L$. Hence, we can re-express the action as $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt = \integral^{t_1}_{t_0} [p \frac{dq}{dt} - H(q, p, t) ]dt$. Multiplying the Hamiltonian by $dt/dt = 1$ and adding a vanishing $0 \cdot \frac{dp}{dt}$, we obtain $ \mathscr{A} = \integral^{t_1}_{t_0} [p \frac{dq}{dt} + 0 \cdot \frac{dp}{dt} - H(q, p, t) \frac{dt}{dt} ]dt$. We notice that the integrand simply equals the inner product of $\vector{\theta}$ with $\vector{S} $, entailing that the Lagrangian $L =\vector{\theta} \cdot \vector{S} $. Consequently, the action $ \mathscr{A} [\gamma] = \integral^{t_1}_{t_0} \vector{\theta} \cdot \vector{S}  dt$. Changing variables, we take the line integral of $\vector{\theta}$ along the trajectory $\gamma $: $ \mathscr{A} [\gamma] =\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$.\footnote{Note that $\vector{S}  dt = d\vector{\gamma}$ only on the actual trajectory in phase space, but not on other possible trajectories.} To derive Hamilton's principle of stationary action, it suffices to show that $\delta \mathscr{A} [\gamma] =\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = 0$. 
% Note on the generality of the argument: in cases where motion equivalence is not satisfied, we can still write the integrand in terms of H(q, p, t), although it would not be an integral over the Lagrangian. Thus, we can still express the `action' as a function of p, q, t. We would calculate this on a path in the extended phase space, although without motion equivalence, not all of the paths on phase space would correspond to a unique trajectory in real, physical space. See if Arnold has a more general name for this generalized principle of stationary action.

By definition, a variation over the integral $\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ consists in taking the difference between the integrand and a corresponding integral over a nearby trajectory $\gamma'$: $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} - \integral_{\gamma'} \vector{\theta} \cdot d\vector{\gamma'}$ . Since the trajectories $\gamma$ and $\gamma'$ have the same endpoints, they define a closed-curve $C$, enclosing a region $\Sigma $. Hence, by Stoke's theorem $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \oint_{C} \vector{\theta}  \cdot dC = \iint_{\Sigma} curl(\vector{\theta}) \cdot \diff \vector{\Sigma}$ . Recalling now that 1) the $curl(\vector{\theta})$ equals the negative of the displacement vector $\vector{S}$, and 2)  $\vector{S}$ is tangent to the trajectory $\gamma $, we see that $\vector{\theta}$ is tangent to the surface $\Sigma $.\footnote{Since $\vector{S}$ is tangent to $\gamma $, and $\gamma'$ defines an infinitesimal variation, it follows that $\vector{S}$ is tangent to the entire surface $\Sigma $.} Hence, the surface integral must vanish, showing that $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ does indeed equal zero. This establishes Hamilton's principle of stationary action, which we have derived by assuming that 1) the displacement vector is divergenceless and 2) using the Legendre transform (assuming motion equivalence).

(XXX: Note how the geometry is better understood in extended phase space, requiring Hamiltonian picture. Note that this applies even when $\theta S$ cannot be expressed as function of $q$ and $\dot q$. )

%Discussion of gauge transformations in the context of Lagrangian vs. Hamiltonian mechanics; contrast with North [77] privileging the momentum variables as significant (look for Wallace 2003 quotations along these lines as well):

\subsection{Gauge redundancy in Lagrangian mechanics}
\label{gauge}

\begin{comment}

% 12-19: I think that I incorporated the relevant points below:

** Overall point: this approach allows us to get a better intuition for what gauge transformations represent.

* From motion equivalence -> we really only care about the trajectory. That is what we want to represent and what we measure.

* A gauge transformation is a change in p variable that leave q and q dot unchanged. Therefore it is simply a change of how states are labeled, without a change in the actual observable dynamics. Add the derivation.

* In Hamiltonian mechanics, the momentum is gauge dependent. In Lagrangian mechanics the kinematic variables are not gauge dependent, but the Lagrangian manifestly is. $L = \theta S$. $S$ is gauge independent but $\theta$ is not. Therefore we will have problems understanding the physicality of the Lagrangian. Similar concerns for the Action. -> neither formulation has the upper hand

* Coordinate transformation: if we restrict to coordinate transformation that only depend on space, both Lagrangian and Hamiltonian are invariant. If we introduce coordinate transformation that do depend on time (e.g. rotating frame), then Lagrangian is still invariant while Hamiltonian manifestly changes as a time component of a co-vector. (because H is the time component of $\theta$). That is why the Lagrangian framework is prevalently used in intro classes.

* This is a problem of the double role of time as ``coordinate'' and ``affine parameter''. To be proper in differential geometry, we have to separate the two uses. So we add a ``temporal'' degree of freedom $(t, E)$ and the evolution is over the affine parameter $s$. In that framework, the Hamiltonian constraint serves both as a global constraint that fixes the energy as a function of $q, p, t$ and as the generator of the affine parameter $s$. The Hamiltonian constraint is invariant. If the affine parameter is proper time, then the Hamiltonian constraint is essentially the mass. -> There is a better formulation of Hamiltonian mechanics, that is typically not found in textbooks, that addresses the problem.

\end{comment}

% Last page of Halvorson/Swanson article seems to broach this issue of gauge redundancy, although they seem to think it is possibly a positive feature of Lagrangian mechanics (a source of additional representational flexibility) rather than a defect.

Understanding both mechanics in terms of our assumptions gives us a way to understand gauge transformations that is both more physically grounded and intuitive. Recall that motion equivalence entails that knowledge of the trajectory in physical space suffices for knowledge of the state in phase space. In other words, we can completely characterize states in terms of the canonical position coordinate $q$ and its first derivative $\dot{q}$. Hence, any transformation that leaves $(q,\dot{q}) $ invariant \textit{ipso facto} leaves the physical state invariant. We want such transformation to be canonical as to preserve the invariant densities. This means that the new state variable $(q, \hat{p})$ will respect the usual Poisson brackets:
\begin{equation}
\begin{aligned}
	\{q^i, q^j\} &= 0 \\
	\{q^i, \hat{p}_j\} &= \delta^i_j \\
	\{\hat{p}_i, \hat{p}_j\} &= 0
\end{aligned}
\end{equation}
Using the second equation we find:
\begin{equation}
\begin{aligned}
	\{q^i, \hat{p}_j\} &= \frac{\partial q^i}{\partial q^k} \frac{\partial \hat{p}_j}{\partial p_k} - \frac{\partial q^i}{\partial p_k} \frac{\partial \hat{p}_j}{\partial q^k} \\
	&= \delta^i_k \frac{\partial \hat{p}_j}{\partial p_k} - 0 \frac{\partial \hat{p}_j}{\partial q^k} = \frac{\partial \hat{p}_j}{\partial p_i} = \delta^i_j
\end{aligned}
\end{equation}
Integrating we have:
\begin{equation}
\begin{aligned}
\hat{p}_j = p_j + A_j(q^i)
\end{aligned}
\end{equation}
where $A_j$ are arbitrary functions of position. Using the third Poisson bracket, we have:
\begin{equation}
\begin{aligned}
0 = \{\hat{p}_i, \hat{p}_j\} &= \{p_i + A_i, p_j + A_j\} \\
&= \{p_i, p_j \} + \{p_i , A_j\} + \{A_i, p_j \} + \{A_i, A_j\} \\
&= 0 - \{A_j, p_i\} + \{A_i, p_j \} + 0 \\
&= - \frac{\partial A_j}{\partial q^i} + \frac{\partial A_i}{\partial q^j} \\
&= curl(A_j) \\
\end{aligned}
\end{equation}
Since $A_j$ is curl free, we can express it as the gradient of a function $f$. Therefore changes of state variable of the form $\hat{p}_j = p_j + \partial_j f$ do not change either the equations of motion or the trajectories. We recognize these as the gauge transformations.

Understanding the action principle in terms of Hamiltonian mechanics also provides a better understanding of gauge transformations. If motion equivalence holds, then gauge transformations leave the variables $(q,\dot{q}) $ invariant. To see this, recall that motion equivalence entails that knowledge of the trajectory in physical space suffices for knowledge of the state in phase space. In other words, we can completely characterize states in terms of the canonical position coordinate $q$ and its first derivative $\dot{q}$. Hence, any transformation that leaves $(q,\dot{q}) $ invariant \textit{ipso facto} leaves the physical state invariant, i.e. is a gauge transformation. [To do: make it more clear how the discussion in this section is intrinsically connected with the earlier geometric treatment of the action and our overall privileging of Hamiltonian mechanics]

In particular, there is a family of gauge transformations that changes the conjugate momentum variable $p$ while leaving $(q,\dot{q}) $ invariant. [To do: illustrate this with general formula for this kind of gauge transformation, and possibly the electrodynamics example.] Consequently, the conjugate momentum in Hamiltonian mechanics is gauge-dependent. The $p $ variables are therefore not of fundamental physical significance [to do: check that this is consistent with Section 2.1's discussion of fundamentality]. Hence, privileging Hamiltonian mechanics does not entail, \textit{pace} North, that ``momentum is a truly fundamental quantity, on a par with position" \parencites*[77]{North}. [To do: see if Wallace 2003 also makes this claim about the fundamentality of conjugate momentum, or the similarly flawed claim that conjugate momentum and position are on equal footing within Hamiltonian mechanics]. Although the canonical position and momentum variables are on equal syntactical footing within symplectic mechanics, this does not entail that they are equally physically significant.

Instead, position and kinetic momentum are physically privileged, since these quantities are gauge-invariant. Consequently, Lagrangian mechanics employs state space variables that are gauge-invariant, while Hamiltonian mechanics uses a gauge-dependent variable. Despite this advantage, Lagrangian mechanics has a compensating disadvantage: the Lagrangian is itself gauge-dependent. The formalism of Section~\ref{action} makes this gauge dependence manifest. Representing the Lagrangian as $L =\vector{\theta} \cdot \vector{S} $, we see that since the vector potential $\vector{\theta}$ is gauge-dependent while the displacement vector $\vector{S}$ is not, $L$ must also be gauge-dependent.\footnote{As shown in Section~\ref{action} we can arbitrarily transform $\vector{\theta}$ by adding the gradient of an arbitrary scalar function, while $\vector{S}=curl(-\vector{\theta}) $ remains invariant under this gauge transformation.} The gauge-dependence of the Lagrangian makes it difficult to provide a clear physical interpretation of any particular Lagrangian.\footnote{The action inherits the gauge-dependence of the Lagrangian. This does not affect the principle of stationary action because the local minima and maxima of the action remain invariant under gauge transformations.} In contrast, the Hamiltonian is defined up to a constant, endowing it with greater physical significance. Overall then, we see that neither Lagrangian nor Hamiltonian mechanics has the upper hand when it comes to invoking only gauge-independent physical quantities. Both formalisms have their trade-offs in this regard (although, it is worth emphasizing that---as argued in Section~\ref{density}---Hamiltonian mechanics provides the natural setting for the invariant phase space density).

Moving from gauge transformations to coordinate transformations, we gain further reasons for using Hamiltonian mechanics within the extended phase space formalism. For many applications, it suffices to use time-independent coordinate transformations. In this special case, both the Lagrangian $L$ and the Hamiltonian $H$ are coordinate-invariant. However, some applications benefit from (or even require) the use of time-dependent coordinate transformations, e.g. a rotating frame of reference. In this case, the Lagrangian remains coordinate-invariant, while the Hamiltonian becomes coordinate-dependent. Again, the formalism of Section~\ref{action} makes this time-dependence manifest. As we saw there, the Hamiltonian $H$ is the time component of the vector potential $\vector{\theta} $, entailing that $H$ transforms as the time component of a co-vector. Methodologically, this makes the Hamiltonian framework less convenient for treating this class of problems.

Nonetheless, formulating Hamiltonian mechanics within the extended phase space provides one way to rectify these deficiencies. At root, the time-dependence of the Hamiltonian stems from the dual role of time as both a coordinate and an affine parameter within the usual Hamiltonian framework. Yet, differential geometry demands that we separate these two roles. We thereby add a temporal degree of freedom $(t, E) $, expanding the phase space degrees of freedom by one (recall that we previously had one degree of freedom for every canonical coordinate pair $q_i, p_i$). Next, we define the phase space evolution over an affine parameter $s $. In this extended Hamiltonian framework, the Hamiltonian constraint functions as both a global constraint---fixing the energy as a function of $q, p, t $---and as a generator of the affine parameter $s $ [to do: spell out what the `Hamiltonian constraint' is in general]. Crucially, the Hamiltonian constraint is now coordinate-invariant, even with respect to time-dependent coordinate transformations. For example, if we define the affine parameter as the proper time, then the Hamiltonian constraint is proportional to the mass (which is manifestly coordinate-invariant). This formalism thereby provides a clear physical meaning for the Hamiltonian constraint. [To do: provide some reference for this modified Hamiltonian formulation]

[To do: not sure how best to end this subsection/transition to the next one]. We have now considered many conceptual and physical advantages of the Hamiltonian formalism. In Section~\ref{Curiel}, we turn to considering Curiel's \parencites*[]{Curiel} arguments against privileging Hamiltonian mechanics.

[Could address the following in this section, rather than in Section 4:] Not all ``state variables" are physical. Hamiltonian mechanics does not really care what the q's and p's are. It is the motion assumption that tells us that q's are position, simply because they define the unit system. P is a gauge quantity (we care about the trajectory, the velocity). In this view, p is simply a ``label'' we can use to express the density over states. Therefore we do care about the physicality of the invariant distribution more than the physicality of the state variables themselves. This entails that the invariant density is really the important physical element/object in Hamiltonian mechanics.

\begin{comment}

I think that I have incorporated most of the following into the above:

I didn't incorporate the following point: In some sense (to be made more precise), gauge transformations are more prevalent in Lagrangian mechanics than Hamiltonian mechanics. This is so despite the fact that the conjugate momentum $p $ is itself a gauge quantity.

% [To do: clarify the following claim:] motion equivalence allows us to understand the gauge redundancy of the Lagrangian. However, it is not true that the Lagrangian has these redundancies in virtue of the system satisfying motion equivalence.

[To do: clarify the following :] From motion equivalence, the Lagrangian has more invariant structure than the Hamiltonian [not sure what this really means]. The Lagrangian has the advantage of being invariant under coordinate transformations, whereas the Hamiltonian is not a coordinate invariant. Instead, the Hamiltonian varies as the time component of a 4-co-vector. Nevertheless, the Lagrangian itself is physically problematic because it is a gauge-dependent quantity [elaborate/illustrate this].

[Clarify the following, if relevant :] But there is a sense in which on the extended phase space, the Hamiltonian constraint is an invariant, yielding a coordinate-invariant way of handling systems with Hamiltonian mechanics.

[To do: connect the following point with the former :] Symplectic geometry indeed does not care about the difference between q's and p's. Classical mechanics does. Misindentification of the math (symplectic geometry) with the physics (classical Hamiltonian mechanics).
%Since the q and $\dot{q} $ variables fix the physical state, this indicates that transformed conjugate momentum is a gauge quantity. 
%North argues that ``momentum is a truly fundamental quantity, on a par with position" [77]. I think that Wallace 2003 article might make a similar point. Look for this.

\end{comment}

\begin{comment}
Issue of the mathematical structure that is necessary/sufficient for Lagrangian mechanics:
North footnote 20 and appendix: it is not generically true that the state space of Lagrangian mechanics comes equipped with a metric. Footnote 22: ``a more general, coordinate-free version" of Lagrangian mechanics ``does not make explicit use of this metric"; discussed in appendix A. North: ``the square of the Riemannian line element is what gives the allowable sets of coordinate transformations, the ones that preserve the Lagrangian equations of motion. A volume element would not suffice for this" [75]. How does this connect with motion equivalence and its associated mathematical structure?

North assumes that formulating a Lagrangian theory requires specifying a Riemannian metric; according to Curiel, this is neither necessary nor sufficient. Furthermore, the physical significance of a Riemannian metric in the context of a Lagrangian theory is unclear: two different Lagrangians can have the same Lagrangian vector field and yet induce inequivalent Riemannian metrics

%Somewhere in this section, should reply to Curiel's privileging of Lagrangian mechanics. Curiel frequently appeals to naturalness and a distinction between natural vs. ad hoc/artificial constructions [he defines `naturalness', specifically a `natural coordinate system', in terms of a particular kind of coordinate transformation. See his equation 4.3, page 22 of the PDF].

% Helpful discussion of differences between Riemannian and symplectic manifolds from Wikipedia (https://en.wikipedia.org/wiki/Symplectomorphism): ``Unlike Riemannian manifolds, symplectic manifolds are not very rigid: Darboux's theorem shows that all symplectic manifolds of the same dimension are locally isomorphic. In contrast, isometries in Riemannian geometry must preserve the Riemann curvature tensor, which is thus a local invariant of the Riemannian manifold. Moreover, every function H on a symplectic manifold defines a Hamiltonian vector field XH, which exponentiates to a one-parameter group of Hamiltonian diffeomorphisms. It follows that the group of symplectomorphisms is always very large, and in particular, infinite-dimensional. On the other hand, the group of isometries of a Riemannian manifold is always a (finite-dimensional) Lie group. Moreover, Riemannian manifolds with large symmetry groups are very special, and a generic Riemannian manifold has no nontrivial symmetries."
%Seems to me that this theoretical result supports North's analysis, and could offer a response to Halvorson/Swanson counterexamples that appeal to Kepler problem and Galilean boosts.

\end{comment}

\subsection{Response to Curiel's criticisms of Hamiltonian mechanics}
\label{Curiel}

Curiel \parencites*[]{Curiel} argues that Hamiltonian mechanics is less physically significant than Lagrangian mechanics. He attempts to show that the Lagrangian framework represents classical mechanical systems in a \textit{natural} matter whereas the Hamiltonian formulation involves an artificial construction \parencites*[270]{Curiel}. His argument hinges on the claim that Hamiltonian mechanics requires an ad hoc condition that is imposed by hand, whereas Lagrangian mechanics follows from physically plausible principles. [Set this up properly: Curiel's definition of abstract classical system; Curiel's privileging of certain kinematical constraints, his equations 4.3; Hamiltonian form of equation 6.2; see below]. Curiel restricts Hamiltonian mechanics to systems whose Hamiltonian takes the form (equation 6.2) $H =\alpha^{m n} p_m p_n + U (q_j) $. Here, the functions $U$ and $\alpha^{m n}$ are arbitrary functions of the generalized position coordinates $q $. The following provides a helpful summary of Curiel's argument against privileging Hamiltonian mechanics:

\begin{quotation}
Hamiltonian mechanics represents abstract classical systems only in so far as we restrict ourselves to a subfamily of all the formally acceptable Hamiltonians by the \textit{ad hoc} use of conditions foreign to Hamiltonian mechanics itself. Its structures do not provide the appropriate concepts and tools to formulate in their terms the required structures that treat configurative quantities differently from momental, nor do they provide any natural justification for the restriction to Hamiltonians of the form (6.2). \parencites[306]{Curiel}
\end{quotation}

We agree with Curiel that his equation 6.2 requires a condition that is imposed by hand, namely the motion equivalence assumption [explain this]. However, this condition has a clear physical meaning, and it corresponds to a physical condition that must hold in order to define/apply Lagrangian mechanics. Pace Curiel, there exist Hamiltonian systems that do not satisfy his equation 6.2 (equivalently, classical systems where the momentum is not a linear function of the velocity), and these will even have a Lagrangian, although they will not be an abstract classical system in Curiel's sense. Hence, there seems to be a burden on Curiel to distinguish---in a principled way---the classical Lagrangian systems from those that are not classical.
% Hence, perhaps we agree with Curiel that ``it is not necessary that [these conditions] hold in order for the theory to be meaningfully applied to model a [classical] system" [305]. Where is Curiel takes this as an argument for the primacy of Lagrangian mechanics, we take it as an argument for the primacy of Hamiltonian mechanics, since Curiel's conditions are equivalent to imposing motion equivalence.

% Curiel's main point: the representation of a classical mechanical system in the Lagrangian framework is \textit{natural} in a way that its representation in a Hamiltonian formulation is not. The former representation depends on intrinsic structures of the classical mechanical system rather than an artificial construction (270).

Rebuttal 2: On its own, the motion of a system in physical space is not sufficient for defining the system's dynamics. The kinematics/motion does not suffice to determine the forces acting on the system. For it is essential to be able to distinguish the forces acting on an inertial system from fictitious forces arising in the description of a non-inertial system. Hence, unless one knows that the motion equivalence principle is satisfied, Lagrangian mechanics faces an underdetermination problem: it is unable to distinguish between certain inertial vs. non-inertial systems that are physically distinct. Distinguishing between these kinds of systems requires a relationship between conjugate momentum and kinematic momentum (or equivalently, knowledge of the energy/Hamiltonian).
% Example of underdetermination in the context of Lagrangian mechanics: a free particle in a non-inertial frame vs. an accelerated particle (in an inertial frame) acted on by a dissipative force. These two distinct classical systems can nevertheless have the same equations of motion.
% The Hamiltonian framework does not face a corresponding underdetermination problem because it does not allow for dissipative systems. Hence, there is a sense in which the Lagrangian formalism is more mathematically general than the Hamiltonian formalism. However, this additional generality is irrelevant when considering conservative classical systems.

Rebuttal 3: The standard predator--prey model from differential equations satisfies Curiel's definition of an abstract classical system. However, it does not satisfy Curiel's notion of the split between configurational vs. velocital coordinates. And it is not true that the velocity will be the first derivative of the position. Moreover, the predator--prey system is neither Hamiltonian nor Lagrangian, because it is not necessarily deterministic and reversible.
% Related to equation 3.1. Make precise.
% Furthermore, the momentum does not need to be a linear function of velocity. For example, the photon does not satisfy this relationship.

Rebuttal 4: make precise and substantiate with textual evidence (might be relevant for responding to objection considered below): implicitly, Curiel assumes that classical systems are defined as Newtonian systems in an inertial frame with Cartesian coordinates. Implicitly, Curiel assumes that energy is a function of position and velocity, entailing that we have a Hamiltonian. The claim in footnote 17 that the kinetic energy is a function of velocity depends on the choice of a particular coordinate system; it is not a coordinate-invariant claim. Curiel's thought experiment assumes that we have a constant force (in this case gravity) and that we are in a reference frame where we can compare forces (e.g. an inertial frame).

[Most sophisticated objection I can think of on behalf of Curiel:] Like our argument, part of Curiel's argument involves appealing to systems outside of the hyper-regular domain. Curiel acknowledges both that only hyper-regular Lagrangians have unique solutions and that Hamiltonian systems have unique solutions even outside of the hyper-regular domain \parencites*[291, 308]{Curiel}. Nevertheless, he argues that this is actually an advantage of Lagrangian mechanics, rather than a disadvantage. On his view, the unique solutions of Hamiltonian mechanics outside the hyper-regular domain is precisely why ``one must impose the quadratic form of the Hamiltonian as an \textit{ad hoc} condition" \parencites*[308]{Curiel}. In contrast, he argues that Lagrangian mechanics precisely delimits the proper domain of abstract classical systems, namely those that satisfy a particular set of physically significant kinematical constraints (his equations 4.3). Although Lagrangian systems do not admit unique solutions outside the hyper-regular domain, these non-unique solutions will still satisfy the kinematical constraints of Lagrangian mechanics \parencites*[311]{Curiel}.

Summary: Hamiltonian mechanics provides solutions outside hyper-regular domain (but not these kinematical constraints) whereas Lagrangian mechanics preserves these kinematical constraints outside hyper-regular domain. Worry: perhaps GC and Curiel are in a verbal dispute about the proper/privileged definition of a classical system (with Curiel privileging these kinematical constraints, and GC privileging deterministic/reversible evolution).

On pages 309-310, Curiel considers a Hamiltonian that could model the photon (so it seems). In considering a Hamiltonian system that does not respect the Lagrangian kinematical constraints, Curiel considers the Hamiltonian $H=p$, which corresponds to a system moving with constant velocity (e.g. velocity of 1). So this could very well be a model of a photon, moving at the speed of light. Curiel points out that the corresponding Lagrangian: $L= vp - H = \dot{q}-1$, yields a trivial Lagrangian problem: $0=0$ upon evaluating the Euler--Lagrange equation. Upshot: ``not all Hamiltonian vector fields map to second-order vector fields under the inverse-Legendre transform" [310]. But then does Curiel consider this system to not be a classical system?

[To do: figure out if the following claims made by Curiel regarding the differences between the symplectic structure in Hamiltonian mechanics (which is fixed) vs. the symplectic structure that arises from the Euler--Lagrange equations (which varies from Lagrangian to Lagrangian) are relevant to our argument or rebuttals: \\ 
Hamiltonian mechanics: ``in Hamiltonian mechanics, the fact that the symplectic structure is fixed once and for all enforces the satisfaction of the kinematical constraints (5.3), and it encodes the existence and uniqueness of solutions to the equations of motion and the conservation of energy, among many other things. All of these features are independent of the individual Hamiltonian used in the equation" [310]. \\
Lagrangian mechanics: ``In Lagrangian mechanics, in contrast, it is the Lagrangian itself that encodes existence and uniqueness of solutions (which will obtain when and only when the Lagrangian is regular), and there is not even a guarantee that energy is conserved, as one can, for example, represent dissipative systems in the Lagrangian framework" [310].

% Either way, presumably we can agree with Curiel's explanation/understanding for why the Lagrangian must be regular: Curiel arguing that he can give an account of ``why it makes physical sense to demand that the Lagrangian be regular", i.e. satisfy equation 4.6 for the Hessian determinant. He notes that by considering the variational problem, we understand ``why the Lagrangian must be quadratic (at least) in the velocities for there to be a unique solution to the Euler--Lagrange equation" [308]. Argument: for the variational problem to be well-posed, you need to be able to ``take non-trivial derivatives of the Lagrangian up to second-order in the velocity, since the crucial integration by parts that yields the Euler--Lagrange equation demands this, on pain of giving, for example, the tautology $0=0$" in some cases, or a contradiction in others.

\section{Material looking for a possible home}

[To do: not sure if following warrants discussion. Could possibly connect this with the earlier discussion of the significance of defining the unit system, since that seems key to our response to this hypothetical worry.]
Teh and Tsementzis \parencites*[]{Teh} provide another argument for the equivalence of Lagrangian and Hamiltonian mechanics within the hyper-regular domain. They develop this using Tulczyjew's reformulation of Hamiltonian mechanics, which shows that ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry" \parencites*[46]{Teh}. Again, there is a worry that this equivalence result might undercut our privileging of Hamiltonian mechanics. However, the equivalence is again just within the hyper-regular domain, requiring that motion equivalence is satisfied. Furthermore, Tulczyjew's formulation relies heavily on the notion of a Lagrangian submanifold, since ``a mechanical system turns out to be a Lagrangian submanifold of an appropriately defined space" \parencites[47]{Teh}.\footnote{A submanifold $N$ of a symplectic manifold $(P, \omega) $ is Lagrangian provided that i) it has half the dimension of P and ii) the pullback of the symplectic form $\omega$ vanishes along the inclusion of N in P \parencites[48]{Teh}.} Yet importantly, Lagrangian submanifolds do not distinguish between the space of velocities and the space of momenta, i.e. they obscure the important physical differences between these two kinds of quantities. Not every constructible Lagrangian submanifold defines units in a physically significant way.\footnote{One could even construct a suitable Lagrangian submanifold by mixing position and momenta coordinates.} It is the physical significance of defining a unit system that breaks the formal (syntactic) symmetry between position and momentum coordinates. 
% [To do: verify the veracity of this discussion of Lagrangian sub manifolds]
%Furthermore, within this reformulation, Lagrangian submanifolds play a very important role. Does this lead to a possible privileging of Lagrangian mechanics?

[To do: Not sure where to put the following material, if anywhere; perhaps after a discussion of why we restrict to hyper-regular domain?:]
In a footnote, Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence. There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" \parencites*[1179]{Barrett2}. Indeed, outside the hyper-regular domain, the Euler--Lagrange equations do not have a single, unique solution. Hence, it is difficult to say what a Lagrangian system even is outside the hyper-regular domain. As an example, consider the case where the Lagrangian $L$ equals zero. Then for this ``system," all paths have the same action, so all paths are equally possible.\footnote{This Lagrangian corresponds to a Hamiltonian that also equals zero, entailing that $dH = 0 $. By Hamilton's equations, the derivatives of the position and momenta are zero, indicating that this system is static.} In contrast,  a Hamiltonian defines a unique solution (unique evolution in phase space) even when outside the hyper-regular domain (i.e. when the Hessian is zero).

\begin{comment}

Notes/questions from 11-16-20 meeting
\begin{enumerate}[1.)]

\item Possible connection with motion equivalence principle: Sameness of empirical content (empirical equivalence) is cashed out in terms of agreeing ``on the allowable trajectories for particles in the system" [1180]. For a Lagrangian system, this is characterized in terms of a base integral curve of the Lagrangian vector field $(X_L)^a$ that ``encodes one way that the positions of particles in a system with Lagrangian L might evolve over time", while the base integral curves of the Hamiltonian vector field perform an analogous role [1180]

\begin{itemize}

\item   Marsden and Abraham 1978, Theorem 3.6.2 proves that (in the hyper-regular domain) the Lagrangian vector field and the corresponding Hamiltonian vector field (obtained via Legendre transformation) have the same base integral curves. This means that they ``agree on how the positions of particles will evolve over time" [1181]. Is this the same as the motion equivalence assumption?

\end{itemize}

\item Barrett footnote 18 wonders whether there could be a ``correspondingly general formulation of Lagrangian mechanics" that is analogous to the symplectic but-not-cotangent-bundle formulation of Hamiltonian mechanics. Does GC derivation show that there cannot be such a correspondingly general formulation of Lagrangian mechanics?

\begin{itemize}

\item: GC in response: defining the Lagrangian presupposes that we have trajectories of positions over time. Velocity is then a derived quantity of these positions. So the tangent bundle seems to be intimately connected with having trajectories. Hence, if we eliminated reference to the tangent bundle, we would seemingly not be able to define trajectories.

\item Furthermore, there is a sense in which GC derivation recovers the tangent bundle and shows why it must be well-defined: provided the evolutions are differentiable (which is needed to preserve densities), it follows that velocities are well-defined, and hence so is the tangent bundle. Note that the trajectories alone do not entail that velocities are well-defined, since the trajectories could in principle not be differentiable.

\item Note that although the symplectic manifold already has some global structure (which corresponds to the global tangent bundle), the cotangent bundle itself is in some sense more global (and hence more specialized). Helpful analogy: a manifold can be locally Euclidean (which is a global future) without being globally Euclidean. The sense in which the tangent bundle and symplectic manifold structure are global is analogous to being locally Euclidean, whereas the global cotangent bundle structure is analogous to being globally Euclidean. So in going from the cotangent formulation to the symplectic formulation, we are becoming less specific (more general). To generalize the Lagrangian, we would seemingly need to lose some structure, but the action seems to be global (integral defined over a trajectory).

\item It might be worth noting these differences, as a way of motivating the claim that there probably is not a correspondingly general formulation of Lagrangian mechanics.

\end{itemize}

\item Barrett also shows that the vector field formulations of Lagrangian and Hamiltonian mechanics are not categorically equivalent [1185]. He interprets this as substantiating Curiel's claim that the ``Legendre transform does not respect the kinematical constraints of Lagrangian mechanics", namely because it shows that ``there are Lagrangians that give rise to the same Lagrangian vector field, but do not translate---via the Legendre transformation---into Hamiltonians that give rise to the same Hamiltonian vector field" [1186]. What does this mean in the context of motion equivalence?

\begin{itemize}

\item Barrett interprets this situation as simply showing that the Hamiltonian vector field contains more information about the Hamiltonian than the Lagrangian vector field does about the Lagrangian. If two Hamiltonian systems agree on the Hamiltonian vector field, then they must agree on the Hamiltonian, up to a arbitrary constant. Yet two Lagrangian systems can have the same Lagrangian vector field but not agree on the Lagrangian or total energy (even up to a constant).

\end{itemize}

\item Teh and Tsementzis discuss Tulczyjew's reformulation of Hamiltonian mechanics, which shows that ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry" [PDF 3]. Does this accomplish a similar moral as the assumptions of physics framework?

\item Furthermore, within this reformulation, Lagrangian submanifolds play a very important role. Does this lead to a possible privileging of Lagrangian mechanics?

\end{enumerate}

\end{comment}

\printbibliography

\end{document}