### 2021-05-30 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: sections 2 and 3 from Andrea

\documentclass[11pt, executivepaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{natbib}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{amsfonts}
\usepackage{MnSymbol}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{geometry}
 \geometry{
 a4paper,
 total={155mm,237mm},
 left=28mm,
 top=32mm,
 }
\usepackage{hyperref}
\hypersetup{colorlinks= true, allcolors=blue}
\setcitestyle{aysep={}}

\newtheorem{prop}{Proposition}

\begin{document}

\title{\textbf{On the Common Logical Structure of Classical and Quantum Mechanics}}

\author{Andrea Oldofredi\thanks{Section de Philosophie, Universit\'e de Lausanne, Switzerland. E-mail: Andrea.Oldofredi@unil.ch} \and Gabriele Carcassi\thanks{Physics Department, University of Michigan, Ann Arbor, MI 48109. E-mail: carcassi@umich.edu} \and Christine A. Aidala\thanks{Physics Department, University of Michigan, Ann Arbor, MI 48109. E-mail: caidala@umich.edu}}

\maketitle

\begin{abstract}
\center Notes on classical and quantum logic
\vspace{4mm}

\noindent \emph{Keywords}: Quantum Mechanics; Quantum Logic; Classical Logic; 
\end{abstract}
\vspace{5mm}
\clearpage

\tableofcontents
\vspace{5mm}

\section{Introduction}

As soon as Quantum Mechanics (QM) achieved a definite and coherent mathematical formulation, several physicists and philosophers claimed that its formal structure does not conform to the laws of classical propositional calculus, a belief that is widespread still to this day (cf.\ \cite{Giuntini:2002}, \cite{DallaChiara:2004} and \cite{deRonde:2016}). From the thirties, in fact, various modifications of Classical Logic (CL) were advanced in order to faithfully represent the physical content of quantum theory. For instance, in 1931 the Polish philosopher Zygmunt Zawirski proposed to apply \L ukasiewicz's three-valued logic to QM, starting from considerations about the wave-particle duality and Heisenberg's uncertainty principle. In virtue of the latter, moreover, the logical empiricists Moritz Schlick and Philipp Frank claimed that the conjunction of two or more statements may have no meaning in quantum theory, concluding that the classical conjunction is not always valid in the context of quantum physics, as recalled in \cite{Carnap:1966}. In their view, a typical example of invalid conjunction is one attributing a definite value for both position and momentum to a single quantum particle at a given time $t$. Furthermore, in 1933 Fritz Zwicky suggested that QM rejects the law of excluded-middle.\footnote{For historical details on quantum logic see \cite{Jammer:1974}.} However, it s generally recognized that 
the founding text on Quantum Logic (QL) is Birkhoff and von Neumann's essay ``\emph{The Logic of Quantum Mechanics}'' (\cite{vonNeumann:1936}), where the authors provided the standard account of the propositional calculus obeyed by quantum propositions.\footnote{It is worth noting that von Neumann in his treatise \emph{Mathematische Grundlagen der Quantenmechanik} published in 1932 anticipated that from the algebraic structure of quantum theory it would have been possible to formulate a new propositional calculus. These ideas would have been fully expressed of his successive collaboration with Garrett Birkhoff.}

A few years after the publication of this paper, the research in quantum logic had a notable development in both the scientific and philosophical communities, opening new research programs searching for the correct formal structures to represent the physics of QM (cf.\ \cite{Reichenbach:1944}, \cite{Mackey:1957}, \cite{Finkelstein:1963}, \cite{Kochen:1965}, \cite{Jauch:1969}, \cite{Giuntini:2002}, \cite{DallaChiara:2004}), and intense metaphysical debates regarding the nature of logic itself and its empirical status (cf.\ notably \cite{Quine:1951}, \cite{Putnam:1968}, \cite{Dummett:1976}, \cite{Hallett:1982}, \cite{Weingartner:2004}, \cite{Bacciagaluppi:2009}). In addition, it is worth noting that research in quantum logic is still very prolific to the present day with many interesting applications in the fields of quantum information, computation and cryptography. Therefore, although QL is no longer viewed as a solution to the foundational problems affecting quantum theory\footnote{This conclusion is now accepted among experts; for details see e.g.\ by \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}. Referring to this, the latter authors stated explicitly that ``quantum logics are not to be regarded as a kind of ``clue'', capable of solving the main physical and epistemological difficulties of QT [quantum theory]. This was perhaps an illusion of some pioneering workers in quantum logic'' (\cite{Giuntini:2002}, p.\ 225).}, it is an important discipline of contemporary physics. In this paper we will neither take a stance in the dispute concerning the ontological status of logic and its laws, nor we will consider recent developments of QL given that they are not strictly relevant to the point we are going to make. Indeed, the present essay is focused on the logical structures common to both quantum and Classical Mechanics (CM), following the spirit of the former discussion.\footnote{In the present essay we will be concerned only with propositional calculus in the context of non-relativistic QM. First order and higher order logics will not be discussed here as well as the logic of relativistic quantum theories.}

Recently one of us faced the question of whether quantum physics necessarily implies a revision of classical logic (cf.\ \cite{Oldofredi:2020}). Alternatively stated, it was asked if classical logic must be abandoned in the quantum realm, or there are quantum theories compatible with a classical propositional calculus. Analyzing Bohmian Mechanics, a well-known interpretation of QM, it has been showed that quantum physics does not necessarily involves a departure from classical logic. More precisely, it has been argued that in the context of this interpretation of QM not only the logical connectives retain their classical meanings---in particular the negation and the disjunction do follow classical rules---but also the distributivity law holds, contrary to the case of standard QM. This fact is a consequence of the clear ontology of the theory and its explanation of quantum measurements.

In this paper we will be concerned with a similar issue, since we ask whether standard quantum mechanics necessarily involves a new logical structure radically different with respect to classical propositional calculus. In what follows we will answer this question in the negative. More precisely, here we have a two-fold aim: in the first place, we will analyze the main motivations and justifications for the introduction of quantum logic arguing that they do not entail the failure of CL in the quantum context. In the second place, we show that from the mathematical structures on which QM is built, it is possible to provide a distributive lattice to quantum propositions, contrary to the received view about QL. Such a distributive lattice of quantum statements, moreover, will allow us to take into account also assertions concerning expectation values, which are generally excluded in the context of the standard approach to quantum logic. Finally, we will argue that it is possible to demonstrate that classical logical propositions can form a non-distributive lattice. All these aspects, to our knowledge, are not sufficiently discussed in literature, thus, the present paper can contribute to shed new light on the relation between logic and quantum mechanics.
\vspace{2mm}

The essay is structured as follows: Section \ref{QL} provides a brief outline of the standard approach to QL (readers familiar with Birkhoff and von Neumann's proposal can skip this part), while in Section \ref{Motivations} we review the main arguments that have been given in literature to introduce a quantum logical calculus. Here we explain  that they do not entail the failure of classical logic in QM. In Section \ref{Math} we show not only that quantum logical propositions generate a distributive lattice, but also that important sentences of quantum theory---i.e.\ those concerning expectations values---are excluded by standard QL.\ In Section \ref{conc} we discuss the main philosophical implications of the preceding sections and conclude the paper.

\section{The Standard Approach to Quantum Logic}
\label{QL}

In order to present the standard approach to QL, let us say a few words about the relation between classical logic and classical mechanics. It is a well known fact that classical propositional logic is equivalent to a Boolean algebra where propositions (atomic and complex) have truth values ``true'' and ``false'' (or $\top$ and $\bot$), and the main operations among them are conjunction ``$\wedge$'', disjunction ``$\vee$'' and negation ``$\neg$'', the only unary operation. From these logical connectives one can introduce secondary operations, as for instance the material implication ``$\longrightarrow$'', the exclusive or ``$\oplus$'' and logical equivalence ``$\longleftrightarrow$''. Furthermore, it is important for our discussion to underline that the laws of propositional logic include commutativity, associativity, identity and distributivity with respect to $\wedge$ and $\vee$.

Analogously, the algebraic structure underlying classical mechanics is Boolean as well, being commutative, distributive and associative; in this context the logical operations of conjunction, disjunction and negation are replaced respectively by multiplication, addition and complementation among the variables of the theory. This fact has an interesting physical significance, since the variables associated to magnitudes of physical systems can be added and multiplied together---i.e.\ one can sum and multiply measurement results. Both propositional logic and classical mechanics, thus, generate a complemented Boolean lattice; from this fact it follows that the observable algebra of CM ``is isomorphic to a Boolean algebra of propositions with $(\wedge, \vee, \neg)$'' (\cite{David:2015}, p. 79).
To this regard, Beltrametti claims that 
\begin{quote}
Boolean algebras are algebraic models of classical logic (more specifically of classical propositional calculus) with the algebraic operations of meet ($\wedge$) and join ($\vee$) corresponding to the logical connectives ``and'', ``or'', and the unary relation of orthocomplementation corresponding to the logical negation. The rules and tautologies of classical logic have their counterpart in equations and identities of Boolean algebras (\cite{Beltrametti:2004}, p.\ 341). 
\end{quote}

Therefore, it is possible to formally characterize the state of a physical system via logical propositions which can assume the truth values ``true'' or ``false''---depending whether such statements describe true or false state of affairs concerning the system under consideration (cf.\ \cite{Jaeger:2009}, p. 61)---and to perform logical operations among propositions via the connectives $(\wedge, \vee, \neg)$. It is worth noting that in the context of CM it is always determined whether a system instantiates a given property or not; alternatively stated, every logical proposition about physical systems is either true or false, and hence in CM the principle of semantic bivalence holds (for more details cf.\ \cite{Bub:2007}, p.\ 642, \cite{Giuntini:2002}, p.\ 130).

Taking instead into account the mathematical structure and the physical content of standard QM, it is generally claimed that the classical propositional calculus is no longer appropriate to represent the logic obeyed by quantum propositions, since quantum theory does not share the same algebraic structure of classical mechanics. In particular, the algebra of quantum observables is non-commutative and, from a logical perspective, generates a non-distributive orthocomplemented lattice, which is non-Boolean. 
Referring to this, given the axioms of QM, Birkhoff and von Neumann showed how the physical content of the theory does not conform to a classical logical structure. The principal aim of their essay on quantum logic was to ``discover what logical structures one may hope to find in physical theories which, like quantum mechanics, do not conform to classical logic'' (\cite{vonNeumann:1936}, p.\ 823). The expression quantum logic in this paper must be clarified, since it refers to a quantum propositional calculus in the form of a ``calculus of linear subspaces with respect to \emph{set products}, \emph{linear sums}, and \emph{orthogonal complements}'' which ``resembles the usual calculus of propositions with respect to \emph{and}, \emph{or}, and \emph{not}'' (\emph{ibid}.), where the logical propositions are associated to measurements, tests on quantum systems.\footnote{Cf.\ \cite{Giuntini:2002} and \cite{DallaChiara:2004} for a systematic introduction to various forms of quantum logics, and to \cite{Engesser:2009} for historical and philosophical discussions on the topic.}

Their analysis begins by defining quantum logical propositions as \emph{experimental propositions}, i.e.\ statements affirming that a certain observable $A$ (or a set of observables) measured on a quantum system $s$ has a given value $a_i$ (or a sequence of values). More precisely, the authors stress that both in classical and quantum mechanics observations of physical systems are given by readings of experimental outcomes ($x_1, \dots, x_n$) of \emph{compatible} measurements ($\mu_1, \dots, \mu_n$). The values ($x_1, \dots, x_n$) are elements of what the authors called the ($x_1, \dots, x_n$)-space, i.e.\ the ``observation-space'' of the system in question, whose elements are all the possible combinations of results of the compatible measurements ($\mu_1, \dots, \mu_n$). Hence, the actual values ($x_1, \dots, x_n$) form a subset of such a space. Birkhoff and von Neumann, then, defined the experimental propositions concerning a physical system as the subsets of the observation space associated with it.

Secondly, the authors underlined that in classical and quantum mechanics the states of physical systems are mathematically represented by points in their state spaces---phase space for the classical case, and Hilbert space $\mathcal{H}$ for the quantum case---which provide the maximal information concerning the system. A point in phase space corresponds to the specification of the position and momentum variables of a certain classical system, whereas in QM the points of $\mathcal{H}$ correspond to wave functions. The authors, then, find a connection between subsets of the observation-space of a system and subsets of its Hilbert space, specifying that quantum experimental propositions are mathematically represented by a closed linear subspace of $\mathcal{H}$; this step is crucial in order to obtain the quantum propositional calculus.\ Alternatively, we can say that quantum mechanical operators correspond to propositions with ``yes/no'' (``true/false'') outcome in a logical system\footnote{Cf.\ also \cite{David:2015}, p.\ 78, where we read that ``An orthogonal projector $\textbf{P}$ onto a linear subspace $P\subset\mathcal{H}$ is indeed the operator associated to an observable that can take only the values 1 (and always 1 if the state $\psi\in P$ is in the subspace $P$) or 0 (and always 0 if the state $\psi\in P^{\perp}$ belongs to the orthogonal subspace to $P$).\ Thus we can consider that measuring the observable $\textbf{P}$ is equivalent to perform a test on the system, or to check the validity of a logical proposition $p$ on the system''.} as underlined by Svozi:
\begin{quote}
Any closed linear subspace of---or, equivalently, any projection operator on---a Hilbert space corresponds to an elementary
proposition. The elementary \emph{true--false} proposition can in English be spelled out explicitly as
\begin{quote}
``The physical system has a property corresponding to the associated closed linear subspace'' (\cite{Svozi:1999}, p.\ 1).
\end{quote}
\end{quote}

Thus, since in quantum mechanics physical systems do not have well-defined values for their properties before the measurement of a given operator (contrary to their classical counterparts), or better that quantum observations do not reveal pre-existing values, we have to stress that such quantum propositions refer always to measurements or to preparation procedures. Hence, the sentence ``the particle $s$ has $x$-spin up'' means either that I have measured the observable $S_x$ and I have found the value $+\hbar/2$ associated with the state ``$x$-spin-up'', or equivalently that I prepared $s$ in the $x$-spin-up state. Thus, one cannot simply say, as in classical mechanics, that a certain system has a certain property without having prepared such a system in a given state or having measured it.\footnote{More on this below.}

Now that we have clarified what quantum propositions are, let us introduce Birkhoff and von Neumann's quantum logic. In order to properly define a propositional calculus for QM, one must define the logical operators for conjunction, disjunction and negation and the notion of logical implication. Following \cite{vonNeumann:1936}, the procedure is rather simple: 
\begin{enumerate}
\item The negation of a proposition $p$ is defined by the authors as follows: ``since all operators of quantum mechanics are Hermitian, the mathematical representative of the negative of any experimental proposition is the orthogonal complement of the mathematical representative of the proposition itself'' (\cite{vonNeumann:1936}, pp.\ 826-827). The orthogonal complement $P^{\perp}$ of the subspace $P$ is the set whose elements are all the vectors orthogonal to the elements of $P$. Such an orthogonal complement satisfies the following property: given a subset $P\subset\mathcal{H}$ and a pure state $\psi$, $\psi(P)=1$ iff $\psi(P^{\perp})=0$ and $\psi(P)=0$ iff $\psi(P^{\perp})=1$. As Dalla Chiara and Giuntini underline, ``$\psi$ assigns to an event probability 1 (0, respectively) iff $\psi$ assigns to the orthocomplement of $P$ [notation adapted] probability 0 (1, respectively). As a consequence, one is dealing with an operation that \emph{inverts} the two extreme probability-values, which naturally correspond to the truth-values \emph{truth} and \emph{falsity} (similarly to the classical truth-table of negation)'' (\cite{Giuntini:2002}, p.\ 132).
\item Concerning the conjunction, Birkhoff and von Neumann notice that one can retain the very same set-theoretical interpretation of the classical conjunction in QL, since the intersection of two closed subspaces $P, Q\subset\mathcal{H}$ is still a closed subspace. Thus, one maintains the usual meaning for the conjunction: the pure state $\psi$ verifies $P\cap Q$ iff $\psi$ verifies both $P$ and $Q$. Thus, quantum logic does not introduce a new logical operator for the conjunction.
\item Contrary to the previous case, the logical operator for disjunction cannot be represented by the set-theoretic union, since the set-theoretical union of the two subspaces $P, Q$ will not be in general a subspace, thus, it is not an experimental proposition. Therefore, in QL one introduces the quantum logical disjunction as the the closed span of the subspaces $P, Q$, which is an experimental proposition. Such a statement corresponds to the ``smallest closed subspace containing both $P$ and $Q$'' (\cite{Bacciagaluppi:2009}, p.\ 55). 
\item The logical implication is defined by set theoretical inclusion: given two experimental propositions $p$ and $q$, 
$p$ implies $q$, means that whenever one predicts $p$ with certainty, one can predict also $q$ with certainty, and this is equivalent to state that $p$ is a subset of $q$. This fact is particularly important since the authors showed that ``it is \emph{algebraically} reasonable to try to correlate physical qualities with subsets of phase-space'' and thus, ``physical qualities attributable to any physical system form a partially ordered system'' (\cite{vonNeumann:1936}, p.\ 828, notation adapted).
\end{enumerate}

In this manner we have defined a propositional calculus for quantum experimental propositions, which generates an orthocomplemented lattice. In what follows we analyze the motivations for such a quantum logic, clarifying why it is generally believed that classical propositional calculus cannot be applied in QM.

\section{Quantum Disjunction and Distributivity Law}
\label{Motivations}

The first remarkable difference between QL and CL is the failure of the distributive law; moreover, analyzing the motivations for which \eqref{DL} does not hold in the quantum context it will be relevant to see also other important differences between these logics. It is a well-known fact that the propositional calculus underlying classical mechanics is distributive in virtue of the commutative algebra of classical observables. Referring to this, \cite{Giuntini:2002} claim that distributivity fails in quantum logic for two main reasons:
\begin{itemize}
\item the non-commutativity of quantum observables,
\item the peculiar behavior of the quantum disjunction---the second major deviation from classical logic---which may be true although neither of the member is true in virtue of the linearity of the Schr\"odinger Equation (SE). 
\end{itemize}
\noindent Let us then have a closer look to these motivations for the failure of the distributivity law, and more generally of CL.

Contrary to the case of classical mechanics, QM relies on a non-commutative algebraic structure\footnote{For details on the mathematical structure of QM and its physical content the reader may refer e.g.\ to \cite{Sakurai1994}, and \cite{Griffiths:2014}. In this essay, it is assumed that the reader has some familiarity with the standard formalism of QM.}: quantum observables generally do not commute, meaning that for any pair of operators $A,B$ we have $[A,B]=AB-BA\neq0$. From a physical perspective this fact entails that if one performs a measurement of $A$ on a quantum system followed by a measurement of $B$, in general one will obtain different results inverting the order of these observations. Moreover, in virtue of the non-commutative algebra of quantum theory, to measure an operator $C=A+B$ is not typically equivalent to measure $A$ and $B$ independently and then add the respective results, since $A,B$ will be in general incompatible (cf.\ \cite{David:2015}, p.\ 77). This feature of QM constitutes the formal basis to prove the Heisenberg uncertainty relation, a theorem of quantum theory which reflects the operational inability to simultaneously measure the values of incompatible operators with arbitrary precision. Such a result, in turn, is usually interpreted ontologically in the sense that quantum system do not instantiate definite properties in non-measurement situations (cf.\ \cite{Sakurai1994}).

In addition, as we will explain in the remainder of this section, the failure of distributivity in quantum logic is due to another particular characteristic of this new propositional calculus, namely the behavior of the quantum disjunction, which is different w.r.t.\ to its classical counterpart in virtue of the presence of quantum superpositions. Indeed, another remarkable difference between QM  and CM is that quantum systems can be in superposition states as a consequence of the linearity of the Schr\"odinger equation---the fundamental dynamical law of quantum theory---which for a single particle reads: 
\begin{align}
\label{SE}
i\hbar\frac{\partial\psi}{\partial t}=\Big(-\frac{\hbar^2}{2m}\nabla_k^2+V\Big)\psi=H\psi,
\end{align}
\noindent where $H$ represents the Hamiltonian operator, defined as the sum of kinetic and potential energy of the system at hand. More specifically, this algebraic property of \eqref{SE} entails that if two wave functions\footnote{In QM the wave function of a system provides the maximal information available about it.} $\psi_1, \psi_2$ are both possible solutions of the same Sch\"odinger equation, then their linear combination (\emph{superposition}) 
\begin{align}
%\label{SSE}
\psi_s=\alpha\psi_{1}+\beta\psi_{2} \nonumber
\end{align} 
is still a solution of the same SE---$|\alpha|^2,|\beta|^2$ (with $\alpha,\beta\in\mathbb{C}$) represent the probabilities to find the system in $\psi_1, \psi_2$ respectively. Notably, the new superposed state $\psi_s$ is also a consistent representation of the system. Referring to this,  \cite{deRonde:2016} explicitly say that in QL 
\begin{quote}
differently from the case in classical semantics, a quantum disjunction may be true even if neither of it members is true. This reflects, for example, the case in which we are dealing with a state such as that of a spin $1/2$ system which is in a linear combination of states up and down.
\end{quote}
\noindent This is exactly the case that we are going to consider in order to review the usual explanation provided for the failure of the distributivity law in the quantum context.

To claim that in QL conjunction and disjunction are not distributive means that given three propositions $p, q$ and $r$ the logical law
\begin{align}
\label{DL}
p\wedge(q\vee r)\longleftrightarrow(p\wedge q)\vee(p\wedge r)
\end{align}

\noindent does not hold.\footnote{In the context of quantum logic, it is usually believed that distributivity must be replaced by a weaker law: $(p\wedge q)\vee(p\wedge r)\longrightarrow(p\wedge(q\vee r)).$}
Let us illustrate this fact with a simple example taken from \cite{Giuntini:2002}. Suppose to consider a $1/2-$spin particle, let's say an electron. From QM we know that there are only two possible spin states in which a particle can be found along each axis after a spin measurement, namely either in the up or in the down state; for the case at hand we assume that this particle has $x$-spin up. Furthermore, since spin operators along different axes do not commute, being incompatible observables---i.e.\ $[S_x, S_y]=[S_x, S_z]=[S_y, S_z]\neq0$---they cannot be simultaneously measured. Then, as a consequence of the Heisenberg uncertainty principle, ``both propositions ``$spin_y$ is up'' and ``$spin_y$ is down'' shall be strongly undetermined. However, the disjunction ``either $spin_y$ is up or $spin_y$ is down'' must be true'' (\cite{Giuntini:2002}, pp.\ 133-134), and the same holds for the case of spin in the $z$ direction. Therefore, if in the distributivity law \eqref{DL} we suppose that
\begin{itemize}
	\item $p$ represents the proposition ``the electron has $x$-spin up'', 
	\item $q$ the proposition ``the electron has $y$-spin up'',
	\item $r$  the proposition ``the electron has $y$-spin down'',
\end{itemize}
\noindent it is possible to see that the left hand side of the bi-conditional expresses a true statement---i.e.\ $p\wedge(q\vee r)=\top$---, since we assume that $p$ is true and we know that the disjunction $(q\vee r)$ is also true, whereas $(p\wedge q)$ and $(p\wedge r)$ are both false in virtue of the incompatibility between the spin operators among different axes. This fact implies that the right hand side of the distributive law asserts a false statement, i.e.\ $(p\wedge q)\vee(p\wedge r)=\bot$. Consequently, this law fails in the context of quantum theory, because we have 
\begin{align}
\label{contradiction}
p\wedge(q\vee r)=\top\longleftrightarrow(p\wedge q)\vee(p\wedge r)=\bot. 
\end{align}

In what follows, we will argue that these motivations for the introduction of QL are not sufficient to claim that CL is rejected by quantum theory, contrary the received view about quantum logic. More precisely, in the remainder of the section we are going to argue that (i) it is not necessarily the case that a quantum disjunction must be true when neither of its members are, and (ii) that \eqref{DL} can be retained in QM, since the argument seen above---usually meant to show its failure---contains serious flaws. In the next section, furthermore, we will show that QL can be given a distributive lattice of propositions.

\subsection{The Quantum Disjunction}

As stated above, there seems to be a strict connection between the features of quantum disjunctions and superposition states. Indeed, in quantum logic it is commonly accepted that a disjunction can be true although neither of its members are, as for instance in the example just discussed. In this case we assumed that the spin of a particle in the $y$ direction was not determined since no measurement of $S_y$ was performed, thereby the electron under consideration was in a superposition of $y$-spin-up and $y$-spin-down states. Then, if we introduce the following propositions
\begin{description}
	\item $q =$ ``The particle has spin $y^+$''
	\item $r =$ ``The particle has spin $y^-$'',
\end{description}
\noindent quantum logic establishes that even though neither of these statements have definite truth-value (or neither of them is actually true), their disjunction $(q \vee r)$ is true---violating consequently the bivalence principle, the third main difference between CL and QL. The explanation of this fact takes into account spin measurements; indeed, when we actually measure the observable $S_y$ on the quantum particle at hand, we will find with certainty only one of the possible outcomes stated above. Thus, it must be the case that after a measurement of the spin along the $y$ direction either $q$ or $r$ will become true.

Nonetheless, it is possible provide an argument claiming that a quantum disjunction may be \emph{not always} true when both its elements have indeterminate truth values, undermining this widespread belief. Our reasoning starts from some remarks concerning the very definition of quantum logical proposition.

In classical mechanics one generally considers physical systems as objects having well-defined values for inherent, dynamical and relational properties, these attributes moreover do not depend on any observation---meaning that measuring a magnitude on a classical particle will reveal a pre-existing value of the observed quantity. Similarly, the logical propositions referring to classical physical systems describe their actual ontic states, i.e.\ these statements regard state of affairs of the systems under consideration which do not depend on any external observer. Alternatively stated, since in the classical case observations reveal the pre-existing values of the qualities attributed to the particles independently of any measurement, we can consider classical experimental propositions as statements about the properties constituting the actual ontic state of classical systems. Thus, statements about measurements and assertions concerning the ontic state of classical systems can be conflated without any dangerous metaphysical consequence.

However, in the context of QM this move is not allowed, given that properties of quantum systems acquire definite values only in measurement situations, as we already stated above. Consequently, one can attribute a definite value for a specific magnitude to a given system only after the measurement of the quantity under consideration, or after having prepared a quantum object in a specific state.\footnote{In operational quantum theory the preparations are a set of instructions that a certain agent has to follow in order to prepare a quantum system in a certain state.} Quantum mechanics, in virtue of the Born's rule, then provides the probability to obtain a certain result among a set of possible outcomes associated with eigenstates and eigenvalues of the measured observable. Thus, quantum logical propositions like ``the particles has spin $y^+$'', ``the velocity of the particle is $v$'', ``the position of the particle is $x$'' \emph{etc.}\ do \emph{not} refer to inherent properties of quantum systems---i.e.\ to their ontic state---but rather they concern the interactions taking place between quantum systems and experimental devices either in preparation procedures or measurement situations. These kind of propositions must be completely indeterminate, being neither true nor false, with respect to the actual state of quantum systems when they do not interact, in complete agreement with the physical content of QM.

The complex proposition $q \vee r$ is considered true before any spin measurement in virtue of the conditional claim according to which if one performs a measurement of $S_y$, then either $y^+$ or $y^-$ \emph{will be found}. Such a statement, then, refers to a classical macroscopic device which, after the performance of a $y$-spin measurement, will indicate with certainty either the outcome $y$-spin-up or $y$-spin-down.\footnote{Clearly we assume that any measurement will yield with certainty only one result.} Nonetheless, neither such quantum disjunction, nor its atomic components tell us something about the inherent property associated with the observable $S_y$ possessed by a quantum particle independently of any measurement of $S_y$.

Referring to this, it is then possible to highlight another peculiar feature of quantum propositions which is relevant for our argument. If we consider the sentence $p=$ ``the particle has spin $x^+$'', we know that this statement can be either true of false only after the actual performance of a $x$-spin measurement on a certain particle\footnote{Or after having prepared the system at hand in a definite $x$-spin state.}: this assertion will be true in the case the experimenter will find the $x^+$ outcome and false otherwise. Before the performance of this observation, however, it has an indeterminate truth value. Nonetheless, we can say that $p$ has a precise meaning even prior to any $x$-spin measurement, since it may be understood as a sentence referring to the expectation value of the $S_x$ operator.  
Hence, in non-measurement situations the propositions $q, r$ above may be similarly regarded as experimentally equivalent to:
\begin{description}
	\item $q' =$ ``The expectation value for $S_y$ is 1/2$\hbar$''
	\item $r' =$ ``The expectation value for $S_y$ is -1/2$\hbar$''.
\end{description}
%As a consequence, proposition $q \vee r$ corresponds to the following statement: 
%\begin{center}
%``The absolute value of the expectation value for $S_y$ 1/2$\hbar$''.
%\end{center}
In this case, it is straightforward to see that the sentence $q' \vee r'$ does not refer to macroscopic devices, but it says something about the possible values of $S_y$. If we have a superposition of the $y$-spin-up and $y$-spin-down states---as in the previous example and thereby a disjunction whose elements are $q'$ and $r'$--- then, the expectation value for $S_y$ will be strictly between $-1/2 \hbar$ and $1/2 \hbar$, and therefore $q' \vee r'$, if referred to expectation values will be false as one would expect, given that $S_y$ may take, before the performance of a certain measurement, any value in this interval. Therefore, contrary to the received view in quantum logic, it is not always the case that a quantum disjunction is true even though neither of its members are.

This discussion highlighted a crucial ambiguity in quantum propositions like ``the particle has spin up/down along such and such direction'', since on the one hand they may refer to measurements results, and on the other hand they may concern expectation values for some spin observable. In order to disambiguate such experimental propositions, one has to take into account whether we are evaluating propositions regarding expectation values or measurement outcomes. This conclusion will be helpful in the following subsection where we show how to keep the distributivity law in quantum logic.

\subsection{Keeping Distributivity in Quantum Mechanics}

As we already said, \cite{Giuntini:2002} have argued that the distributivity law fails in quantum mechanics in virtue of (i) the alleged peculiar behavior of the quantum disjunction and (ii) the non-commutativity among spin operators defined along different axes. Recalling their example, we introduced the following three propositions

\begin{description}
    \item $p =$ ``The state of particle is $x^+$''
    \item $q =$ ``The state of particle is $y^+$''
    \item $r =$ ``The state of particle is $y^-$''
\end{description}

\noindent assuming that $p$ is true. This fact entails two consequences that should be explicitly stated: on the one hand, at a certain time $t$ an experimenter actually performed a measurement of $S_x$ and found the result $x^+$, and on the other hand, no observations of $S_y$ have been made at $t$. Moreover, the story goes, if one would measure the spin along the $y$ axis at later time $t'$, one would either get $y^+$ or $y^-$, hence, we say that $q \vee r \equiv \top$. Therefore, $p \wedge (q \vee r)$ is true, since both $p$ and $(q \vee r)$ are.

However, both the conjunctions on the right hand side of the distributivity law are false because $p \wedge q$ and $ p \wedge r$ refer to incompatible observations: QM does not allow a simultaneous measurement of these quantities, therefore, we have $(p \wedge  q) \vee (p \wedge  r) \equiv \bot$. As a consequence, the two sides of the bi-conditional expressing the distributive law generate a contradiction, so that one concludes that \eqref{DL} is not valid in quantum logic, and thereby that the lattice of quantum proposition is non-distributive.

In order to understand the reasons why this example does not show that distributivity fails in quantum theory, it is worth noting that the left hand side of the distributivity law contains propositions referring to different instants of time: indeed, in our propositions $p, q$ and $r$ we are considering the very same electron, and we assume that it has spin $x^+$, meaning that an agent performed at particular time $t$ a measurement of $S_x$ and found $x^+$ as a determinate outcome. Moreover, in virtue of the commutation relation $[S_x, S_y]\neq0$, we know that the particle at hand has indeterminate spin along the $y$ axes at $t$, hence, this quantum object is not in a definite eigenstate of $S_y$. Nonetheless, the disjunction $(q \vee r)$ is assumed to be true since if we would measure $S_y$ \emph{at a later time} $t'>t$ we would find only one of the possible results contained stated in $q$ and $r$---i.e.\ either spin $y^+$ or $y^-$. Taking into account the right hand side of the distributivity law, however, one is evaluating all the propositions $p, q$ and $r$ \emph{at the very same time} $t$: indeed, we ask whether at $t$ the particle has spin $x^+$ and simultaneously either spin $y^+$ or $y^-$. As already said, quantum theory forbids a simultaneous measurement of non-commuting quantities, and thus, we obtain the desired contradiction. 
In sum, \eqref{DL} fails because we considered propositions evaluated at different times $t$ and $t'$ in the left hand side of the distributivity law, whereas the propositions on the right hand side are considered uniquely at time $t$. Rewriting \eqref{DL} making explicit the temporal dependence of the three propositions, we immediately say that the above example showing the failure of distributivity in quantum logic is flawed:

$$p_t \wedge (q \vee r)_{t'} \longleftrightarrow (p_t \wedge q_t) \vee (p_t\wedge r_t).$$

\noindent It is in general true that $q, r$ will have a definite truth value at time $t'$, since a measurement of $S_y$ is suppose to happen and, therefore, at $t'$ the disjunction $(q \vee r)_{t'}$ will be true. The distributivity law, thus, is violated not in virtue of an inherent quantum mechanical motivation or principle, but simply because we evaluate the three assertions at different times---and thereby we have to do with different statements---in the two sides of the bi-conditional.

Now, as in the previous case of the quantum disjunction, it is possible to argue that the distributivity law is not necessary violated in quantum logic if we take carefully into consideration the meaning of quantum propositions in relation to (i) the physical content of QM, and (ii) their temporal order. In the first place, let us introduce a temporal parameter $t$ in order to evaluate the above sentences at the very same instant of time:

\begin{description}
    \item $p_t =$ ``The state of particle at time $t$ is $x^+$''
    \item $q_t =$ ``The state of particle at time $t$ is $y^+$''
    \item $r_t =$ ``The state of particle at time $t$ is $y^-$''.
\end{description}

\noindent Also in this case let us assume that $p_t$ is true. Then, in virtue of the formalism of quantum theory, we conclude that at time $t$ both $q_t$ and $r_t$ must be false, since the quantum object under consideration is not in a definite eigenstate of $S_y$ at $t$. This fact implies that the left hand side of the distributivity law $p_t \wedge (q_t \vee r_t)$ is false, as well as the right hand side $(p_t \wedge  q_t) \vee (p_t \wedge r_t)$ since both elements of the disjunction are false. Thus, taking into account a temporal parameter, it is easy to see that both sides of the bi-conditional expressing the distributivity law have the same truth value

$$p_t \wedge (q_t \vee r_t) \equiv \bot \equiv (p_t \wedge q_t) \vee (p_t\wedge r_t)$$

\noindent  establishing that \eqref{DL} is \emph{not} violated.

Finally, let us conclude this section by showing that a failure of the distributivity can happen also in the context of classical mechanics if we take into account three propositions evaluated at different temporal instants in the two sides of the law. Suppose we have a ball sitting at position $p$ over a hatch at a certain time $t$ (see Figure 1). If the hatch opens at a later time $\hat{t}$, the ball drops, bounces around, until it will rest at either position $q$ or $r$ with equal chance.

\begin{center}
\includegraphics[scale=.8]{Ball.eps}
\end{center}

Firstly, let us consider the following statements:

\begin{description}
    \item $p =$ ``The position of the ball at time $t$ is $p$''
    \item $q =$ ``The position of the ball at time $t$ is $q$''
    \item $r =$ ``The position of the ball at time $t$ is $r$''.
\end{description}

\noindent It is straightforward to see that these are three incompatible assertions: since the ball can have only one possible determinate position in space at a given time $t$---in this example it is assumed that it is located at $p$---this object cannot be localized in two different locations, $p$ and $q$ or $p$ and $r$. Since both sides of the bi-conditional defining the distributivity law are false, we conclude that \eqref{DL} holds.

Now, if we assume that the hatch is actually opened and that at later time $\hat{t}$, the ball has settled either in $q$ or $r$. Then, we introduce the new sentences:

\begin{description}
    \item $\hat{q} =$ ``The position of the ball at time $\hat{t}$ is $q$''
    \item $\hat{r} =$ ``The position of the ball at time $\hat{t}$ is $r$''.
\end{description}

In this case, at time $\hat{t}$ the disjunction $(\hat{q} \vee \hat{r}) \equiv \top$ because the ball must be in one of the two positions after it dropped. Hence, taking into account the left hand side of the distributivity law, if we evaluate $p$ at $t$ and $q, r$ at $\hat{t}$ we see that $p_t \wedge (\hat{q} \vee \hat{r}) \equiv \top$, since $p$ is hold true as well as the disjunction $(\hat{q} \vee \hat{r})$. However, if we consider the right hand side of \eqref{DL}, evaluating $p, q$ and $r$ at time $t$---as done in the quantum case---we will trivially obtain that $(p \wedge  q) \vee (p \wedge r)\equiv \bot$. Thus, a contradiction can be deduced also in classical mechanics if we evaluate sentences at different times in the two sides of the bi-conditional. Nonetheless, one would not say that distributivity fails in the classical context, but rather that we have to evaluate propositions taking into account the same temporal instant in both sides of \eqref{DL}.

In conclusion, contrary to the received view in quantum logic, we have shown that the distributivity law is not violated in quantum context once the correct temporal order of quantum propositions---and consequently their actual meaning---is considered.

\section{On the Common Logical Structure between Quantum and Classical Mechanics}
\label{Math}

After having critically assessed the general motivations to introduce quantum logic as a consequence of the physical content of QM, in this section we present our main argument according to which it is possible to assign a distributive lattice to quantum propositions. Remarkably, in what follows we will see that the non-distributivity of the quantum logic lattice does neither originate from non-commuting observables nor from superpositions of quantum mechanics: it is a property of the lattice of subspaces of \emph{any} vector space, and of the lattice of subgroups of any group more in general (cf.\ \cite{Davey:2002}). To this regard, indeed, we also show that taking into account the space of classical probability distributions one can apply a non-distributive lattice to classical propositions. This fact is due to the similar structures of the state space of quantum mechanics and the space of classical probability distributions.

In what follows, then, it will be argued that both theories---quantum and classical---may be assigned \emph{two} lattices of propositions: a distributive one that follows the laws of classical logic, and a non-distributive one obeying the rules of quantum logic. Contrary to the widespread opinion, therefore, we will claim not only that quantum physics does not necessarily entails a revision of classical propositional calculus, but also that it is actually possible to show that quantum propositions can form a distributive lattice. Furthermore, in agreement with the example of the violation of the distributivity law in the classical context seen a few lines above, in the remainder of the section we will argue also that the propositions of classical mechanics may obey a non-classical type of logic.

Before we proceed, it is worth noting that the classical space we will be using is not phase-space itself, but the space of probability distributions over phase-space, which is a real Hilbert space and thus has a similar structure to its quantum counterpart. To this regard, it should be stressed that probabilistic theories are important even in classical mechanics. Due to the nature of experimental verification, also in the context of classical mechanics there are always going to be uncertainties in the preparation and measurements, and repeated measurements will directly lead to statistical constructs.
However, we want to emphasize that this paper aims primarily at showing that QM can be given a distributive lattice of propositions. Here we want to enter neither into debates concerning the ontology of quantum mechanics, nor into discussions concerning the similarities between the commitments of QM and those of classical mechanics.  In this essay we simply point out a formal similarity between the above mentioned mathematical structures that has implications for the logic of both theories.

From a mathematical perspective, the distributive lattice will be provided in both classical and quantum case by a Borel algebra, which is the smallest $\sigma$-algebra on a set $X$ equipped with a topological structure that contains the topology; the non-distributive lattice will be provided by the set of closed subgroups ordered by inclusion. We will claim that the non-distributive lattice is a strict subset of the distributive one; this result entails that everything that can be described in quantum logic can also be described in classical logic. Moreover, the distributive lattice will include propositions, like sentences concerning the expectation value of observables that are of high scientific interest, since in an experimental context by ``measurement'' one typically refers to the statistical collection of many ``single takes''. On the contrary, it is well-known that standard quantum logic deals only with single-take measurements, remaining silent w.r.t.\ propositions involving statistical statements. Hence, measurements of cross sections, the most typical type of measurement in the context of particle physics, are not part of the quantum lattice, and neither are measurements of the probabilities themselves. Therefore, showing that one may apply a distributive lattice to quantum propositions can be seen as an important result in order to extend the set of meaningful logical propositions of quantum mechanics.

\subsection{The Distributive Lattice of Quantum Mechanics}

As anticipated above, in this section our arguments are fundamentally based on the notion of $\sigma$-algebra, which is a significant structure from a foundational perspective, given that physically relevant mathematical branches like measure theory (cf.\ \cite{Cohn:2013}) and probability theory (cf.\ \cite{Grimmett:2001}) are built on top of it. Let us then review some of its essential features.

In order to begin our discussion, let us consider in the first place a set $X$: in the quantum case it will correspond to the set of all possible quantum states; similarly, in the classical case it will correspond to the set of all possible probability distributions over phase space.\footnote{We should stress, however, that $\sigma$-algebras are defined independently of what the elements of $X$ actually are. For a simpler example, $X$  could correspond to the set of the real numbers for a continuous quantity.} By definition, a $\sigma$-algebra is a collection of subsets of $X$ closed under countable union, countable intersection and complement with respect to $X$; the closure condition requires that  performing those operations on the sets contained in a $\sigma$-algebra will return another set in the same $\sigma$-algebra. Interestingly, these sets represent statements concerning the objects described by $X$, i.e.\ quantum states and classical probability distributions.\footnote{In the quantum case this is exemplified also in Section \ref{QL}.} Then, in virtue of the equivalence between sets and statements, it will be more convenient for us to conceive $\sigma$-algebras as collections of propositions about $X$ instead of sets. Referring to this, since the $\sigma$-algebra is closed under countable union, countable intersection and complement with respect to $X$, the corresponding statements will be closed under countable conjunction, countable disjunction and complement. Hence, the $\sigma$-algebra generates a Boolean algebra of statements.\footnote{Note that the interpretation used here is exactly the same used in probability theory. In fact, a probability space is defined by three elements: a sample space $\Omega$ which represents the set of all possible outcomes; a $\sigma$-algebra over $\Omega$ which represents all events, all propositions; and a measure that assigns a probability to each event.}

In the second place, it is crucial to underline that every topological space---as for instance the complex Hilbert space of quantum mechanics and the real Hilbert space of classical probability distributions considered in this essay---comes already equipped with a $\sigma$-algebra called Borel algebra.\ A topology, in fact, is also a collection of sets---called \emph{open sets}---and the Borel algebra is defined to be the smallest $\sigma$-algebra that contains all the open sets.\footnote{Here the topology is important for us only as a mean to construct the Borel algebra.} Complete normed spaces, such as the mentioned Hilbert spaces, have a canonical way to construct a topology and, therefore, a $\sigma$-algebra. Let us see how this procedure works in the easier case of a standard three-dimensional Euclidean space. Given two points $x$ and $y$, we can write the square of distance $\epsilon^2 = |y - x|^2$ as the square of the norm of the vector between the two points. The statement ``the position is within $\epsilon$ meters of $x$'' corresponds to the set $U = \{y \; | \; |y - x|^2 < \epsilon^2 \}$ of all the points whose distance from $x$ is less than $\epsilon$. These are exactly the open sets that generate the topology, and the Borel algebra will be the smallest $\sigma$-algebra that contains these sets. This works analogously for all normed spaces. A Hilbert space, in particular, comes with an inner product which also defines a norm. We can define a distance $\epsilon^2 = |\psi - \phi|^2=\langle \psi - \phi , \psi - \phi \rangle$ between two states, and the statement ``the state of the system is within $\epsilon$ of $\psi$'' corresponds to the set $U = \{\phi \; | \; |\psi - \phi|^2 < \epsilon^2\}$. As we have already stated, the space of classical probability distributions is also a Hilbert space where each point corresponds to the square root of the density.\footnote{The choice of the square root of the density over the density itself is merely a technical choice.} In this case we have a distance $\epsilon^2 = |\sqrt{\rho_1} - \sqrt{\rho_2}|^2=\langle \sqrt{\rho_1} - \sqrt{\rho_2} , \sqrt{\rho_1} - \sqrt{\rho_2} \rangle$ and we can proceed in the same exact way. In all normed spaces, the topology, and therefore the $\sigma$-algebra, is constructed from sets of the form ``the object is within $\epsilon$ of $x$''.

It is a simple result that the Borel algebra thus constructed contains all propositions of quantum logic, and we state it in the following proposition:

\begin{prop}
	Let $\mathcal{H}$ be a Hilbert space. Let $\Sigma$ be the Borel $\sigma$-algebra induced by its inner product. Let $L(\mathcal{H})$ be the lattice of closed subspaces of $\mathcal{H}$. Then $L(\mathcal{H}) \subset \Sigma$.
\end{prop}

Proof. Let $p \in L(\mathcal{H})$ be a closed sub-space of $\mathcal{H}$. Then $p$ is a closed set, the complement of an open set, and therefore a Borel set: $p \in \Sigma$. Let $v \in \mathcal{H}$ be a vector in the Hilbert space and consider the singleton $U = \{ v \}$. This is not a closed sub-space and therefore $U \notin L(\mathcal{H})$. As the topology of a Hilbert space is Hausdorff, all singletons are closed and therefore $U \in \Sigma$.\footnote{To give more context, the set $U = \{ v \}$ can be written as $U = \{w \; | \; |w - v|^2 = 0\}$ the set of all elements with zero distance from $v$. This can be understood as the limit (i.e.\ countable intersection) of a sequence of propositions ``the object is within $\epsilon$ of $v$'' where $\epsilon$ tends to zero.} This means $L(\mathcal{H}) \subset \Sigma$. This concludes the proof.

The above proposition contains our first crucial conclusion. As we stated in Section \ref{QL}, in quantum logic the propositions correspond to closed subspaces, which are closed sets and, thus, are Borel sets. All the propositions of quantum logic are consequently readily found in the Borel algebra of the Hilbert space. The Borel algebra, in turn, is a Boolean algebra that follows classical logic, and contains all the statements of quantum logic. This construction also applies identically for the classical cases, where we will also be able to talk about the set of all closed subspaces.

While all of this may seem suspect to some reader, we would like to stress that these are very standard constructions. Alternatively stated, all of what has been mentioned so far are the standard mathematical techniques used to study these spaces and can be found in any textbook (cf.\ \cite{Rudin:1991}, \cite{Vasudeva:2017}). Here we are bringing out the full physical significance of these mathematical constructs, which is too often overlooked. Understanding the link between the mathematical structures and the physics they represent is crucial to clearing out possible misunderstandings.

Now that we have briefly seen the definitions and basic constructions, we can gain more insights into these structures and what they represent. First, let us understand why the lattice of subspaces (i.e. quantum logic) fails to be a Boolean algebra. Stone's representation theorem tells us that any Boolean algebra can be expressed as an algebra of sets using the standard set operations (i.e.\ intersection, union and complement) (cf.\ \cite{Davey:2002}). For the lattice of subspaces, note that the disjunction $p \vee q$ does not correspond to the set union of the subspaces, but the span. In fact, the union of subspaces is not part of the lattice.  This is why the lattice, even if it is a complemented $0-1$ lattice, fails to be a Boolean algebra: it does not use the standard set operations.

Clearly, $\sigma$-algebras are Boolean algebras because they are exactly an algebra of sets with standard set operations. As we said before, the Borel algebra will contain the lattice of subspaces plus all the set operations between them, including the union. In a way, the lattice of subspaces is not Boolean because it does not contain enough sets, which are instead contained in the Borel algebra.

Arrived at this point, one may ask: do these extra sets correspond to physically meaningful propositions? To answer this question, let us first demonstrate the following proposition.

\begin{prop}
	Let $\mathcal{H}$ be a Hilbert space and $\langle \, , \, \rangle$ be its inner product. Let $X$ be a self-adjoint bounded linear operator. Let $F_X : \mathcal{H} \to \mathbb{R}$ be the map defined by $F_X(\psi) = \langle \psi , X \psi \rangle$. Let $U \subseteq \mathbb{R}$ be a Borel set. Then $F_X^{-1}(U)$ is a Borel set.
\end{prop}

Proof. Let $X$ be a self-adjoint bounded linear operator. Recall that a linear operator is continuous if and only if it is bounded. Therefore $X$ is continuous. Note that $F_X$ is the composition of $X$ with the inner product. Recall that the inner product is a continuous function. Since $F_X$ is the composition of continuous functions it is also continuous. Recall that all continuous functions are also Borel measurable, therefore $F_X$ is Borel measurable. Let $U \subseteq \mathbb{R}$ be a Borel set. Since $F_X$ is Borel measurable, the reverse image $F_X^{-1}(U)$ is also a Borel set. This concludes the proof.

Since $\langle \psi , X \psi \rangle$ expresses the expected value for operator $X$, the proposition shows that the Borel algebra can express statements about the expectation value of observables,\footnote{Technically, position and momentum are not bounded operators. However, one can argue that measured position and momentum are indeed bounded as the acceptance of a detector is always limited. Note that all operators for which the spectral theorem applies can be written as the limit of a sequence of bounded operators.} such as ``the expected position $\langle x \rangle$ of the particle is between $x_0$ and $x_1$.'' This would correspond to the Borel set $\{ \psi \in H \, | \, \langle \psi , X \psi \rangle \in (x_0, x_1) \}$.\footnote{In most cases, one assumes states to be normalized, which is what we have done here for simplicity. Renormalization would lead to the set $\{ \psi \in H \, | \, \langle \psi , X \psi \rangle / \langle \psi , \psi \rangle \in (x_0, x_1) \}$, which is still a Borel set since renormalization is a continuous function over its domain $H \setminus \{ 0 \}$.}

Borel sets also allow us to express statements such as ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''. This is \emph{not} equivalent to the statement ``the particle is in the eigenstate $x_0$''. This latter proposition will be true only if the whole function is at $x_0$. That is, if the $x_0$ is the support of the wave-function. The former is a lot less stringent: the wave-function can even be distributed over an infinite range (e.g.\ a Gaussian wave packet) just as long the average is $x_0$. This means that we can have statements like ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''$\wedge$ ``the expected momentum $\langle p \rangle$ of the particle is exactly $p_0$'' without having contradictions (e.g.\ a Gaussian wave packet can be centered on any value of position and momentum).

We can go one step further. Note that all statements that characterize the position and momentum of the center of mass form a sub-algebra of the Borel algebra. Recall that the Ehrenfest theorem states that $m \frac{d}{dt}\langle x \rangle = \langle p \rangle$ and $ \frac{d}{dt}\langle p \rangle = - \langle V(x) \rangle$. If the potential for a quantum particle can be considered constant over the wave-function (i.e. $\langle V(x) \rangle = V(\langle x \rangle)$ ), the motion of the center of mass will reduce to the case of a point-particle. The Borel algebra contains as a sub-algebra the statements we need for (at least one form of) the classical limit. And this is a purely classical sub-algebra, because the average position and average momentum can both be well defined. Moreover, note that this is not dissimilar to what happens in classical mechanics when, for example, we study planetary motion using only the position and momentum of the center of mass. The algebra of statements for the position and momentum of the center of mass can be independently described and studied. In this regard, the logical structure of the Borel algebra of quantum mechanics works in the same way.

\subsection{The Non-Distributive Lattice of Classical Mechanics}

Now that we have argued that quantum mechanics can be given an algebra that looks more like the one of classical mechanics, let us show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic. Given that quantum mechanics is inherently a probabilistic theory, let us compare it to a classical probabilistic theory. Thus, let us consider the space of all possible probability distributions $\rho(x,p)$ over phase-space. As already stated, we can formalize this as a Hilbert space, where a vector $\psi$ corresponds to $\sqrt{\rho(x,p)}$. This will be a real vector space, and we are only going to consider operators that can be expressed as a multiplicative function $f(x,p)$. Note that $\langle \psi | F \psi \rangle = \int f(x,p) \rho(x,p) dx dp$ as one would expect.\footnote{To be clear, we are not arguing that this construction is the most appropriate or useful. Simply that it can be done.}  We can now take the algebra of subspaces and, as in the quantum case, it will be non-distributive: the disjunction of two subspaces will not be the set union, but the span.

To really understand the source of the non-distributivity, let us study the simplest possible example. Let us assume that there are only two possible classical states of a certain physical system $s$, $1$ and $2$. A state is identified by a two-valued distribution $[\rho_1, \rho_2]$, where $\rho_1$ is the probability to find the object in state 1 and $\rho_2$ is the probability to find the object in state 2. We note $[k,0]$ the subspace spanned by the first state only. This corresponds to the statement ``the system $s$ is certainly in state $1$'' since the probability for state $2$ is zero. Conversely, $[0,k]$ corresponds to the statement ``the system $s$ is certainly in state $2$''. The disjunction $[k,0] \vee [0,k]$ will correspond to the subspace spanned by both, which is the whole space $[k,j]$ where we can assign different non-zero probability to either case. We can also take the subspace $[k,k]$, the subspace spanned by vectors that have equal components in $1$ and $2$, which corresponds to the statement ``the system is equally likely to be in either state $1$ or $2$''. Note that $[k,0]$, $[0,k]$ and $[k,k]$ are all pairwise disjoint: no pairs have a vector in common apart from zero. Therefore their conjunctions $[k,0] \wedge [0,k] = [0,k] \wedge [k,k] = [k,k] \wedge [k,0] = [0,0]$ are zero dimensional. On the other hand, $[k,k]$ is a subspace of the whole space $[k,j]$ and therefore $[k,k] \wedge [k,j] = [k,k]$. Putting it all together, we have $[k,k] \wedge ( [k,0] \vee [0,k] ) = [k,k] \wedge [k,j] = [k,k]$. But we also have $( [k,k] \wedge [k,0] ) \vee ( [k,k] \wedge [0,k] ) = [0,0] \vee [0,0] = [0,0]$. Therefore the lattice is not distributive.

Note that during the discussion there was no mention as to whether $k$ and $j$ where real numbers, complex numbers or quaternions. The argument is formally the same in all cases. Therefore the lack of distributivity has nothing to do with the complexity of the Hilbert space or non-commuting observables. It's a property of the lattice of subspaces of any vector space and, more generally, of the lattice of subgroups of any group. Indeed, as a further example, consider the group of boost transformations in relativity. The boosts along one direction form a subgroup. If we take the subgroups of boost along $x$ and $y$ respectively, the smallest subgroup that contain both, the disjunction, is the subgroup of all boosts along any direction in the $x$-$y$ plane. As before, the disjunction is the span, so the resulting set is different from the union, and the union of all boosts along the x-axis and the y-axis is not a subgroup and therefore not contained in the lattice. Thus, this has the same non-distributive structure as quantum logic.

We want to be clear that we are not claiming that the lattice of subspaces of quantum logic is not of interest, or that it is of equal interest in both classical and quantum mechanics. Here we have pointed out that the argument that quantum mechanics \emph{requires} necessarily a radical departure from classical logic is untenable, given that both theories are built on similar structures. We conclude that such claims should be substantially weakened, in the light of the arguments that we proposed in this section.

Finally, let us conclude by saying that once we understand the similarity between the theories, we are in a much better position to understand their differences, which naturally exist. For example, quantum mechanics can only be expressed as a probabilistic theory and therefore \emph{always} requires a structure that in classical mechanics is not mandatory. In quantum mechanics, all subspaces of dimension one represent a pure state while in classical mechanics this is not the case. If we take two intervals $U_x$ and $U_p$ in position and momentum, and form the statement ``the state is wholly in $U_x$ and $U_p$'', meaning that the vector components are zero outside of those regions, in the classical case we always find at least one $\rho(x,p)$ that satisfies the bounds, while in quantum mechanics, if the bounds violate the uncertainty principle, no suitable $\psi(x)$ can be found. The point is that the differences are in the content of the logical structure, on the particular relationships among the particular statements, not on the type of structure.

\section{Discussion and Concluding Remarks}
\label{conc}
\vspace{5mm}

\textbf{Acknowledgements:} withheld from draft.
\clearpage

\bibliographystyle{apalike}
\bibliography{bibliography}
\end{document}