### 2023-01-21 Notes/2023-QuantumGeneralizedCoordinates/QuantumGeneralizedCoordinates.tex
Notes on variable changes in QM

\documentclass[11pt]{article}

\usepackage[margin=1.25in]{geometry}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{enumitem}

\def\>{\rangle}
\def\<{\langle}

\begin{document}

\title{Notes on QM in generalized coordinates}
\author{Gabriele Carcassi}

\date{\today}

\maketitle

\begin{abstract}
\end{abstract}

\section{Change of coordinates in manifolds}

\begin{equation}
	\begin{aligned}
		\hat{x}^i &= \hat{x}^i(x^j) \\
		d\hat{x}^i &= \partial_j \hat{x}^i dx^j \\
		d\hat{x}^n &= | \partial_j \hat{x}^i | dx^n = |J| dx^n \\
		|J| \hat{\rho}(\hat{x}^i) &= | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i)  = \rho(x^j) \\
		\int \rho(x^j) dx^n &= \int | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i) dx^n = \int \hat{\rho}(\hat{x}^i) d\hat{x}^n
	\end{aligned}
\end{equation}

Mathematically, $\rho$ can be understood as the (only) component of a volume form. In that case, $\rho$ changes sign under change of orientation. Alternatively, $\rho$ can be understood as the (only) component of a density.\footnote{Effectively, the density is like a volume form that keeps track of orientation.} In this case, $\rho$ does not change sign under change of orientation and the transformation rule should use the absolute value of the Jacobian determinant.

\section{Change of coordinates for wave-function}
\begin{equation}
	\begin{aligned}
		\rho(x^j) &= \psi^\dagger(x^j) \psi(x^j)\\
		||J|| \hat{\rho}(\hat{x}^i) &= || \partial_j \hat{x}^i || \hat{\psi}^\dagger(\hat{x}^i) \hat{\psi}(\hat{x}^i) = \psi^\dagger(x^j) \psi(x^j) = \rho(x^j) \\
		\sqrt{|| \partial_j \hat{x}^i ||} \hat{\psi}(\hat{x}^i) &= \psi(x^j) \\
		\hat{\psi}(\hat{x}^i) &= \sqrt{|| \partial_i x^j ||} \psi(x^j) \\
	\end{aligned}
\end{equation}
Note that $\rho$ can't be negative. Mathematically,  $\rho$ should be the component of a density. Therefore $||J||$ represents the absolute value of the Jacobian determinant.

\section{Change of coordinates in symplectic manifolds}
\begin{equation}
	\begin{aligned}
		\hat{p}_i &= \partial_i x^j p_j \\
		d \hat{p}_i &= \partial_i x^j d p_j \\
		d\hat{p}^n &= | \partial_i x^i | dp^n = |J^{-1}| dx^n \\
	\end{aligned}
\end{equation}

\section{Properties of Fourier transform}

Definition of Fourier transform
\begin{equation}
	\begin{aligned}
		\phi(p_j) &= F(\psi(x^i)) = \frac{1}{\sqrt{2 \pi \hbar}} \int e^{\frac{p_j x^j}{\imath \hbar}} \psi(x^j) dx^n \\
	\end{aligned}
\end{equation}

Definition of convolution
\begin{equation}
	\begin{aligned}
		(g * h)(t) &= \int_{0}^{t} g(\tau) h(t-\tau) d\tau\\
	\end{aligned}
\end{equation}

Convolution theorems
\begin{equation}
	\begin{aligned}
		F((g * h)(t)) &= F(g(t)) F(h(t))\\
		F((g h)(t)) &= F(g(t)) * F(h(t))\\
	\end{aligned}
\end{equation}
TODO: find explicit confirmation for complex functions

\section{Change of Fourier transform}
\begin{equation}
	\begin{aligned}
		F(\hat{\psi}(\hat{x}^i)) &= F\left(\sqrt{|| \partial_i x^j ||} \psi(x^j)\right) = F\left(\sqrt{|| \partial_i x^j ||}\right) * F \left( \psi(x^j)\right)  \\
	\end{aligned}
\end{equation}
TODO: think of cross-checks

\end{document}