### 2023-02-10 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
HarriganSpekkens: another pass

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{tikz}

\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\frenchspacing

\newcommand\partitle[1]{\textsc{#1}.}

\begin{document}

\title{On the reality of the quantum state once again: A no-go theorem for $\psi$-ontic models}
\author{Gabriele Carcassi}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
\author{Andrea Oldofredi}
\affiliation{Centre of Philosophy, University of Lisbon, Portugal}
\author{Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
\vspace{2mm}

\date{\today}

\begin{abstract}
In 2010 Harrigan and Spekkens (HS) proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic. Building on the HS framework Pusey, Barrett and Rudolph (PBR) proved a theorem claiming that $\psi$-epistemic models contradict quantum mechanics, and thus quantum states cannot be taken to represent mere information about quantum systems. In this essay we propose a simple no-go theorem that explains why also $\psi$-ontic models cannot be expected to reproduce quantum mechanics. We will see that, using information theoretic considerations, the lack of overlap of epistemic states requires all states to be orthogonal, which openly contradicts quantum theory. Combining our result with the PBR argument, our conclusion is that both $\psi$-epistemic and $\psi$-ontic models contradict quantum mechanics.
\end{abstract}

\maketitle

\section{Introduction}

In 2010 Harrigan and Spekkens (HS) proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic \cite{Harrigan:2010}. While their original aim was to clarify Einstein's view of Quantum Mechanics (QM), the HS framework has been widely employed in the literature not only to categorize different formulations of QM, but also to argue what types of interpretations are admissible (\cite{Leifer:2013, Leifer:2017, Branciard:2014, Hermens:2021, Wood:2015, Ringbauer:2015, Mazurek:2016, Bartlett:2012}; cf.\ \cite{Oldofredi:2020b, Ladyman:2021} for critical discussions).

Referring to this, one of the most influential works based on HS classification is due to Pusey, Barrett and Rudolph who published a formal result in \emph{Nature Physics}---widely known as the PBR theorem---showing that ``if the quantum state merely represents information about the real physical state of a system, then experimental predictions are obtained that contradict those of quantum theory'' (\cite{PBR:2012}, p.\ 475). Alternatively stated, PBR argued that in every model reproducing the statistics and predictions of QM the quantum state $\psi$ must represent real physical properties of the system under consideration and not agents' knowledge---i.e.\ models must be $\psi$-ontic. Consequently, quantum theories cannot be $\psi$-epistemic.

Such a theorem had a remarkable resonance (\cite{Leifer:2014, Leifer:2014b, Lewis:2012, Renner:2012, Colbeck:2017, Hardy:2013, Maroney:2014, Patra:2013, Mansfield:2016, Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014, Aaronson:2013}), and questions about its actual meaning are still discussed today: on the one hand, it rules out interpretations of QM where $\psi$ merely represents information. On the other, scholars recently showed that non-trivial epistemic as well as statistical approaches to QM are not refuted by the PBR argument (\cite{Ben:2017, Rizzi:2018, Oldofredi:2021, DeBrota:2019}).

While the discussion has been focused mainly on whether $\psi$-epistemic models are problematic or whether the PBR argument itself is problematic, we ask a different question: what if the HS classification itself is problematic? In particular, the framework assumes that once a full description of the transition probability is given for pure states, mixed states are constructed simply as a classical statistical mixture. But classical mixtures and quantum mixtures have different properties. For example, classical mixtures have a single decomposition in terms of pure states while quantum mixtures do not. Studying the behavior of states under mixing, then, opens a new venue of possible violations that are not as easily seen by looking at transition probability.

The core insight is that the degree of overlap between two probability distribution affects how the entropy of their mixture relates to the entropy of the components. Consequently, we find that the lack of overlap required by $\psi$-ontic models forces all pure states to be orthogonal. Since this openly contradicts quantum theory, we conclude that $\psi$-ontic models are not viable.

Combining our result with the PBR argument, we conclude that the HS classification itself is fundamentally flawed: both $\psi$-epistemic and $\psi$-ontic models do contradict quantum mechanics.\footnote{In principle, PBR leaves open the possibility of $\psi$-epistemic models that violate statistical independence. Given that statistical independence is linked to entropy additivity, that loophole can be closed as well.} Consequently, HS framework should be employed neither to classify interpretations of the quantum formalism, nor to draw conclusion about the nature of $\psi$.

\section{Summary of the Harrigan \& Spekkens Model}

We briefly review the main features of the classification provided by HS starting with the usual operational setting for QM. We have a preparation protocol $P$, associated with a density operator $\rho$ on the relevant Hilbert space, and a measurement protocol $M$, represented by a POVM $\{ E_k\}$ where each $k$ represents a possible measurement outcome. The probability of obtaining a particular $k$ given a particular preparation $P$ and measurement $M$ is given by the generalized Born rule
\begin{equation}
	p(k|M, P)=\textrm{tr}(\rho E_k).
\end{equation}

An ontological model, as defined by HS, additionally assumes that there exists a set of states $\Lambda$, called \textbf{ontic states}, that provide the complete specification of the properties of a given physical object. A preparation $P$ will prepare a particular ontic state $\lambda$ according to the probability distribution $p(\lambda | P)$, which is referred to as \textbf{epistemic state}. The measurement outcome will depend only on the ontic state with probability $p(k|\lambda, M)$ (\cite{Harrigan:2010}, p.\ 128). This leads to the following expression:
\begin{equation}
	\int_\Lambda d\lambda p(k|\lambda, M) p(\lambda| P)= \textrm{tr}(\rho E_k).
\end{equation}
It is also assumed that a mixture of pure states $\{ \psi_i \}$ with probabilities $\{ w_i \}$ will be represented by
\begin{equation}\label{epistemic_mixing}
	\sum_i  w_i p(\lambda| P_{\psi_i}).
\end{equation}
As the epistemic states for mixed preparations are simply linear combinations of epistemic states for pure preparations, the HS model concentrates on the latter.

To classify an ontological model, we look at the relationship between quantum and ontic states. There are two broad categories. In a \textbf{$\psi$-ontic} model the wave function is a physical property of the ontic state, in the sense that given an ontic state $\lambda$ there is only one pure state preparation $P_\psi$ that could have prepared $\lambda$. As we can see in figure \ref{overlap} (a), this happens if the probability distributions do not overlap, i.e.\ if we have
\begin{equation}\label{ontic_condition}
	p(\lambda | P_{\psi})p(\lambda|P_{\phi})=0
\end{equation}
for all pairs of states $\psi$ and $\phi$. If a model is not $\psi$-ontic, then it is \textbf{$\psi$-epistemic}. In this case, $\lambda$ can be described by more than one quantum state and the wave function is taken to represent knowledge about the state preparation---in such models quantum states generate overlapping probability distributions over $\Lambda$ as shown in figure \ref{overlap} (b). A $\psi$-ontic model is said \textbf{$\psi$-complete} if the quantum states and the ontic states coincide. More precisely if 
\begin{equation}\label{complete_condition}
	p(\lambda|P_\psi)=\delta(\lambda-\lambda_{\psi}).
\end{equation}
All other models are \textbf{$\psi$-incomplete}.

To make this classification more concrete, let us give an example from classical mechanics. Consider the case where we prepare a particle according to a specific value of energy. The energy partitions phase space into mutually exclusive regions, and therefore we can understand the energy of the preparation as a property of the particle itself. According to HS, this would be an \emph{ontic property}. Consider now the case where we prepare a particle according to a specific temperature. When we take the particle from the oven, we are sampling from a Boltzmann distribution over different energies. However, unlike energy, temperature does not partition phase space because the same particle state could have been prepared by ovens at different temperature. Temperature is a property of the preparation and therefore an \emph{epistemic property}.

\begin{figure}
\includegraphics[scale=.7]{ontic}
\caption{\footnotesize{Harrigan and Spekkens' distinction between $\psi$-ontic (a) and $\psi$-epistemic (b) ontological models.}}\label{overlap}
\end{figure}

\section{Entropy and the ontological model}

The exposition of the HS classification makes it seem one has complete freedom in choosing the overlap between the overlap between epistemic states. In this section we will show that this is not the case: this is constrained by entropy of mixtures. These constraints cannot be replicated by a $\psi$-ontic model with standard information theoretic techniques. Since quantum statistical mechanics and quantum information theory depend on the correct value of entropy on mixtures, $\psi$-ontic models will violate quantum theory.

To show the violation, we will first study how the entropy behaves when mixing two states, both for the Shannon/Gibbs entropy and the von Neumann entropy.\footnote{Full calculations with all steps are included in the appendix.} The first will be used to calculate entropy of statistical mixtures of epistemic state, since equation \ref{epistemic_mixing} tells us the mixture is done according to a standard (classical) measure of probability. The second will be used to calculate the entropy on the quantum case.

\textbf{Information entropy of mixed non-overlapping probability distributions}. Let $\rho_1$ and $\rho_2$ be two probability density function over a space $\Lambda$ with measure $\lambda$. That is, they are the Radon-Nikodym derivative of the respective probability measures and the measure $\lambda$. The entropy for each distribution is given by the usual formula\footnote{The logarithm is assumed to be in base 2.}, for example
\begin{equation}\label{shannon_entropy}
	H(\rho_1) = - \int_\Lambda \rho_1(\lambda) \log \rho_1(\lambda) d\lambda.
\end{equation}
Suppose the two distributions are disjoint, and let $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ be a uniform mixture of the two distributions. More precisely, $\rho$ is the Radon-Nikodym derivative of the average of the two probability measures with respect to $\lambda$. The entropy of $\rho$ is given by
\begin{equation}\label{entropy_nonoverlap}
	H(\rho) = 1 + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2).
\end{equation}
Note how the non-overlapping assumption fixes the entropy of the mixed state.

\textbf{Quantum information entropy of quantum mixed states}. Now suppose $\psi$ and $\phi$ are two pure quantum states and let $p = | \langle \psi | \phi \rangle |^2$ be the probability of transition from one to the other. Consider the mixed state $\rho = \frac{1}{2} | \psi \rangle \langle \psi | + \frac{1}{2} | \phi \rangle \langle \phi |$. Its entropy is given by
\begin{equation}\label{entropy_mixed}
	H(\rho) = H\left(\frac{1+\sqrt{p}}{2}, \frac{1-\sqrt{p}}{2}\right).
\end{equation}
where the right side of the equation is the Shannon entropy calculated on the given values. Therefore the entropy of an equal mixture of two pure states depends only on the probability of measuring one having prepared the other.

\textbf{Theorem: Since non-overlapping distributions can only represent orthogonal states, $\psi$-ontic models cannot be consistent with quantum theory.}

\emph{Proof}: Suppose we have a $\psi$-ontic model defined according to \cite{Harrigan:2010}. The epistemic states $p(\lambda|P_\psi)$ and $p(\lambda|P_\phi)$ consist of non-overlapping probability distributions over a space $\Lambda$. That is, they are the two probability density functions defined to be the Radon-Nikodym derivatives between the respective probability measures over $\Lambda$ and the measure $\lambda$. Equation \ref{epistemic_mixing} tells us that the probability measure combine in the usual way, and therefore an equal mixture of the two states must obey \ref{entropy_nonoverlap}. Given that $\psi$ and $\phi$ are pure states, and the entropy for pure states in quantum mechanics is zero, we must have
\begin{equation}\label{entropy_pure}
	H(p(\lambda|P_\psi)) = H(p(\lambda|P_\phi)) = 0,
\end{equation}
and therefore
\begin{equation}\label{required_entropy}
	\begin{aligned}
		H\left(p(\lambda|\frac{1}{2}P_\psi + \frac{1}{2}P_\phi)\right) &= \\
		H\left(\frac{1}{2}p(\lambda|P_\psi) + \frac{1}{2}p(\lambda|P_\phi)\right) 
		&= 1.
	\end{aligned}
\end{equation}
If we compare the above with eq.~\ref{entropy_mixed}, it follows that $p$ must be zero. In fact, the only case the Shannon entropy for a two element distribution is equal to one is for the uniform distribution $H(1/2, 1/2) = 1$.  That is, we must have that
\begin{equation}\label{orthogonal}
	\langle \psi | \phi \rangle = 0
\end{equation}
no matter what $\psi$ and $\phi$ are.

Hence, the non-overlapping assumption built into the $\psi$-ontic model necessarily implies that all pure states are orthogonal. Since this is not true in quantum mechanics, any $\psi$-ontic model will fail to reproduce the results of quantum information, quantum statistical mechanics and, therefore, quantum theory in general. $\square$

If we combine this result with PBR, the only models that are not ruled our are $\psi$-epistemic models where the objects are not independently prepared. That is, where the joint distribution over ontic states is not the product of the marginals. However, this loophole can also been closed with information theoretic considerations. It is a well-known result in both classical and quantum information theory that the entropy of a joint distribution is the sum of the marginal if and only if the subsystems are independent (\cite{Ash:2010}, \cite{Nielsen:2010}). Therefore an epistemic state that represents independent quantum states must also be the joint distribution of statistically independent epistemic states of the individual systems. This tells us that $\psi$-epistemic models must represent composite states of independently prepared systems with independently distributed epistemic states, and therefore PBR applies. This closes the last loophole: no model in the HS categorization can reproduce quantum theory in its entirety, which includes quantum information and quantum statistical mechanics.

Thus, one concludes the following:
\begin{quote}
	\textbf{no ontological model can reproduce quantum statistical mechanics and quantum information with standard information theoretic tools}.
\end{quote}
That is, even if one had an ontological model that mapped the result of single outcome correctly, it would still give the wrong answer when trying to calculate ensembles and their entropy.

\section{Throwing out information theory}

Note that our conclusion relies on the use of standard information theoretic tools. In principle, we do not exclude that someone may find a different construction to reproduce the correct entropies. That is, one may give a different definition of entropy to use on the ontic space. We believe this to be highly unlikely as there are, at least, two objective difficulties that cannot be circumvented.

First of all, information entropy is strictly concave, meaning that $H(p \rho_1 + (1-p) \rho_2) \geq p H(\rho_1) + (1-p) H(\rho_2)$ with the equality valid only if $\rho_1$ and $\rho_2$ are the same exact state. That is, we can check whether two states $\rho_1$ and $\rho_2$ are the same state simply by creating a mixture $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ and see whether we end up with the same entropy or not. In other words: \emph{entropy fixes equality}. %spiega meglio
The issue here is that the mapping between epistemic states and density matrices is not injective. Quantum mixtures, in fact, do not have a single decomposition in pure states. For example, the maximally mixed state of a qubit can be achieved with either an equal mixture of spin up and down or an equal mixture of spin left and right. Therefore the same quantum mixture must correspond to an equivalence class of epistemic states, which means whatever definition we have on entropy for epistemic states must not satisfy strict concavity to be able to agree with the quantum case.\footnote{The degree of degeneracy one has to require is also extreme. For a qubit, in the simplest case, we would have a $\psi$-ontic state for each direction of spin, meaning the space of ontic states is the sphere $S^2$. The space of all possible epistemic states is the set of all possible distribution integrable to one. This is an infinite dimensional function space. However, the space of mixed state for a quibit is the interior of a sphere, which is a bounded three-dimensional manifold. The equivalence classes of epistic states, then, must be infinite dimensional.} That is, one can make mixtures of different ensemble without increasing entropy, which does not make physical sense.

Secondly, as equation \ref{entropy_mixed} shows, computing the entropy requires the inner product. In fact, if we had the entropy for all even mixtures of all pairs of pure states, we could reconstruct the inner product. Therefore, whatever other structure one puts on top of epistemic states, essentially redefines the inner product. Also, it has to do it in a way compatible with the other definitions of the model, therefore this structure will likely be ad-hoc for each model. In other words, the idea that the inner product just represents transition probabilities, which, in our view, is the key idea the ontological model uses to explain ``what really happens'' does not work.

One, naturally, will object that we haven't strictly proven that any of this is impossible or that there is no reasonable way to do it. Granted. However, it is not up to us to show that HS models cannot be made to work outside of the scope of standard information theory. And in any case, we believe we put forth enough evidence that the HS framework has serious flaws which, at the very least, must be understood by those that use it to draw any conclusion about the interpretation of quantum theory and the nature of $\psi$.

\section{Conclusion}

By using standard information theoretic techniques, we have shown that $\psi$-ontic models are not compatible with quantum information theory, and therefore with quantum mechanics itself. Given that the PBR theorem already ruled out $\psi$-epistemic models, the HS categorization of ontological models is fundamentally empty. Thus, such a framework should be employed neither to classify quantum interpretations, nor to draw conclusions on the nature of the quantum state.

The key problem of the HS categorization is that it is ``not quantum enough''. The framework tries to hide the non-classicality at the level of each ontic state, which, by construction, cannot be decomposed further using a standard probability measure. However, it posits that statistical mixtures are simply classical mixture in the sense that they are captured by a standard probability measure. But quantum mixtures are non-classical in the same way that the quantum transition probabilities defined by the inner product are non-classical. Therefore the HS categorization prevents one to describe quantum statistical mixing.

Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are not sustainable not because there is a problem with $\psi$-epistemic models in particular, but rather there is an underlying problem in the HS definitions. If understood in this manner, the PBR result is the first part of a more general argument showing the inadequacy of the HS framework to categorize quantum theories. Argument that we have completed with the proposed no-go theorem against $\psi$-ontic models.

\section{Acknowledgements}

Andrea Oldofredi is grateful to the Funda\c{c}$\tilde{\mathrm{a}}$o para a Ci\^encia e a Tecnologia (FCT) for financial support (Grant no. 2020.02858.CEECIND).  This work is in connection to Assumptions of Physics, a larger project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).

%\bibliographystyle{ieeetr}
\bibliography{bibliography}
\clearpage

\section*{Appendix: calculation}
\label{A}
\textbf{Entropy of mixed non-overlapping distributions}. We want to show that, given two non-overlapping probability distributions $\rho_1$ and $\rho_2$, the entropy of $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ is given by
\begin{equation}
	H(\rho) = 1 + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2). \tag{\ref{entropy_nonoverlap}}
\end{equation}

Let $U_1, U_2 \subset \Lambda$ be the respective supports of the distributions, which are non-overlapping i.e.\ $U_1 \cap U_2 = \emptyset$. We then have
\begin{align*}
	H(\rho) &= - \int_\Lambda \rho \log \rho d\lambda \\
	&= -\int_{U_1} \rho \log \rho d\lambda -\int_{U_2} \rho \log \rho d\lambda \\
	&= -\int_{U_1} \frac{1}{2} \rho_1 \log \frac{1}{2} \rho_1 d\lambda -\int_{U_2} \frac{1}{2} \rho_2 \log \frac{1}{2} \rho_2 d\lambda \\
	&= - \frac{1}{2} \int_{U_1} \rho_1 \log \frac{1}{2} d\lambda - \frac{1}{2} \int_{U_1} \rho_1 \log \rho_1 d\lambda \\
	&- \frac{1}{2} \int_{U_2} \rho_2 \log \frac{1}{2} d\lambda - \frac{1}{2} \int_{U_2} \rho_2 \log \rho_2 d\lambda \\
	&= - \frac{1}{2} \log \frac{1}{2} - \frac{1}{2} \log \frac{1}{2} + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2) \\
	&= 1 + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2). \\
\end{align*}

\textbf{Entropy of quantum mixed states}. We want to show that, given two states $\psi$ and $\phi$, the entropy of the mixed state $\rho = \frac{1}{2}|\psi\rangle\langle\psi| + \frac{1}{2}|\phi\rangle\langle\phi|$ is
\begin{equation}\label{entropy}
	H(\rho) = H\left(\frac{1+|\langle\psi|\phi\rangle|}{2}, \frac{1-|\langle\psi|\phi\rangle|}{2}\right).
\end{equation}

\begin{center}
	\begin{tikzpicture}[scale = 1]
		\draw (0,0) circle (2);
		\node at (-2.3,0) {$-$};
		\node at (2.3,0) {$+$};
		\node at (2,1.4) {$\psi$};
		\node at (2,-1.4) {$\phi$};
		\draw (-2,0) -- (2,0);
		\begin{scope}
			\clip(0,0) circle (2);
			\draw (0,0) -- (2,1.4);
			\draw (0,0) -- (2,-1.4);
			\draw (1.634,1.4) -- (1.634,-1.4);
		\end{scope}
		\fill (1.634,0) circle (0.05);
		\node at (1.8,.2) {$\rho$};
	\end{tikzpicture}
\end{center}

Note that $\psi$ and $\phi$ will identify a two-dimensional subspace which can be thought, without loss of generality, as a qubit and therefore can be represented by a Bloch sphere. The picture represents the intersection of the Bloch sphere with the plane identified by $\psi$ and $\phi$. As $\rho$ is an equal mixture of the two states, it will be represented by the midpoint between the two. Taking the line that goes through $\rho$ and the center of the sphere, we can see that $\rho$ can also be seen as the mixture of the states $+$ and $-$ which, since they represent equal and opposite directions, form a basis. To diagonalize $\rho$, then, means to express it in terms of $+$ and $-$.

If $\theta_{\psi\phi}$ is the angle between $\psi$ and $\phi$, we have
\begin{equation}
	|\langle \psi | \phi \rangle |^2 = \cos^2 \frac{\theta_{\psi\phi}}{2}.
\end{equation}
The angle is divided by two because the angle on the Bloch sphere (i.e. in physical space) is double the angle in the Hilbert space. For example, for $z^+$ and $z^-$ the angle on the Bloch sphere would be $\pi$ and the inner product is zero (i.e. opposite directions in physical space correspond to orthogonal states).

Now we express $\psi$ and $\phi$ in terms of $+$ and $-$, remembering that they form a basis. Given that $\rho$ is at the midpoint, the figure is vertically symmetric. The angle between $\psi$ and $+$, then, is half of $\theta_{\psi\phi}$. The inner product between $\psi$ and $+$ is
\begin{equation}
	\begin{aligned}
		|\langle \psi | + \rangle |^2 &= \cos^2 \frac{\theta_{\psi +}}{2} \\
		&= \cos^2 \frac{\theta_{\psi\phi}}{4}.
	\end{aligned}
\end{equation}
Keeping in mind that we are composing vectors in the Hilbert space (and not in the geometry of the physical space) we have
\begin{align*}
	\left|\psi\right>&=\cos\frac{\theta_{\psi\phi}}{4}\left|+\right>+\sin\frac{\theta_{\psi\phi}}{4}\left|-\right> \\
	\left|\phi\right>&=\cos\frac{\theta_{\psi\phi}}{4}\left|+\right>-\sin\frac{\theta_{\psi\phi}}{4}\left|-\right>.
\end{align*}

The density matrices corresponding to the pure states are
\begin{align*}
	\left|\psi\right>\left<\psi\right|&=\cos^2\frac{\theta_{\psi\phi}}{4}\left|+\right>\left<+\right|\\
	&+\cos\frac{\theta_{\psi\phi}}{4}\sin\frac{\theta_{\psi\phi}}{4}\left(\left|+\right>\left<-\right|+\left|-\right>\left<+\right|\right) \\
	&+\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right| \\
	\left|\phi\right>\left<\phi\right|&=\cos^2\frac{\theta_{\psi\phi}}{4}\left|+\right>\left<+\right|\\
	&-\cos\frac{\theta_{\psi\phi}}{4}\sin\frac{\theta_{\psi\phi}}{4}\left(\left|+\right>\left<-\right|+\left|-\right>\left<+\right|\right) \\
	&+\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right|.
\end{align*}

We can now calculate the mixture
\begin{align*}
	\frac{1}{2}(|\psi\rangle\langle\psi| &+ |\phi\rangle\langle\phi|) \\
	&=\cos^2\frac{\theta_{\psi\phi}}{4}\left|+\right>\left<+\right| +\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right| \\
	&=\frac{1+\cos\frac{\theta_{\psi\phi}}{2}}{2}\left|+\right>\left<+\right| +\frac{1-\cos\frac{\theta_{\psi\phi}}{2}}{2}\left|-\right>\left<-\right| \\
	&=\frac{1+|\langle\psi|\phi\rangle|}{2}\left|+\right>\left<+\right| +\frac{1-|\langle\psi|\phi\rangle|}{2}\left|-\right>\left<-\right|. \\
\end{align*}

As $\rho$ is in a diagonal form, the entropy is given by \ref{entropy}.

\end{document}