### 2025-09-22 Proposals/2025-NSF-EntropicGeometry/ProjectSummary.tex
NSF proposal

\thispagestyle{empty}
\noindent
\textbf{Overview:}\\

\noindent
\textbf{Intellectual Merit:}  \\

\noindent
\textbf{Broader impacts:}  \\

The proposed activities additionally include significant outreach through social media, primarily through YouTube.  Videos will be produced to disseminate the results of the research, with some videos aimed at the level of an advanced undergraduate in mathematics and/or physics, furthering education.  Other videos will be aimed at active researchers in various fields of mathematics and physics, providing a new mathematical framework as well as a new perspective linking traditionally disparate realms of physics that may impact tools or thinking in their own research.  In the spirit of an open-research approach that moreover leverages ``crowd'' contributions, YouTube videos aimed at researchers will additionally be used to advertise open problems that the project personnel may not be able to expediently resolve themselves. Further dissemination of the open-research approach may inspire researchers in completely different areas to adopt a similar approach for their own projects, with potential acceleration of research progress in particular within interdisciplinary areas, where not all necessary expertise may be available within a single research community.