### 2025-11-08 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Various further updates

\begin{center}
	\textbf{An axiomatic basis for entropic structures in physical theories} \\
	Christine A. Aidala (PI) and Alejandro Uribe (Co-PI), University of Michigan
\end{center}
%\vspace{-8mm}

\section{Introduction}

\iffalse
Physics is currently a patchwork of different theories with no common underlying foundation. As part of our larger Assumptions of Physics program, we aim to develop an axiomatic basis for states and processes that can be equally applied to different theories, such as classical and quantum mechanics. Different theories are then recovered as specializations of the more general structure.

While the state spaces of classical (i.e.~symplectic manifolds) and quantum mechanics (i.e.~projective Hilbert spaces) are rather different, the space of ensembles of both have a similar structure. They are both convex subsets of a topological vector space, equipped with a strictly concave entropy function. The core idea, then, is to use ensembles as a way to generalize the notion of physical states, and the entropy to characterize the geometry in a way that is independent of the particulars of each physical theory. In the past year, we have done significant preliminary work that shows that not only is this approach possible, but it is extremely productive. The proposed work is to turn those preliminary insights in an axiomatic specification that is fully physically motivated. If successful, such an approach will provide the following benefits:
\begin{enumerate}
	\item clarify exactly what physical premises guarantee the basic geometric structures of physical theories, allowing a more principled formulation of current physical theories as well as the search for new ones;
	\item provide a generalized structure for information theory, possibly allowing future generalization of results from both classical and quantum information theory
	\item provide useful constraints to more precisely formulate physical theories in the infinite-dimensional case.
\end{enumerate}
\fi

The proposed work starts from the observation that the spaces of statistical ensembles of classical and quantum mechanics have multiple features in common. They are both convex subsets of a locally convex topological vector space whose metric is induced by the information entropy, or more specifically, by the Jensen-Shannon divergence. This is in contrast to the different structures of the spaces of classical and quantum states, which rely on symplectic manifolds and projective spaces of complex Hilbert spaces, respectively. Even more notably, our preliminary work shows that the symplectic form of classical mechanics and the inner product (i.e.~Born rule) of quantum mechanics can indeed be recovered by the entropy. In general, the states defined by mechanics can be understood as special cases of ensembles with the lowest possible entropy or the extreme points of the convex set.

The goal is to find minimal requirements (i.e.~axioms) for an entropic structure that is common to all physical theories in such a way that known results from classical and quantum theories, in areas like mechanics and information theory, can be generalized. In particular, the relationships among the four mathematical structures required by a space of ensembles---the topology, the convex structure, the proposed entropic structure and the Lie algebra---need to be understood. Moreover, it is crucial that these premises are well founded on physical requirements, so that the generalization is of value to physics.

\section{Proposed activities}

\subsection{Intellectual merit}

As a guide to the reader, let us recall how states and ensembles are defined in both classical and quantum mechanics. For a discrete classical system, the state space $X$ is a countable set with the discrete topology, the space of ensembles consists of all probability measures over the Borel algebra of $X$ and the entropy is given by $S(p_i) = - \sum_i p_i \log p_i$. Note that all logs here are assumed base 2. In classical mechanics, instead, the state space $X$ is a symplectic manifold, which is therefore equipped with the  Liouville measure $\mu(U) = \int_U \omega^n$, where $\omega$ is the symplectic form. The space of statistical ensembles is the space of all probability measures over the Borel algebra of $X$ that are absolutely continuous with respect to the Liouville measure. For these measures, the Radon-Nikodym derivative $\rho = \frac{dp}{d\mu}$ exists and the entropy is defined as $S(\rho) = - \int \rho \log \rho d\mu$. In particular, for a uniform distribution with support $U$, the entropy reduces to $S(\rho_U) = \log \mu(U)$. In quantum mechanics, the state space is given by the projective space $P(\mathcal{H})$ of a complex Hilbert space $\mathcal{H}$, the space of ensembles is given by the density operators (i.e.~positive semi-definite, self-adjoint trace-one operators) and the entropy is given by $S(\rho) = -\tr(\rho \log \rho)$.

While the structures of the state spaces of classical and quantum mechanics are different, their spaces of ensembles have a rather similar structure: they both consist of a bounded convex subset of a metrizable and second-countable locally convex topological vector space, where the metric can be derived from the entropy through the square root of the Jensen-Shannon divergence $JSD = S\left(\frac{1}{2} \ens[a] + \frac{1}{2} \ens[b]\right) - \left(\frac{1}{2}S(\ens[a]) + \frac{1}{2}S(\ens[b])\right)$~\cite{Jensen-ShannonOriginal,HayashiQuantumInfo,majtey2005jensen}. The JSD is, in both cases, the square of a distance function and can be interpreted as how much the average entropy increases under an equal mixture of two ensembles. There are, therefore, three common structures at play: a topology (which physically represents experimental verifiability), a convex structure (which physically represents the ability to make statistical mixtures) and an entropic structure (which physically represents the variability of the elements within each ensemble). While the axioms for the topology and the convex structure are readily available, to our knowledge, the correct axioms for the entropic structure still need to be identified. Moreover, the interplay between the topology, the convex structure and the entropic structure opens additional mathematical questions which have not been investigated.

The goal of this proposed work is to develop a common axiomatic basis for the information-theoretic structure underlying all physical theories, classical and quantum. The immediate aim is to generalize results of both classical and quantum information theory to a common, more abstract theory that can also handle the infinite-dimensional case. Longer term, we aim to use the axioms developed here as part of a larger effort to develop a unified base theory for physical theories with a clean mathematical foundation.

Note that in all above three cases the definition of the entropy is tightly bound to the definition of the space. Therefore, the biggest problem is understanding how one can give a single characterization that is general enough to encompass the three cases, while constrained enough to recover the correct expression in each case. Fortunately, in our preliminary work we already have found the solution. In all cases, the entropy possesses an upper bound to its increase under convex combination: $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p) + p S(\ens[a]) + \bar{p} S(\ens[b])$ where $\bar{p} = 1-p$ and $I(p) = -p \log p - \bar{p}\log \bar{p}$. This maximum increase is given precisely when the two ensembles are orthogonal with respect to the inner product of the respective space. In the classical case, this means that the measures have disjoint support; in the quantum case, it means that the eigenspaces of non-zero eigenvalues are orthogonal. In other words, characterizing the interaction of the entropy function with the topology and the convex structure is enough to characterize the entropic structure.  That is, the form of $I(p)$---the Shannon entropy formula~\cite{Shannon}---can be characterized, and since in both classical and quantum mechanics ensembles can always be decomposed into orthogonal components, the different entropy formulas can be recovered in each case.

Another important insight from our preliminary work is that the entropy is not only responsible for the geometry of the ensemble space, but it is also intimately related to the geometry of the state space of both classical and quantum mechanics. That is, the entropic structure of the ensemble space is equivalent to the geometric structure of the state space.

As we saw in classical mechanics, the symplectic form gives us the Liouville measure which in turn is used to define the entropy. Therefore, the entropy of an ensemble is defined from the geometry of the state space. However, note that the entropy for a uniform distribution $\rho_U$ over a set $U$ is given by $S(\rho_U) = \log \mu(U)$. Therefore, if we are given the entropy for all probability measures, we can recover the Liouville measure by setting $\mu(U) = 2^{S(\rho_U)}$. While this does not recover the symplectic form directly, note that the symplectic form can also be used to define the entropy of the marginal distributions along each degree of freedom. Around a point, in fact, we can define Darboux coordinates and define joint probability distributions that are the product of independent marginals along each pair of position and momentum. One finds a similar relationship between the entropy of uniform marginal distributions and the integral of the symplectic form over the support of the marginal. From these, we believe we can recover the symplectic form.

In quantum mechanics, the geometry is given by the Born rule
\begin{equation}
	p(\phi|\psi) = \frac{\< \phi | \psi \>\< \psi | \phi \>}{\< \psi | \psi \>\< \phi | \phi \>}.
\end{equation}
Given two states $\psi$ and $\phi$, the ensemble given by their equal mixture, $\rho_{\psi\phi} = \frac{1}{2}|\psi\>\<\psi|+\frac{1}{2}|\phi\>\<\phi|$,  has von Neumann entropy $S(\rho_{\psi\phi}) = I\left(\frac{1 + \sqrt{p(\phi|\psi)}}{2} \right)$ where $I$, as before, is the Shannon entropy for two elements. As before, the entropy is defined in terms of the inner product. However, the relationship between $S(\rho_{\psi\phi})$ and $p(\phi|\psi)$ is invertible. This means that, given the entropy of the equal mixtures of all pairs of states we can reconstruct the Born rule and, therefore, the geometry of the state space.

Given that \emph{all} physical theories must admit a notion of ensembles that can be statistically mixed via convex combination, these insights from our preliminary work show that the entropic structure plays a crucial role in all physical theories and merits further study. Moreover, while typically information geometry is cast in finite-dimensional spaces, there is nothing preventing its generalization to infinite dimensions. If the entropic structure, in fact, imposes a distance function whose open balls are convex, then this provides us with a locally convex topology which is exactly what is needed to define calculus in infinite dimensions. Moreover, the Lie algebras of both classical and quantum mechanics, Poisson brackets and commutators, respectively, have a special relationship with the geometric and entropic structures, which may provide us with a template to generalize those structures as well.

Given these very promising preliminary findings, the goal of the proposed research is to find a minimal set of axioms to characterize the entropic structure of physical theories. Minimal in the sense that, in the spirit of analogous work on the foundations of mathematics, like set theory or reverse mathematics~\cite{RevMathBook}, the axioms can be shown to be satisfied by the different theories (which now become models of the larger theory) but also allow one to rederive the entropy formula for each specific theory. Let us, then, go through the current state of our preliminary work, and show the type of questions that the overall approach touches.

%The goal is to find a very limited set of axioms to define the entropic structure of a physical theory that can be justified from physical considerations, show that this set of axioms imposes a geometric structure on a generic space of statistical ensembles and that it recovers the standard geometric structures of classical and quantum mechanics. Furthermore, the same structure should be connected to topological and measure theoretic structures used in physics.

%and Other that are required by any physical theory to define its entropic structure, which  and construct a general connection to geometry, calculus and topology, thus giving a generalized precise mathematical characterization of physical theories. In our numerous conversations with mathematicians, physicists and philosophers, it is clear that such work is sorely needed, and, since it lies at the intersection of different fields, it requires an interdisciplinary approach.

%We want to stress that finding the correct starting point is not just a mathematical issue, but also a physics problem. Let us explore this problem through an example. The space $M_1([0,1])$ of probability measures defined over the $[0,1]$ interval is an infinite-dimensional vector space. As Aliprantis and Border note in the context of their own field of economics [add reference],  ``Since there is more than one topology of interest on an infinite dimensional space, the choice of topology is a key modeling decision that can have economic as well as technical consequences.'' Unbeknowst to many physicists, this is true for physics as well. If we choose a weak topology, we are saying the convergence of expectation values of all continuous functions is enough to determine convergence of ensembles. While this may seem plausible at first glance, we can find a weakly convergent sequence for which the entropy does not converge. If the entropic structure of a physical theory is really equivalent to its geometric structure, not only should the entropy be continuous, but it is likely that the topology should be the one generated by the entropic structure. Ultimately, answering these questions means understanding both the physical and mathematical consequences of those decisions.

The central focus is the intuitive notion of statistical ensemble (i.e.~the collection of all outputs of a preparation procedure), as this seems to be particularly suitable for generalization for a number of both practical and conceptual considerations. In terms of the mathematics, as we have seen, the spaces of statistical ensembles of classical and quantum mechanics can be made to look very similar. From the experimental side, statistical ensembles are what we produce and measure experimentally. On the conceptual side, the reproducibility requirement of scientific investigation means that physical laws are of the form ``whenever I prepare a system according to this procedure, I will find this range of outcomes.'' That is, physical laws are relationships between ensembles. Therefore, it makes sense to require that any physical theory must \emph{at least} posit a space of ensembles. This provides us with a strong conceptual starting point, and properties of ensembles are what need to be specified through axioms.

The first axiom establishes that an ensemble space $\Ens$ is a $\mathsf{T}_0$ second-countable topological space. This is based on our previous work~\cite{aop-book,aop-topExpDisting}, which established a link between the experimental verifiability of the theory and its topology. The open sets of the topology correspond to experimentally verifiable statements (i.e.~statements associated with an experimental test that terminates successfully in finite time if and only if the statement is true) while Borel sets correspond to statements associated with an experimental test without termination guarantees~\cite{aop-book,aop-topExpDisting}. Real and integer numbers, for example, are recovered from an idealized model representing references of a reference system. The key insight is that the set-theoretic operations represent a two-valued logic system (i.e.~truth or falsehood of statements) while the topology keeps track of a three-valued logic system (i.e.~interior, exterior and boundary represent respectively successful termination, unsuccessful termination and non-termination). The topology, therefore, plays an important role.

The second axiom imposes a convex structure, which represents the ability to perform statistical mixtures. Given two ensembles $\ens[a], \ens[b] \in \Ens$, we can always choose a probability $p \in [0,1]$ and select $\ens[a]$ $p$ percent of the times and $\ens[b]$ $\bar{p} = 1-p$ percent of the times. This gives the ensemble space a convex structure as $p \ens[a] + \bar{p} \ens[b] \in \Ens$ for all $\ens[a], \ens[b] \in \Ens$ and $p \in [0,1]$. This insight is already used in the literature, particularly in Generalized Probabilistic Theories (GPTs); see e.g.~Refs.~\cite{plavala2023,wilce2025generalizedprobabilitytheorynotes,DArianoBook} \textcolor{blue}{Anything else to cite?}. The main innovation in this respect is that we do not limit ourselves to finite combinations and we do not simply allow all infinite combinations (which can be shown to include physically untenable cases~\cite{baccetti2013infinite,aop-unphysHilbert}). Rather, we let the topology do the work. That is, infinite convex combinations exist only if they are the topological limit of finite convex combinations. On physical grounds, the mixing operation must be continuous in the natural topology defined by experimental verifiability, much like addition and scalar multiplication are continuous in a topological vector space, leading to the idea of a topological convex space.

Surprisingly, topological convex spaces do not seem to have been studied, and this structure alone presents interesting sets of questions.  For example, it is established that a convex space embeds into a vector space if and only if it is cancellative, that is, if $p \ens[a] + \bar{p} \ens = p \ens[b] + \bar{p} \ens$ for some $p \in (0,1)$ implies $\ens[a] = \ens[b]$. The question now is what are necessary and sufficient conditions for a topological convex space to embed continuously into a topological vector space? Can we have, for example, a Hausdorff second-countable topological convex space that is not a continuous embedding of a topological vector space? Topological vector spaces present a number of nice results which may or may not hold in topological convex spaces. For example, any $\mathsf{T}_1$ second-countable topological vector space is metrizable. Does that same result hold for topological convex spaces? Another interesting aspect is the notion of closure. In a topological vector space, the topological closure of a convex set is convex. We found that this is true in topological convex spaces as well. Additionally, there is another type of closure that may be more interesting physically. We can define the $\sigma$-hull of a set $A$ as the set of possible infinite convex combinations, which can also be shown to be a convex set. This will be a superset of the hull of $A$, because it will include all finite convex combinations, and it will be a subset of the closed hull of $A$, as all infinite convex combinations are topological limits. Moreover, it can be shown that, in some cases, the hull, the $\sigma$-hull and the closed hull are all different. From the physics side, the $\sigma$-hull may be a more physically meaningful closure as it represents the true limit process of statistical mixing, the operation the axiom is characterizing. The subtle issue is that convex combinations always return elements ``inside'' the set, while linear combinations can also return the ``edges.'' Therefore, there are already interesting topics to study even without imposing the entropic structure.

The third axiom, central to the proposed work, imposes an entropy function that quantifies the variability of elements within a statistical ensemble. Although we imagine an ensemble to be a collection of instances, we do not formalize these instances, as they need to be left unspecified in a general theory. However, the variability of the elements within an ensemble is physically significant as it will influence, for example, the variability of statistical measurements. As we showed in previous work~\cite{Carcassi:2021}, entropy quantifies the variability of the elements within a distribution, which gives us a physically meaningful starting point. The primary goal of the proposed work, then, is to find what axiomatic properties this variability indicator must have, and to understand their conceptual, physical and mathematical implications.

In our preliminary work, for example, the axiom equips each ensemble space with an entropy function $S : \Ens \to \mathbb{R}$ that currently has the following properties:
\begin{itemize}
	\item Continuity
	\item Strict concavity: $S(p\ens[a] + \bar{p} \ens[b]) \geq p S(\ens[a]) + \bar{p} S(\ens[b])$ with the equality holding if and only if $\ens[a] = \ens[b]$
	\item Upper variability bound: there exists a universal function $I : [0,1] \to \mathbb{R}$ (i.e.~the same for all ensemble spaces) such that $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p) + p S(\ens[a]) + \bar{p} S(\ens[b])$; if the equality holds, $\ens[a]$ and $\ens[b]$ are orthogonal, noted $\ens[a] \ortho \ens[b]$
	\item Mixtures preserve orthogonality: $\ens[a] \ortho \ens[b]$ and $\ens[a] \ortho \ens[c]$ if and only if $\ens[a] \ortho p \ens[b] + \bar{p} \ens[c]$ for any $p \in (0,1)$
\end{itemize}
These properties can be justified by the physics. For example, if we are performing a statistical mixture between two ensembles, the average variability cannot decrease, which justifies the concavity. The goal is to have a full physical justification for each property to make sure that we are not leaving out any possibly interesting physical theory. Similarly to what one does in the foundations of mathematics, we typically construct different models of the various formulations of the axioms to check that each property is truly independent from the other, and that it is physically required.

Though we do not believe the current version of the axiom lists all necessary properties of the entropic structure, as discussed below, it does list enough of them to already prove interesting results. For example, we can show that $I(p)$, the maximum entropy increase during mixing, must be, up to a multiplicative constant, the Shannon entropy~\cite{Shannon} $I(p)= - p \log p - \bar{p} \log \bar{p}$. The proof is general in the sense that it does not assume whether the specific model is classical, quantum or something else entirely. Moreover, in the case of both classical and quantum mechanics, states can be decomposed into orthogonal components, which means the upper variability is an equality and can be used to calculate the entropy from that decomposition. This will allows us to recover the expressions for the entropy in the three different cases, classical discete, classical continuous, and quantum. Similarly, we will be able to show that, from this structure alone, the geometric structure of both classical and quantum mechanics can be recovered.

Our preliminary work also shows that an entropy satisfying those properties forces the convex space to be cancellative, meaning that it embeds into a vector space. That is, a convex space that does not embed into a vector space cannot posses an entropic structure. We can, in fact, produce a model of that case and see exactly why the entropic structure fails. We can also show that an ensemble space must be ``directionally bounded,'' meaning its intersection with any affine line is a finite interval. It is an open question as to whether the ensemble space must be a bounded set. We believe we can also show that some examples of Generalized Probabilistic Theories are inconsistent with an entropic structure and therefore cannot represent a physically meaningful space of ensembles. In other words, the entropic structure constrains the types of spaces possible for physical theories in surprising and interesting ways.

Another line of inquiry is the relationship between the entropic structure and a metric over the ensemble space. For example, we can define the mixing entropy as $MS(\ens[a], \ens[b]) = S\left(\frac{1}{2}\ens[a] + \frac{1}{2} \ens[b]\right) - \left(\frac{1}{2} S(\ens[a]) + \frac{1}{2} S(\ens[b])\right)$. If we express the entropy in bits, the mixing entropy satisfies the following:
\begin{itemize}
	\item non-negativity: $MS(\ens[a], \ens[b]) \geq 0$
	\item identity of indiscernibles: $MS(\ens[a], \ens[b]) = 0 \iff \ens[a]=\ens[b]$
	\item unit boundedness: $MS(\ens[a], \ens[b]) \leq 1$
	\item maximality of orthogonals: $MS(\ens[a], \ens[b]) = 1 \iff \ens[a] \ortho \ens[b]$
	\item symmetry: $MS(\ens[a], \ens[b]) = MS(\ens[b], \ens[a])$
\end{itemize}
It is almost a distance function, as it does not satisfy the triangle inequality, but it recovers the Jensen-Shannon divergence (JSD)~\cite{Jensen-ShannonOriginal} in both classical and quantum mechanics. While the JSD is the square of a distance function in both classical and quantum information theory (see e.g.~\cite{JSDProperties,HayashiQuantumInfo}), our current axioms do not guarantee $MS$ to have this property in general. An interesting question, then, is whether enough physical constraints exist for that to be the case. The $MS$ can be used to create open balls and show that the topology must be Hausdorff. It is not yet clear whether this topology is metrizable and whether it must coincide with the topology of the ensemble space.

A related line of questions relates the entropic structure of a physical theory to elements of differential geometry and the calculus of variations. We are aware that the current axioms do not guarantee the entropic open balls to be convex, even though they are convex in both the classical and quantum cases. We suspect the open balls to be convex in general, even though we have not yet found a physically satisfying premise. If the convexity of the open balls is established, then the topology generated by entropic open balls is locally convex, which gives us enough structure to define calculus in infinite-dimensional spaces. That is, the entropic structure would in fact be responsible for the ability to define derivatives in physical theories.

In the finite-dimensional case, and if the entropy is twice differentiable with respect to the convex coefficients, the strict concavity of the entropy means that the Hessian is negative definite, which allows us to define a metric tensor. Let $V_{\ens[a]}$ be the embedding vector space with $\ens[a] \in \Ens$ as origin. This will be a differentiable manifold. On a linear coordinate chart we can define $g_{\ens}( v, w ) = - \frac{\partial^2 S}{\partial \ens ^2}( v, w )$ where $v$ and $w$ are elements of the tangent space at $\ens \in \Ens \subset V_{\ens[a]}$. Since all Hessians are symmetric, $g_{\ens}$ will be bilinear, symmetric and positive definite: it is a metric tensor and $V_{\ens[a]}$ is a Hessian manifold~\cite{hessianmanifoldsShima1997,hessianmanifoldsliu2025}. In classical and quantum mechanics it recovers the Fisher information metric (see e.g.~Refs.~\cite{facchi2010,AmariInfoGeom}). The immediate goal, then, is to see how much of information geometry can be generalized to this case. If a locally convex topology can be established, the further goal is to extend this to the infinite-dimensional case.

Another goal is to recover the geometric structure of the state spaces of both classical and quantum mechanics from the entropic structure, to formally show the equivalence. For quantum mechanics, one can simply define the Born rule from the entropy as we have seen before. For classical mechanics, the situation is a bit more complex as we need to recover first the Liouville measure. This can be done by noting two things: for a uniform distribution with support $U \subseteq X$, $\mu(U) = 2^{S(\rho_U)}$, and the uniform distribution over $U$ has the highest entropy of all distributions with support $U$. Given a set $A \subseteq \Ens$ of ensembles, we can then define the state capacity as the supremum of the exponential of the hull of $A$, which physically means the highest entropy achievable with a statistical mixture of elements of $A$. This set function has nice properties: it is non-negative, monotonic, subadditive in general and additive on orthogonal sets of ensembles. Note that if $A_U \subseteq \Ens$ is the subspace of all probability measures with support $U \subseteq X$, the state capacity of $A_U$ recovers the Liouville measure of $U$. Moreover, $A_U$ is orthogonal to $A_V$ if and only if $U$ and $V$ are disjoint. This means that the state capacity restricted to the lattice of those subspaces is an additive measure. Interestingly enough, one can follow the same construction in the quantum case, and the measure will return the dimensionality of the subspace. Concerning non-additive measures, there are additional interesting questions as one can create a non-additive version of probability measures, which may point to a non-additive version of probability theory. We have already looked at this problem, and, unfortunately, the standard techniques that have been developed in that area are not sufficient to handle the cases of interest to us within physical theories.

A possible fourth axiom may be needed to conclude all the requirements for physical theories. In both classical and quantum mechanics, there is a Lie algebra that characterizes the relationships between generators of time evolution. In classical mechanics, this is captured by the Poisson bracket, while in quantum mechanics it is captured by the commutator. The unsolved difficulty is that these structures are defined on the state space and not on the ensemble space. While there are attempts in this direction~\cite{caticha2008,HallReginattoBook}, we have found no general solution in the literature. Since the set of the probability measures defined on an odd-dimensional manifold satisfies the previous axioms, at least in the current form, it may be that the previous three axioms are not sufficient since no symplectic structure can exist over odd-dimensional manifolds. However, there is an intimate relationship between this Lie-algebraic structure and the entropic structure, since in both cases, the Lie-algebraic structure describes the relationship between the generators of transformations that leave the geometric structure (i.e.~symplectic form and inner product respectively) unchanged.  This means, as outlined before, they must leave the entropic structure unchanged. That is, it describes the group of deterministic and reversible motions, those that leave the entropy unchanged~\cite{aop-HamConsInfoEnt,aop-HamQuantInfo,aop-book}. Part of the investigation to identify the correct specification of the entropic structure, then, will need to consider its impact on the definition of the Lie-algebraic structure. For example, in our reverse physics approach to classical mechanics~\cite{aop-book,Carcassi:2022bpm}, we were able to show that a symplectic structure is necessary and sufficient to define an entropy that is invariant under frame transformations. If such a  requirement could be specified in general, the principle of relativity at equal time could be shown to be equivalent to this structure.

\noindent
\textbf{Expected outcomes and significance:}
During the three-year period of the proposed research, we expect to achieve \textcolor{blue}{and submit manuscripts on} at least the following:
\begin{itemize}
	\item Demonstrate that the Shannon entropy can be recovered as the maximum entropy increase under mixture for all ensemble spaces, and that the Shannon, Gibbs and von Neumann entropies can be recovered in the specific cases of discrete classical, continuous classical, and quantum mechanics.
	\item Show that the geometric structures of classical and quantum mechanics can be recovered from the entropic structure alone.
	\item Demonstrate that information geometry can be generalized to the ensemble space of any physical theory of finite dimensions. 
	\item Show that the entropic structure imposes constraints on the space of possible physical theories and rules out some Generalized Probabilistic Theories.
\end{itemize}
The above expected results should lead naturally to a range of additional possible directions and questions that could be addressed during or subsequent to the grant period, depending on the rate of advancement.  For example:
\begin{itemize}
	\item Identify what results from information geometry and information theory could be generalized to ensemble spaces; this would lend itself to various self-contained contributions by the grad student, undergrads, or individuals outside the project.
	\item Determine whether the convexity of the mixing entropy can be physically motivated and whether that guarantees a continuous embedding into a locally convex topological vector space.
	\item Generalize information geometry to the infinite-dimensional case; this would require a locally convex topology to be able to define variations.
	\item Determine whether the ensemble spaces can be in general proven to be metrizable and characterize the connection between the metric and the entropy.
	\item Find a suitable function space that can be used as a model for the case of classical field theories.
	\item Generalize classical Hamiltonian mechanics and the principle of stationary action to a generic ensemble space.
\end{itemize}

\textcolor{blue}{The immediate significance of the proposed research stems from the realization that, though classical information theory and quantum information theory are presented as two separate subjects in textbooks, there are a lot of commonalities that make them an interesting target for generalization.  
The broader significance lies in its approach as well as its specific potential to provide a common axiomatic foundation for \emph{all} physical theories.  Currently within physics, we typically do not know from where the various mathematical structures emerge, making it unclear in which directions they can be generalized in order to develop new theories that are physically meaningful.  Our approach instead seeks to \emph{find axioms from physical requirements}, i.e.~the axiomatics are justified from the physics.  Such an axiomatic foundation common to all physical theories would serve not only to better ground current theories and elucidate their relationships, but also to constrain potential future physical theories.}

%We're checking that the axioms are independent of each other.

While there exists other work attempting to reconstruct quantum mechanics from physical principles, for example in Refs.~\cite{hardy2001quantumaxioms,caticha2008,dakic2009,Chiribella2010,Masanes_2011}, we are not aware of any literature attempting to relate the geometric and entropic structures of either classical or quantum mechanics, much less those of a general physical theory. Works trying to reconstruct quantum mechancis from physical premises are typically in finite dimensions, do not generalize to classical mechanics, and do not provide a mathematically rigorous path to generalize to classical or quantum field theories.

\noindent
\textbf{Summary of proposed research program:} As we have seen, the program is rich and connects to many areas of physics, information theory and mathematics. Let us reiterate the primary goals. We want to mathematically formalize the intuition that specifying the geometry in physics is equivalent to specifying the entropy. To do so, we need to give a precise, physically motivated, axiomatic definition of what an entropic structure is for a general physical theory. The fundamental grounding concept will be the notion of statistical ensemble, as we can argue that any physical theory must \emph{at least} provide us with a space of statistical ensembles. All axioms will be stated as properties of ensemble spaces. Then we must show that the entropic structure characterizes the geometry of the general space in a way that recovers the geometry of classical and quantum mechanics in these specific cases. While pursuing this work, we will gain insight on how to handle topological issues properly, and hopefully show that ensemble spaces must embed into a second-countable metrizable locally convex vector space, so that the notion of differentiability can be extended to the infinite-dimensional case. We will also keep an eye open to connections to measure theory and Lie algebra, so that more features of physical theories can be generalized in the future.

\noindent
\textbf{Assessment of intellectual merit:} The proposed research program will be evaluated by the progress made in achieving the stated goals and outcomes and in disseminating the results to technical and academic audiences.

\noindent
\textbf{Relationship to other current funding:} PI Aidala's current funding from the NSF Division of Physics is for work on LHCb experiment at CERN, and from the NSF AccelNet program is to increase collaboration on quantum chromodynamics (QCD) research between the U.S.~and Latin America.  The PI's funding from the Department of Energy is for work at the Relativistic Heavy Ion Collider (RHIC) on the PHENIX and sPHENIX experiments.  These grants are for a completely separate line of research. Her grant from the John Templeton Foundation, titled ``Why classical probability and classical information theory are incompatible with quantum mechanics and quantum contextuality,'' supports the development of novel non-additive measures (set functions) suitable for quantifying the number of states in quantum mechanics~\cite{aop-nonaddmeas}.  There is no direct overlap with the activities proposed here.  Co-PI Uribe does not have any current funding.

\noindent
\textbf{Relationship to other commitments and activities of the PI and Co-PI:} PI Aidala has been continuously involved in high-energy experimental QCD since 2001.  Her separate line of theoretical research on the mathematical foundations of physics is relatively new, with her first publication in 2018~\cite{aop-phys-blueprint}. She has successfully split her efforts between her different research lines for the past seven years, with her work related to the Assumptions of Physics program producing 15 peer-reviewed publications, 1 more submitted paper, and a published technical book since 2018.  She is moreover in the process of reducing the number of personnel she supervises on QCD research in order to have more time for her theoretical work---four of her Ph.D.~students in QCD completed their degrees in the last 18 months, and two more are expected to graduate before the start of the proposed project.

Co-PI Uribe has an active research program studying the semi-classical limit in various geometric contexts, currently with a focus on phase spaces with a K\"{a}hler structure.  This limit relates the mathematics of quantum and classical mechanics, as Planck's constant tends to zero.
There is a similarity between Uribe's research program and the research proposed here, as the latter unifies the formalisms of classical and quantum mechanics.  Uribe is very interested in exploring this similarity, as it may give rise to new perspectives and suggest new theorems.

\subsection{Broader impacts}
In addition to disseminating the goals and results of the project through traditional means such as peer-reviewed publications and presentations at workshops and conferences, the goals and results of the project will be disseminated to wider audiences by leveraging existing social media and other outreach infrastructure that has already been developed for the broader Assumptions of Physics program, in particular, Carcassi's two YouTube channels, as described below in the section on qualifications of the research team.  One of them is aimed at the level of individuals with upper-level undergraduate training in physics and/or mathematics, and the second, more technical channel is aimed at professional researchers, for which he creates videos presenting detailed technical results or describing open technical questions, leveraging the expertise of others to help advance research goals.  Regarding the impact of posting open technical problems, our first collaborative paper that resulted from someone reaching out to us after seeing a video, on classical mechanics as the high-entropy limit of quantum mechanics, is currently under review at a journal~\cite{aop-classicallimit}.  Aidala and Carcassi were previously completely unaware of the researcher and his work at the University of Innsbruck, Austria. Carcassi's August 2025 video on ``The need for physical mathematics'' has already received 49k views and led directly to an invitation to speak in a special session at the Joint Mathematics Meeting in Washington, DC in January 2026. Carcassi additionally writes monthly blog-style essays related to the AoP program and advertised through social media platforms such as Facebook and LinkedIn.  Thus, the results of the proposed research will reach a much wider audience than through the publications and conference talks alone.  Results from this project will moreover be added to a future edition of Aidala and Carcassi's open-access book \textit{Assumptions of Physics} published online through Michigan Publishing~\cite{aop-book}. As part of making results accessible to others as soon as possible and leveraging input and expertise from others, real-time compilation from the public github repository of the latest draft of the book is set up on the AoP website~\cite{aop-website}. Further dissemination of the open-research approach may inspire researchers in completely different areas to adopt a similar approach for their own projects, with potential acceleration of research progress in particular within interdisciplinary areas, where not all necessary expertise may be available within a single research community.

As part of this project, Carcassi and Aidala will host an annual online school that includes an introduction to generalized ensemble spaces and the mathematical relationships between geometric and entropic structures, showcasing various results from the proposed work in a didactic manner.  The school will consist of six hour-long lectures, two per day over three days, each followed by 30 minutes of questions and discussion, from 10 am - 1 pm U.S. Eastern Time, enabling individuals from across the Americas, Europe, Africa, and parts of Asia to participate at a reasonable hour.  The school will be advertised primarily through social media.  Registrants will be given a Zoom link for participation, and those who choose not to register will be able to follow a livestream on YouTube (on Carcassi's technical channel) and ask questions through the livestream chat.  There will be no fee to participate.  All lectures and discussion will also be recorded and available through YouTube later. With 27 registrants for the inaugural Assumptions of Physics online school in June 2024, this grew to 167 registrants in June 2025, mainly undergraduate and graduate students and recent graduates in physics and/or mathematics, but also a handful of professors and other senior researchers.  Among the registrants were individuals from India, Bangladesh, Pakistan, Ethiopia, Egypt, Sudan, Bolivia, Brazil, and Singapore, in addition to North America and Europe. During the period of proposed work, we expect the school to reach several hundred individuals across similar demographics.

Beyond the immediate research goals, the proposed work will serve to unify what are currently disparate branches of physics---classical and quantum, and aspects of statistical mechanics---while making them more mathematically rigorous.  Developing a common mathematical framework will pave the way for eventual new approaches to physics education, with less compartmentalized teaching across the core advanced undergraduate courses.  Finding general mathematical structures that any physical theory must have will also help constrain the search for new physical theories.

\noindent
\textbf{Assessment of broader impacts:}  The reach and impact of videos related to the proposed work will be assessed by the number of views they receive and by other measures of audience engagement such as comments posted or subsequent email contacts, as well as potential talk invitations or proposals for collaboration.  The reach and impact of the school will be assessed by the number of registrants and their demographics, as well as the subsequent views of the recorded lectures.  The impact of developing a mathematical framework unifying disparate branches of physics will need to be assessed on a longer-term timescale than that of the grant, but early indicators could include talk invitations and paper citations. \\

\section{Personnel}
The proposed research will be performed by PI Christine Aidala (U-M Physics), Co-PI Alejandro Uribe (U-M Mathematics), staff researcher Gabriele Carcassi (U-M Physics), and a Ph.D.~student to be recruited from the Applied and Interdisciplinary Mathematics (AIM), Physics, or Mathematics Ph.D.~program.  See also the section on ``Qualifications of the Research Team'' below.  \\

Aidala as PI will be responsible for coordinating all of the work.  She and/or Uribe will serve as the formal academic (co-)advisor for the Ph.D.~student.  Students in the AIM program are required to have a primary advisor in Mathematics and a co-advisor in another Department.  Aidala and Uribe will identify suitable journals for publication and conferences at which to disseminate the results, and oversee the direction of the work.  Uribe has served as a mathematical consultant for Aidala and Carcassi's broader Assumptions of Physics program since 2018 and has become more involved since 2024, in particular regarding the efforts to develop a generalized theory of ensemble spaces for all physical theories. He is bringing mathematical expertise and will be able to contextualize the results of the project with respect to other areas within mathematical physics.  He is interested in what mathematical research questions this project may generate and what interesting mathematical structures may emerge.

Carcassi will dedicate 20\% of his time to the project and has been the individual most immersed in the technical details of the preliminary work leading to this proposal.  He is (currently part-time) staff in the U-M Department of Physics, a Research Area Specialist Senior, with his salary fully supported through grants.  He, Uribe, and Aidala will together identify the most promising technical directions, he will take the lead in performing a significant fraction of the work, and he will connect the results of the project to broader results so that the impact of the research is maximized.  With formal training in computer science and engineering, he will additionally serve as the primary contact to interface with others in these disciplines who may be interested in the work. Aidala, Uribe, and Carcassi, and once identified, the Ph.D.~student, will generally meet at least weekly to discuss the project.

Depending on the interests, background, and skills of the Ph.D.~student, there will be different ways for the student to contribute to the research. One possibility is to analyze what constructions and theorems from topological vector spaces can be generalized to topological convex spaces. This will help us better understand the relationship between the two, and possibly give insights into what properties the entropic structure may impose to bridge the gap. Another possibility is to see what results from classical and quantum information theory can be generalized to the entropic structure alone. Yet another possibility, since the JSD is the square of a distance function, is to identify sufficient conditions for a pseudo-distance to be the square of a distance, which may clarify what properties the entropic structure imposes. Developing proficiency in particular in information theory and information geometry will lead to a wide range of possible next steps in the student's career after their Ph.D., with applications for example in quantum computing and in machine learning across academia and the private sector.  We additionally anticipate involving one or more undergraduates in the project, identifying self-contained pieces of the project to which they could contribute.  For example, starting from a few specific results from information geometry or information theory, they could work on trying to generalize these results to ensemble spaces.

%\begin{figure*}[htb]
%	\centering
%	\includegraphics[scale=.6]{figures/PersonnelTimelineNSF2022.png}
%	\caption{Estimated timeline of research activities.}
%	\label{fig:personnelTimeline}
%\end{figure*}

\section{Qualifications of the research team}
PI Christine Aidala is a highly recognized leader within physics, having received an NSF CAREER Award, a Sloan Fellowship, and a Presidential Early Career Award for Scientists and Engineers. She was a 2019-20 U.S.~Fulbright Scholar and is a Fellow of the American Physical Society (APS).  She served on the National Academy of Sciences Committee on a U.S.-Based Electron-Ion Collider Science Assessment 2016-18 and on the U.S.~Nuclear Science Advisory Committee 2023-25.  Much of her work has been in quantum chromodynamics; however, in 2015 she initiated the Assumptions of Physics program with her collaborator and husband Gabriele Carcassi, and since 2018 they have so far published together 15 peer-reviewed articles ~\cite{aop-phys-blueprint,aop-topExpDisting,aop-HamQuantInfo,Carcassi:2021,aop-spacetimeStruct,aop-HamConsInfoEnt,Carcassi2021four,Carcassi:2022bpm,aop-HamPriv,aop-commonLogical,aop-action,aop-nogo,aop-nonaddmeas,aop-nogoresponse,aop-unphysHilbert} as well as a technical book on the mathematical foundations of physics~\cite{aop-book}, with one additional article currently under review and publicly available as a preprint~\cite{aop-classicallimit}.  She is the PI on a grant from the John Templeton Foundation that supports the development of novel measure-theoretic structures for quantum mechanics.

Co-PI Uribe is a Fellow of the American Mathematical Society, with a long track record of publication in first-rate journals going back decades.   He has held the positions of Professeur Invit\'{e} at the \'{E}cole Normale Sup\'{e}rieure (Paris) and at the Universit\'{e} de Paris-Saclay, and of Research Professor at MSRI (now the Simons Laufer Mathematical Sciences Institute).  He was the PI on various NSF grants (DMS) from 1989 to 2014.

Carcassi is the co-initiator with Aidala and technical lead for the Assumptions of Physics program.  He is the Co-PI on the Templeton Foundation grant with Aidala.  Related to the AoP program, he has given seminars to audiences of physicists, mathematicians, and philosophers at places including Harvard, Caltech, Oxford (UK) the University of Pavia (Italy) the Munich Center for Mathematical Philosophy (Germany) and the Wolfram Institute.  He and Aidala gave a joint public talk on misconceptions in the foundations of physics for the Michigan State University Advanced Studies Gateway, and he will give a U-M Saturday Morning Physics public talk on ``What is classical and what is quantum?'' in November 2025.  He was invited and will deliver an overview talk on the AoP program at a special session of the Joint Mathematics Meeting in January 2026.  His training in computer science and engineering allows him to bring ideas from fields not typically studied by individuals trained primarily in mathematics or physics, for example from information theory, theoretical computer science, systems theory, signal processing, and operations research.  His position as U-M Physics Department staff without teaching responsibilities provides him with time and flexibility to dedicate to research related to the AoP program such as the work proposed here. See also his biographical sketch included as an Other Supplementary Document.\\

Together, Aidala, Uribe, and Carcassi are uniquely positioned to advance the mathematical foundations of physics, with Aidala's and Uribe's traditional training in physics and mathematics, respectively, complementing Carcassi's formal training in computer science and engineering.  Carcassi's university staff position has allowed him to audit a number of graduate Mathematics and Physics courses at U-M, after having previously audited multiple graduate Physics courses at Stony Brook University.  Uribe has served as a mathematical consultant to Aidala and Carcassi for the broader Assumptions of Physics program since 2018.  For the AoP program, Aidala and Carcassi's published work so far has covered a wide range of topics.  The ``physical mathematics'' approach within their work is an approach to the mathematical foundations of physics that seeks to construct mathematical structures strictly from axioms and definitions that can be rigorously justified from physical requirements~\cite{aop-book}.  For example, they have shown that experimental distinguishability is mathematically captured by Hausdorff topologies~\cite{aop-book,aop-topExpDisting}, and they have argued that at the smallest space-time scales, at which geometric information becomes experimentally inaccessible, there should still in principle be a regime in which topological structure remains experimentally accessible~\cite{aop-spacetimeStruct}.  Their ``reverse physics'' approach is instead an approach to the foundations of physics that analyzes known theories to identify those physical principles and assumptions that can be taken as their conceptual foundation~\cite{aop-book,Carcassi:2022bpm}. Their reverse physics approach has led for example to a geometric and physical interpretation of the principle of least action~\cite{aop-action} and to a demonstration that the uncertainty principle in quantum mechanics emerges solely from the lower limit on the entropy~\cite{Carcassi:2022bpm}. In other work, they have for example reduced the number of postulates of quantum mechanics, proving that the tensor product postulate for composite systems can be derived from the state and measurement postulates~\cite{Carcassi2021four}, and they have offered a clearer characterization of Shannon information entropy~\cite{Carcassi:2021}.

Regarding success in mentoring and education, in 2019 Aidala received the U-M Imes and Moore Mentorship Award ``for exceptional contributions toward recruiting and mentoring graduate students in the sciences from disadvantaged and nontraditional backgrounds,'' and in 2025 the U-M Rackham Distinguished Graduate Mentor Award.  She has served as the dissertation chair for 10 graduate students in Physics and Applied Physics who have already completed their doctorates, as co-chair for another in Applied Physics and as co-chair with Uribe for one in Applied and Indisciplinary Mathematics.  She is currently the dissertation chair for an additional 4 Ph.D.~students in Physics.  Aidala and Carcassi have also successfully published with former U-M Ph.D.~students in Mathematics~\cite{aop-topExpDisting} and Philosophy~\cite{aop-HamPriv,aop-unphysHilbert} whose primary dissertation work was with other faculty members.  In 2020 Aidala received the U-M John Dewey Award for long-term commitment to the education of undergraduate students.  Between Aidala and Carcassi, they have mentored more than 50 undergraduate students in research since coming to U-M in 2012, with Carcassi serving as the primary mentor for three undergraduate senior theses related to the AoP program.  One of Carcassi's former undergraduate research mentees is now pursuing a Ph.D. in Math at UT Austin.

Uribe has graduated six Ph.D.~students and has mentored 13 postdoctoral assistant professors.  He has directed two Master's theses in the U-M Mathematics bridge program (MLB Scholars), as well as numerous REU projects. He has additionally served on the dissertation committee of four of Aidala's students thus far.

Regarding the qualifications of the research team to achieve the broader impacts beyond academia as described above, Carcassi has a YouTube channel (@gcarcassi) focused on physics and mathematics with 18k subscribers currently.  He posts a new video monthly, generally targeted at individuals with an advanced undergraduate background in physics and/or mathematics.  Collectively, his videos on this channel have received over 630k views. He additionally has a second, more technical YouTube channel (@AssumptionsofPhysicsResearch) aimed at professional researchers, with more than 1600 subscribers.  He uses this channel as a venue to disseminate research results at a more technical level, including video versions of preprints or published papers as well as conference presentations related to the AoP program.  He moreover uses the technical channel as an innovative approach to try to ``crowd-source'' technical expertise to advance the research more efficiently.  Given the breadth of the broader AoP program, there is no single community with the appropriate technical background across all relevant and potentially relevant areas of mathematics and physics.  So far this ``crowd-sourcing'' approach has led to multiple instances of viewers with whom we have had no previous contact pointing us to valuable reference materials, and in 2024 it led to our first coauthored paper with a researcher who reached out to us after seeing an open problem video, on ``Classical mechanics as the high-entropy limit of quantum mechanics''~\cite{aop-classicallimit}. We were previously completely unaware of the researcher and his work in Innsbruck, Austria.  In other forms of reach beyond academia, Carcassi was interviewed in 2024 by Curt Jaimungal for the ``Theories of Everything'' podcast~\cite{Carcassi-ToEInterview}, which has more than 500k subscribers.  Carcassi additionally writes monthly blog-style essays related to the AoP program and advertised through social media platforms such as Facebook and LinkedIn.  In more creative outreach work, in April 2024 at U-M he performed an original musical comedy lecture, ``An Engineer in the Foundations of Physics,'' which drew approximately 100 attendees. He recently performed a slightly updated version of the comedy lecture at the American Physical Society meeting in Chicago in October 2025.

\section{Results from prior NSF support}
\subsection{Intellectual merit of prior NSF support}
Aidala has been a successful steward of NSF funds since receiving a CAREER Award in 2015.  Her prior research awards have been through the Physics Division, Nuclear Physics - Experiment program: \emph{CAREER: Valence and sea quark dynamics at Fermilab} (Award Number 1452636, Jul 2015-Jun 2021, \$738,082), and \emph{Studying quantum chromodynamics at LHCb} (Award Number 2012926, Jul 2020-Jun 2023, \$697,969; renewal Award Number 2309922, Jul 2023-Jun 2026, \$736,406).  She is additionally a Co-PI on an NSF AccelNet award, \emph{AccelNet-Design: Inter-American Network of Networks on Quantum Chromodynamics Challenges} (Award Number 2114669, Sep 2021-Aug 2026, \$38,556 allocated to U-M), aimed at further developing collaboration in quantum chromodynamics research between the U.S.~and Latin America.  Her QCD research has focused on advancing the multidimensional quark-gluon structure of the proton and on studying the processes by which scattered quarks or gluons form new bound states (hadrons).  Her group's NSF-funded work includes results published in Nature~\cite{SeaQuest:2021zxb}, Physical Review Letters~\cite{LHCb:2022tbc,LHCb:2019qoc}, Physical Review D Letters~\cite{LHCb:2022rky}, Physical Review D~\cite{LHCb:2025mcq}, Physical Review C~\cite{SeaQuest:2022xdu}, Physics Letters B~\cite{LHCb:2025tvf,SeaQuest:2024qdw}, and Computing and Software for Big Science~\cite{Aidala:2023dai}, among other journals~\cite{SeaQuest:2023tcr,SeaQuest:2017kjt}, and it includes one additional paper submitted for review~\cite{LHCb:2025rxf}.  Regarding the specific scientific advancements of some of the work, Aidala and her collaborators have studied antimatter in the proton and found surprising evidence that there is an excess of down ``flavored'' antiquarks relative to up flavored antiquarks in the proton for antiquarks carrying as much as 45\% of the proton's total momentum~\cite{SeaQuest:2021zxb,SeaQuest:2022xdu}.  Her group also led the first-ever measurements of identified hadron species production by high-energy light-mass quarks in proton-proton collisions, multidifferential in multiple kinematic variables simultaneously, showing that existing models poorly predict the relative production rates of some species~\cite{LHCb:2022rky}, and recently discovered striking evidence of suppressed gluon radiation from beauty (bottom) quarks due to their large mass~\cite{LHCb:2025mcq}. In the spirit of making research more open, her group has championed and led the open data efforts of the 1200-scientist LHCb collaboration~\cite{Aidala:2023dai}.  This proposal is the second time she is submitting to the Applied Mathematics program, based on the research direction in the mathematical foundations of physics that she initiated more recently, which has so far produced 15 peer-reviewed papers since 2018 as well as a technical book, as described above in the section on qualifications of the PI and research team.  Uribe was the PI on various NSF grants from 1989--2014.

\subsection{Broader impacts of prior NSF support}
Aidala's group has been leading the open data efforts of the LHCb experiment at the Large Hadron Collider at CERN.  Of particular note, they have developed a user-friendly web-based application to securely access and provide analysis documentation for large-scale data sets in high-energy physics, the ``NTuple Wizard''~\cite{Aidala:2023dai}.  It is currently in beta-release to provide access to open data sets from LHCb, and individuals from other experimental collaborations have expressed interest in potentially adapting it to create a similar application to facilitate access to their own open data sets and corresponding documentation.

In addition, a variety of activities under Aidala's prior NSF grants have contributed directly to training junior individuals from a wide range of backgrounds for careers in STEM.  Including individuals with support from on and off of the grant, 3 postdocs, 8 Ph.D. students and 6 undergrads have been involved in her NSF-funded projects since July 2020. An additional 3 distinct Ph.D. students and 7 undergraduates worked on the projects supported by Aidala's CAREER grant 2015-20.  Her former grad students have gone on to permanent staff scientist positions at Fermi National Accelerator Lab and Brookhaven National Lab, a faculty position at Nantes Universit\'{e} (France), a Lecturer position at U-M, the Head of Special Projects and Research Manager at Entanglement, Inc., and to found their own consulting firm, with others in postdoctoral positions.

\newpage